### 2023-11-14 rmds/figures-supp.Rmd
refresh

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")

# Load results
source(here::here("code", "setup.R"))

data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples)))
```

## Figures
```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "m50")) +
  geom_point(aes(as.factor(Nsamples), conv_success, col = repro_freq, group = repro_freq)) + 
  geom_line(aes(as.factor(Nsamples), conv_success, col = repro_freq, group = repro_freq)) +
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) +
  scale_color_brewer(palette = "Spectral") + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Number of samples", y = "Convergence success (%)", title = "Model convergence") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S1.

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "m50")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative error", title = "Bias - Length at 50% maternity") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")

```
Figure S2.

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "c")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative error", title = expression("Bias - "~italic(P)[Max])) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")

```
Figure S3. Note 2PLF-maturity and 3PLF-fixed methods are identical

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "m50")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = "Precision - Length at 50% maternity") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S4.

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "c")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - "~italic(P)[Max])) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S5.
\newpage

```{r fig.width = 12, fig.height=7}
ggplot(data %>% filter(par_name %in% "m50", method_long %in% c("3PLF - free", "3PLF - fixed"))) +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "% interval coverage", title = "Interval coverage - Length at 50% maternity") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S6.
\newpage

```{r fig.width = 12, fig.height=5}
ggplot(data %>% filter(par_name %in% "c", method_long %in% c("3PLF - free"))) +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "% interval coverage", title = expression("Interval coverage - "~italic(P)[Max])) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S7. 
\newpage

```{r, fig.width = 11, fig.height = 8.5}
filter(data, method_long %in% c("3PLF - free", par_name %in% "c")) %>% 
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = n_maternal)) + 
  geom_text(aes(x = as.factor(Nsamples_sel), y = repro_freq, label = round(n_maternal, 0))) +
  scale_fill_viridis_c(trans = "log10") + 
  guides(fill = guide_colorbar(title = bquote(atop("Median", "n. maternal")))) + 
  facet_wrap(~species, nrow = 2) +
  theme_half_open() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure S8. Median number of females in maternal condition in simulated data that were used to estimate maternity parameters subject to varying sample size, gear selectivity, reproductive frequency. 250 iterations were carried out for each simulation. 
\newpage

```{r, fig.width = 11, fig.height = 8.5}
filter(data, method_long %in% c("3PLF - free", par_name %in% "c")) %>% 
ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = n_immature)) + 
  geom_text(aes(x = as.factor(Nsamples_sel), y = repro_freq, label = round(n_immature, 0))) +
  scale_fill_viridis_c(trans = "log10") + 
  guides(fill = guide_colorbar(title = bquote(atop("Median", "n. immature")))) + 
  facet_wrap(~species, nrow = 2) +
  theme_half_open() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure S9. Median number of immature females in simulated data that were used to estimate maternity parameters subject to varying sample size, gear selectivity, reproductive frequency. 250 iterations were carried out for each simulation. 
\newpage

```{r fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "r_0")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative error", title = "Bias - Reproductive Output") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```

```{r fig.width = 12, fig.height = 10}
  ggplot(data %>% filter(par_name %in% "r_0")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative absolute error", title = "Precision - Reproductive Output") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```

### 2023-11-14 rmds/figures.Rmd
refresh

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load results
source(here::here("code", "setup.R"))

```

## Figures

```{r out.width = '100%', fig.cap = "Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. "}
knitr::include_graphics(here("img","figure1.png"))
```

\newpage
```{r fig.width = 7, fig.height= 5.5}
data <- read_rds(here("data", "simulations-pivot.Rds"))

ggplot() + 
  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
  scale_color_viridis_d(option = "D", direction = 1) + 
  facet_grid(species~method_long) + 
  ylim(-30, 50) +
  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
```

Figure 2.  Per cent relative error (bias) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods. Each point represents parameter estimates one iteration of simulated data. Bias tended to increase as the true underlying $P_{Max}$ decreased. 
\newpage

```{r}
data <- read_rds(here("data", "simulations-slim.Rds"))

data <- data %>% 
    mutate(method_long = case_when(
    method %in% "2" ~ "2PLF - maturity",
    method %in% "2a" ~ "2PLF - maternity",
    method %in% "3" ~ "3PLF - free",
    method %in% "3a" ~ "3PLF - fixed"
  )) %>%
  mutate(mesh_name_long = case_when(
    mesh_name %in% "high" ~ "High selectivity",
    mesh_name %in% "low" ~ "Low selectivity"
  )) %>% 
  filter(mesh_name %in% "high", par_name %in% "r_0") %>% 
  filter(Nsamples %in% c(50, 250, 1000))

data <- data %>% 
  mutate(method_long = fct_relevel(
    method_long, "3PLF - free",
    "3PLF - fixed",
    "2PLF - maternity",
    "2PLF - maturity"
  ))

ggplot(data) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(aes(x = Nsamples, y = abs_error, fill = repro_freq), outlier.shape = NA) +
  facet_grid(method_long ~ species) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(fill = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Median % absolute error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  ylim(0, 80)
```

\newpage
```{r fig.width = 10, fig.height= 7}
# Load data - Sandbar
data <- read_rds(here("data", "empirical-plumbeus.Rds"))
preds <- read_rds(here("data", "empirical-predictions.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 220, 8)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - free") %>% 
  mutate(label2 = "italic('C. plumbeus ')~`- 3PLF  free`")

p1 <- ggplot()  + 
  geom_ribbon(data = preds1, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds1, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "italic('C. plumbeus')~` - 3PLF fixed`")

# Plot
p2 <- ggplot()  + 
  geom_ribbon(data = preds2, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds2, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

# Load data - blacktip
data <- read_rds(here("data", "empirical-limbatus.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 160, 8)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds3 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - free") %>%
  mutate(label2 = "italic('C. limbatus')~` - 3PLF free`")

# Plot
p3 <- ggplot()  + 
  geom_ribbon(data = preds3, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds3, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

preds4 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "italic('C. limbatus')~` - 3PLF fixed`")

p4 <- ggplot()  + 
  geom_ribbon(data = preds4, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds4, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL)  + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

grid.arrange(p3, p4, p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 16, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 16, font = 8)))
```
*Figure 3. Comparison of 3PLF-free and -fixed methods used to estimate maternal parameters for blacktip shark, *C. limbatus*, and sandbar shark, *C. plumbeus* in the Gulf of Mexico and Western North Atlantic. Sold line is the expected proportion in maternal condition at fork length. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. A fixed value of $P_{Max} = 0.5$ was used for both species.*

### 2023-11-14 rmds/manuscript.Rmd
refresh

---
title: |
  Quantifying maternal reproductive output of chondrichthyan fishes
subtitle: |
author:
- name: Alastair V Harry
  affiliation: Fisheries & Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia
  email: alastair.harry@gmail.com
  number: 1
- name: Ivy Baremore
  affiliation: MarAlliance
  number: 2
- name: Andrew Piercy
  affiliation: Valencia College
  number: 3
abstract: |
  For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to calculate the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, relatively few studies report maternity functions. Instead, maturity functions have typically been used as a proxy, despite evidence of a delay between attainment of functional maturity and pregnancy in some species. This study examined the relative performance of alternative approaches for quantifying maternal reproductive output. Applying logistic models to a combination of simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used. At sample sizes < 250 individuals a fixed-value for $P_{Max}$ was more effective at reducing bias. Maturity parameters could be estimated with greater precision however substituting them in place of maternity parameters overestimated lifetime reproductive output. The extent to which maturity functions can provide a good approximation for maternity functions may be species specific and requires futher research. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. We proposed that estimation of $P_{Max}$ is generally preferable to using a fixed value. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.

toc: false
output:
  word_document: default
  bookdown::pdf_document2:
    pandoc_args:
      - --csl=extras/cjfas.csl
      - --citation-abbreviations=extras/abbreviations.json
      - --template=extras/mytemplate.tex
bibliography: "extras/maternal-output.bib"
keep-latex: true
linestretch: 2 
header-includes:
  - \fontsize{12}{20}
  - \usepackage{lineno}
  - \linenumbers
  - \linenumberfont
  - \renewcommand\linenumberfont{\normalfont\small\sffamily}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{\thepage}
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Setup code
source(here::here("code", "setup.R"))

# Load results
empirical <- read_rds(here("data", "empirical-summary.Rds"))
plumbeus <- read_rds(here("data", "empirical-plumbeus.Rds"))

```

\newpage

# Introduction

Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass [@quinn_quantitative_1999]. Calculating reproductive capacity requires data on the sexually mature proportion of the population contributing to reproduction at a given size or age. For most populations of teleost fishes a simple two-parameter logistic regression function (2PLF) is sufficient to model this process using dichotomous sexual maturity stage data (immature / mature) from a sample of individuals [@jennings_marine_2001; @king_fisheries_2007]. While the mathematical formulae and methods of statistically estimating parameters for a 2PLF vary, the estimated values are similar and can be obtained with the built in generalised linear model (GLM) programs in most modern statistical software.

In chondrichthyan fishes (sharks, rays, skates, and chimaeras), the process of maturation can likewise be modeled with a simple 2PLF. However for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015]. Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017]. This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014]. Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stock-recruitment_2013]. This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

## Maternity function

Walker [-@walker_reproduction_2005] recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: $$Y_i \sim B(x_i; n_i, P_i)$$ $$E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x)}}$$ where the proportion of individuals in a given length or age category, $x_i$, that are in maternal condition, $Y_i$, is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, $P_i$, multiplied by the total number of individuals in the category, $n_i$. If $x$ is a continuous rather than categorical variable (i.e. $n = 1$), $Y_i$ reduces to a Bernoulli random variable [@zuur_analyzing_2007]. $\alpha$ is the upper asymptote of the curve, controlling the maximum value of $Y$ as $x$ approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when $\alpha = 1$, as is appropriate for most teleost and chondrichthyan populations if modeling maturity [@roa_estimation_1999; @quinn_quantitative_1999]. When $\alpha = 1$, the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows $Y$ to be modeled as a linear function of $x_i$, where $b_0$ and $b_1$ are the intercept and slope, respectively [@zuur_analyzing_2007]. Alternatively, the model can be expressed in more biologically relevant terms as: $$E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} $$ where $x_{50}$ and $x_{95}$ are the sizes or ages at which 50% and 95% of the maximum proportion of individuals ($P_{Max}$) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

Walker [-@walker_reproduction_2005] defines individual females as being in maternal condition if they would have given birth or laid eggs by the end of a given year such that they contribute to annual recruitment (age 0+ cohort) at the beginning of the next year. Pregnancy is a necessary, but not sufficient, condition for maternity, and the two conditions are not always synonymous. For example, reproduction in the Australian population of school shark, *Galeorhinus galeus*, occurs triennially, such that approximately one third of mature females give birth annually and $P_{Max} \approx 1/3$ [@walker_reproduction_2005]. Gestation lasts 20 months, and includes a protracted period of ovulation. Newly pregnant females are not considered to be in maternal condition until the year of pre-recruitment. Walker's [-@walker_reproduction_2005] definition is specific to species that reproduce during a single, well-defined period of the year, but the concept can be generalised to accommodate asynchronously reproducing species or those that reproduce more than once a year.

Numerous authors have adopted Walker's [-@walker_reproduction_2005] maternity function, including for species with biennial or longer reproductive cycles [@huveneers_reproductive_2007; @rochowski_reproductive_2015; @trinnie_reproductive_2016] as well as those with annual or shorter reproductive cycles [@trinnie_reproductive_2009; @mejia-falla_reproductive_2012; @harry_age_2013; @taylor_living_2016]. Techniques have been described for determining maternal output in asynchronously reproducing species [@braccini_determining_2006; @colonello_reproductive_2016], and maternity functions are also increasingly being used directly in shark and ray population assessments [@sedar_sedar_2012; @sedar_sedar_2017].

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions, potentially due to a range of factors. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles. Such data can be difficult to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

For practitioners that are aware of maternity functions, lack of information on implementation may also have discouraged use. Walker's [-@walker_reproduction_2005] approach to estimating parameters in the maternity function is difficult to reproduce, apparently due to the constraints of the proprietary statistical program first used to implement the analysis. Specifically, the method as described involves adjusting the raw data prior to parameter estimation and then weighting it during analysis [@walker_reproduction_2005]. Additionally, the use of a fixed value for $P_{Max}$, as in the original analysis, has the potential to bias parameter estimates and reduce their standard errors [@motulsky_fitting_2004]. Alternative approaches to estimating maternal parameters have also arisen in the literature, indicative of a lack of guidance on implementation. For example, some authors have obtained maternity parameters by fitting a 2PLF to maternity data [@baremore_reproduction_2012; @baremore_reproduction_2013].

A consequence of the low uptake in use of maternity functions is that most practitioners undertaking population assessments invariably take an *ad hoc* approach to quantifying maternal reproductive output. This typically involves approximating maternal output using a maturity function. Under such an approach all mature females are tacitly assumed to reproduce in each breeding season [@cortes_demographic_1998]. Non-annual reproduction is accounted for by weighting fecundity or the maturity function by the assumed (but often unknown) duration of the reproductive cycle [-@walker_reproduction_2005]. In some cases, calculations of reproductive output may use the age-at-first-reproduction, defined as the mean age at maturity plus the gestation period [@mollet_comparative_2002]. This accounts for the protracted gestation period of many chondrichthyans, but still assumes that all individuals begin reproducing immediately after maturity. Measures of annual reproductive output that follow this approach are hence built upon a series of assumptions that are rarely tested in practice.

## Revisiting the maternity function

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions [@cortes_population_2012]. It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g. annual, biennial), despite observations of plasticity in this trait within discrete populations [@higgs_reproductive_2020]. Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing $P_{Max}$ in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles. To date only a single study on spiny dogfish, *Squalus acanthias*, has statistically estimated $P_{Max}$ from data [@colonello_reproductive_2016], and the feasibility of doing so for a wider range of species has not been investigated.

This study revisits the use of maternity functions with the objectives of providing guidance on implementation and appropriate use. To address this objective a combination of simulated and empirical data were used to 1) evaluate the performance of two alternative methods for estimating maternity parameters, subject to varying reproductive frequency, sample size, and gear selectivity, 2) illustrate the effect of not using using maternity functions on calculations of lifetime reproductive output, and 3) outline strategies for making inferences on reproductive frequency from maternity data.

# Methods

## Approach

To assess the feasibility of estimating maternity parameters a simulation study was done using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1). Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs. The performance of three methods in estimating 'true' maternity parameters from simulated data was tested by examining precision, bias, and interval coverage. To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters. Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency. All simulation modelling and data analysis was conducted using the R language [@rcoreteam_r:_2018] and the Template Model Builder (TMB) R package [@kristensen_tmb:_2016].

```{r out.width = '100%', fig.cap = "Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. "}
knitr::include_graphics(here("img","figure1.png"))
```

## Simulation

### Data generation

Data were generated for the gummy shark, *Mustelus antarcticus*, and school shark, *Galeorhinus galeus*, using parameters from studies conducted on southern Australian populations (Table 1) [@grant_estimation_1979; @kirkwood_gill_1986; @moulton_age_1992; @walker_fishery_1992; @punt_stock_1998; @walker_reproduction_2005; @walker_spatial_2007]. Both species are well-studied and were chosen to be representative of r- and K-selected chondrichthyan life history styles, respectively [@stevens_variable_1999].

Catch at length data were simulated using a female-only, length-based catch curve model [@L3Assess]. The approach involved calculating expected survival and catches per recruit subject to specified life history parameters, gear selectivity, and total mortality. Growth was modelled using a von Bertalanffy growth function with length *l*, of a fish of age, *a*, calculated as $$l(a) = L_\infty (1- e^{-K(a-a_0)})$$

where $L_\infty$ is asymptotic length, *K* is the growth coefficient and $a_0$ is the hypothetical age at zero length. Recruitment into the population was assumed to occur at age 0 with length conforming to a normal distribution. The proportion of fish in length class, *l*, at age, $a = 0$, was calculated as $$\theta_k = \int_{l^-_k}^{l^+_k}f_{a=0}(l)dl$$ where $l^-$ and $l^+$ are the lower and upper limits of each 1cm length class, *k*. $f_{a=0}(l)$ is the value of the normal probability density function at age $a = 0$ for a given length, *l*,$$f_{a=0}(l) = \frac{1}{\sigma\sqrt{2{\pi}}}\exp\bigg[\frac{(l-\mu)^2}{2\sigma^2}\bigg]
$$

where mean length, $\mu$, was caclulated from the growth curve. The standard deviation, $\sigma$ was not available for either species so was assumed to be directly proportional to length, with a constant of proportionality of 5% (Table 1), i.e. $\sigma = 0.05 \mu$.

Growth in subsequent age classes was modelled using a length transition matrix [@punt1997; @hall2000], $G = g_{j,k}$ that represents the probability that a fish in length class, $j$, will grow into length class, $k$, over a specified time interval $$ G = \begin{bmatrix}
 g_{1,1}&0& 0 &0\\ 
 g_{2,1}& g_{2,2} & 0  &0 \\ 
 \cdots&\cdots  & \ddots  & 0\\ 
 g_{n,1}& g_{n,2} & \cdots & g_{n,n}
\end{bmatrix}$$

where the elements of the matrix followed the general form of Punt *et al* [-@punt1997] and are given by$$ g_{k,j} = \left\{
    \begin{array}{ll}
        \int_{{\infty}^-}^{\mathrm{l_{k}^+}}f[\phi(l,j)]dl & \mbox{if }k=1 \\
        \int_{l_{k}^-}^{\mathrm{l_{k}^+}}f[\phi(l,j)]dl &\mbox{if }k=1<k<n \\
\int_{l_{k}^-}^{\mathrm{\infty^+}}f[\phi(l,j)]dl &\mbox{if }k=n \\
    \end{array}
\right.
$$

where *l* is mean length, *f* is the specified normal distribution, $\mathrm{l_{k}^-}$ and $\mathrm{l_{k}^+}$ are the lower and upper limits of length class *j*, and $\phi$ is a vector of parameters.

Selectivity in the model was assumed to follow a gamma function based on gillnet selectivity experiments [@kirkwood_gill_1986; @punt_stock_1998], where the selectivity of an individual in length class, $l$ was given by $$S(l) = \left [ \frac{l}{\theta_1 \theta_2} \right ]^{\theta_1 }exp(\theta_1-\frac{l}{\theta_2}) $$ where $\theta_1$ and $\theta_2$ are estimated parameters [@kirkwood_gill_1986]. Two selectivity scenarios were simulated corresponding to low and high selectivity of the maternal component of the population. For consistency between species, gillnet mesh sizes were chosen that would result in peak relative selectivity occurring at the lengths equal to 5% and 95% of the population in maternal condition (Table 1).

Fishing mortality, *F*, in length class *l*, was calculated as $$F(l) = S(l) \cdot F$$where $F$ was assumed to be constant and nominally set to 25% of natural mortality, $M$ (i.e. $0.25M$). The effects of varying levels of $F$ were not investigated further in this study. Total mortality, *Z,* in length class, *l*, was further given by $$Z(l) = F(l)\cdot{M} $$

The number of fish, $N$, per recruit in length class, *l,* that survived to age, *a*, was calculated as

$$ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=1<a<A \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A \\    \end{array}\right.$$

The estimated numbers of fish caught in length class, l, at age, a, was calculated using the Baranov catch equation as

$$
\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
$$

To examine the effect of varying sample size, six scenarios were conducted with catch ranging from 50 to 2500 individuals, reflecting a gradient from data poor to rich.

Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This processes was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l(a) - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 250 iterations were run for each of the 144 unique variables combinations.

### Estimation approaches

Two approaches to estimating maternity parameters were compared, the 3PLF itself (3PLF-free) and the 3PLF function with a fixed asymptote (3PLF-fixed). Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF maturity). The 3PLF-free method was examined to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016]. The 3PLF-fixed is the method described by Walker [-@walker_reproduction_2005] and most commonly used in practice. The 2PLF-maternity method was used to examine what effect simply ignoring the upper asmyptote had (i.e. $P_{Max}$ fixed at 1). The 2PLF-maturity method is tacitly assumed if using mature females to approximate maternal ones. The resulting maturity curve can then by weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially).

Only the 3PLF-free method involves statistically estimating $P_{Max}$, which must be subjectively chosen in the case of the 3PLF-fixed and the 2PLF-maturity. $P_{Max}$ is ideally chosen based on detailed study of the ovarian and uterine cycles [@walker_reproduction_2005], or alternatively based on the proportion of mature females observed to be in maternal condition during sampling [@baremore_reproduction_2012; @harry_age_2013; @trinnie_reproductive_2016]. For the purposes of the simulation, the timing and duration of the uterine and ovarian cycles were assumed to be unknown. The following procedure was used to 'guess' the fixed value of $P_{Max}$ in each simulation. $P_{Max}$ was chosen as the proportion of females in maternal condition above a the length at which $99\%$ of females were mature, $L > \Psi(L_{99\%})$. If this failed (e.g. zero pregnant females), $L_{95}$ was used, followed by $L_{50}$, and finally the proportion of all mature females in the sample that were in maternal condition. Statistical estimation of all non-fixed parameters in each of the methods was undertaken using maximum likelihood. Nonparametric bootstrapping was used to calculate approximate 50% confidence intervals for the best-fit parameters from 250 resampled data sets.

### Reproductive output

For each iteration, the per-generation rate of multiplication, $R_0$, or lifetime female pup production, was calculated as: $$R_0 =\int_{0}^{\infty}R\beta(a)e^{\int_{0}^{a}M ds}da $$ where $R$ and $M$ are the embryonic sex ratio and natural mortality rate (Table 1) and $\beta(a)$ is the age-specific reproductive rate which incorporates $\Psi'(L)$ [@xiao_demographic_2000].

### Performance

Model performance was evaluated in each iteration where 1) the fitted model successfully converged as indicated by a positive definite Hessian matrix, and 2) the estimated parameters $L_{50}'$ and $L_{95}'$ remained within specified bounds ($0 - 10^4$). An iteration was also not attempted if there were no maternal females in the randomly generated sample. Model performance was evaluated by quantifying precision, bias, and interval coverage for $L'_{50}$, $P_{Max}$ and $R_0$. Bias was quantified using the mean relative error, $\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]$ and precision using the mean absolute relative error, $\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|$. Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50% confidence intervals [@rudd_accounting_2018]. For a well-performing model, approximately 50% of parameters would be expected to fall withing the nominal 50% confidence intervals.

### Empirical case study

An empirical case study was also undertaken to show how maternity functions can be used in practice. Data from two studies of the reproductive biology of the sandbar shark, *Carcharhinus plumbeus*, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods. The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{yr}^{-1}$) that was thought to have a biennial reproductive cycle [@springer_natural_1960]. The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s [@sminkey_age_1995] and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes [@baremore_reproduction_2012; @piercy_histological_2016]. In their study, Baremore and Hale [-@baremore_reproduction_2012] estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle. Piercy *et al* [-@piercy_histological_2016] did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles. To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with $P_{Max}$ estimated and fixed at 0.5 and 0.33, respectively.

# Results

## Simulation study

Parameter estimation was generally straightforward; with sample sizes $> 100$ a success rate of $\geq 95\%$ was achieved in all scenarios except the 2PLF-maternity method for gummy shark (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where $P_{Max}$ was always fixed at 1. In addition to convergence failures, 7 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.

Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased. The nature of this problem can be seen in the positive correlation between the $\hat{L'_{50}}$ and $\hat{P_{Max}}$ parameters whereby $P_{Max}$ affects the magnitude and direction of bias in these parameters (Figure 2). With lower values of $P_{Max}$ it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying $P_{Max}$ decreased there was a greater potential for positive bias in $\hat{P_{Max}}$ (Figure 2).

```{r fig.width = 7, fig.height= 5.5, fig.cap= "Bias (per cent relative error) in parameter estimates for $\\hat{L'_{50}}$ and $\\hat{P_{Max}}$ for 3PLF methods. Each point represents parameter estimates from one iteration of simulated data. Bias tended to increase as the true underlying $P_{Max}$ decreased."}
data <- read_rds(here("data", "simulations-pivot.Rds"))

ggplot() + 
  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
  scale_color_viridis_d(option = "D", direction = 1) + 
  facet_grid(species~method_long) + 
  ylim(-30, 50) +
  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
```

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). Fixing the asymptote was effective at reducing bias at sample sizes $\leq$ 250, although above this neither method was clearly preferable. The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). For $L'_{50}$ gains in precision tended to diminish at sample sizes $\geq$ 500. $P_{Max}$ was comparatively more challenging to estimate accurately, with precision continuing to improve noticeably at the largest sample sizes (Figure 3).

```{r fig.width = 7, fig.height= 5.5, fig.cap = "Precision (per cent absolute error) in parameter estimates  of $\\hat{P_{Max}}$ for 3PLF methods. Large sample sizes were needed to accurately estimate $\\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased "}
data <- read_rds(here("data", "simulations-summarised.Rds"))

data %>% 
  filter(par_name %in% "c") %>% 
  filter(method_long %in% c("3PLF - free", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "Mean % absolute relative error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for $L'_{50}$ was well below the expected level of 50%, irrespective of sample size. In contrast, for the 3PLF-free method interval coverage oscillated around 50% for both $L'_{50}$ and $P_{Max}$.

```{r fig.width = 7, fig.height= 5.5, fig.cap = "Interval coverage for $\\hat{L'_{50}}$ for 3PLF methods. Figure shows the percentage of simulations where the true parameter value fell within the 50% bootstrap confidence interval. "}
data %>% 
  filter(par_name %in% "m50", method_long %in% c("3PLF - free", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "% interval coverage") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```

Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in low numbers of females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in high selectivity scenarios for gummy sharks \~10% of individuals were immature (Figure S9) compared to \~50% in the corresponding scenarios for school sharks. As $P_{Max}$ decreased the proportion of both maternal females and immature females in the analysis also decreased. The latter effect was due to the right shift in length structure associated with the proportionally fewer juveniles in the population.

Overall, the 2PLF-maternity method performed the worst. Ignoring $P_{Max}$ was equivalent to introducing a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3). This manifested in an overestimation of $L'_{50}$ that was exacerbated by gear selectivity effects (Figure S2). In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance. $L_{50}$ could usually be estimated with a higher precision than $L'_{50}$. For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2% and exceeding the performance of the 3PLF-methods at sample sizes \< 250.

The effect of the different methods in ultimately calculating $R_0$ varied considerably and was strongly influenced by life history and selectivity characteristics. Greater precision was achieved with the 3PLF methods, with the 3PLF-free method generally performing best (Figure 5, Figure S11). For gummy sharks the 3PLF-fixed method was the preferred model at lower sample sizes and lower values of $P_{Max}$, consistent with decreasing performance of 3PLF methods as $P_{Max}$ decreased (Figure 2, Figure 3).

```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in accurately calculating $R_0$. The best performing method was that which minimised mean absolute error across 250 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs_error) %>%
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```

Calculations of $R_0$ for the 3PLF methods were generally unbiased with sample sizes $\geq 250$, and neither method clearly outperformed the other (Figure S10, Figure 6). As anticipated, the 2PLF-maternity method performed the worst overall and ignoring $P_{Max}$ consistently led to $R_0$ being overestimated. This effect was particularly pronounced for gummy sharks under low selectivity scenarios where derived $R_0$ was more than double the true value. Using maturity data to approximate maternal data (2PLF-maturity method) also consistently overestimated $R_0$, as expected since $L_{50}$ was less than $L'_{50}$. However, the magnitude of bias was relatively small in some cases. For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, this method frequently outperformed the 3PLF methods (Figure 6).

```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in minimising bias in calculations of $R_0$. The preferred method was that which minimised bias, $|\\text{relative error}|$ across 250 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs(rel_error)) %>% 
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```

## Empirical case study

Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$Stage > 2))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$Stage > 2))))`% were in maternal condition. Using the 3PLF-free method the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2). Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence. Comparison of 3PLF-fixed models with values of $P_{Max} = 0.33$ and $P_{Max} = 0.5$ showed much greater support for a biennial cycle ($\Delta_{AIC}$ = 9.4). The model with $P_{Max} = 0.5$ also outperformed the 3PLF-free method ($\Delta_{AIC}$ = 1.66). Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data [@burnham_model_2002].

```{r fig.width = 4.5, fig.height= 5, fig.cap= "Comparison of 3PLF-free and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C. plumbeus*, in the Gulf of Mexico and Western North Atlantic. Solid line is the expected proportion in maternal condition at length, $\\Psi'(L)$. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. $P_{Max}$ was fixed at 0.5 in the lower panel"}
# Load data - Sandbar
data <- read_rds(here("data", "empirical-plumbeus.Rds"))
preds <- read_rds(here("data", "empirical-predictions.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 220, 7.5)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - free") %>% 
  mutate(label2 = "3PLF - free")

p1 <- ggplot()  + 
  geom_ribbon(data = preds1, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds1, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2) + 
  theme(strip.text = element_text(hjust = 0.5))

preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "3PLF - fixed")

# Plot
p2 <- ggplot()  + 
  geom_ribbon(data = preds2, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds2, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2) + 
  theme(strip.text = element_text(hjust = 0.5))

grid.arrange(p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 12, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 12, font = 8)))
```

# Discussion

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, $P_{Max}$. Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that precision, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated. Using a fixed value for $P_{Max}$ resulted in lower bias at low sample sizes. This study also demonstrated that recruitment was overestimated when maturity data were used to approximate maternal data and if $P_{Max}$ was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

## Implementing maternity functions

Using simulated data to compare the relative performance of the 3PLF-free and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$. In most scenarios sample sizes of at least 250 were needed for the 3PLF-free method to approach or exceed the performance of the 3PLF-fixed method. At sample sizes below 250 it is therefore advisable to fix $P_{Max}$. Due to the decline in performance associated with longer reproductive cycles, sample sizes of at least 500 are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle.

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of $\geq1000$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999]. In the case of the gummy shark this imprecision led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$. Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a sample size of \> 1000. A similar sample size was used by Colonello *et al* [-@colonello_reproductive_2016] to successfully estimate $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].

In light of the data requirements suggested by this study, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close $L_{50}$ is to $L'_{50}$. While several studies have shown these parameters to be similar [@walker_spatial_2007; @soto-lopez_reproductive_2018], $L'_{50}$ is more frequently shifted to the right of $L_{50}$ [@braccini_determining_2006; @palacios-hernandez_temporal_2020; @montealegre-quijano_sexual_2014; @colonello_reproductive_2016], and this length difference can equate to one or more years. For example, Harry *et al* [-@harry_age_2013] found both spot-tail shark, *C. sorrah*, and Australian blacktip shark, *C. tilstoni*, began reproducing the year after reaching sexual maturity. Similarly, Fujiyama found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.

More research is needed to understand the relationship between size at maturity and maternity and variability within and among taxa. In general, this study confirms Walker's [-@walker_reproduction_2005] assertion that the common practice of weighting the maturity curve by the frequency of parturition overestimates recruitment. This study also confirms that fitting a 2PLF to maternal data when $P_{Max} < 1$ (effectively ignoring $P_{Max}$) is likely to result in biased parameters and can also overestimate recruitment. While not widely used, this approach has been undertaken in several studies [@mejia-falla_reproductive_2012; @baremore_reproduction_2012; @baremore_reproduction_2013; @rambahiniarison_life_2018].

The success of any analysis using the 3PLF-free method, and its ability to outperform the 3PLF-fixed method, ultimately depends on the analyst choosing a suitable value for $P_{Max}$. In this study, the value for $P_{Max}$ used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data. The potential to introduce a greater level of bias in the analysis through the incorrect selection of $P_{Max}$ should also be considered if using this approach.

## Empirical study

The empirical analysis carried out using data for the sandbar shark illustrates the potential for the 3PLF-free method to provide novel insights from maternal data. Detailed reproductive studies of the western North Atlantic population by Baremore and Hale [-@baremore_reproduction_2012] and Piercy *et al* [-@piercy_histological_2016] were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or longer at the population level. Reanalysis of these combined data sets using the 3PLF-free method, however, resulted in a maximum likelihood estimate of $P_{Max} = 0.48$, most consistent with biennial cycle. Using AIC values to compare the relative plausibility of 3PLF-fixed models corresponding to biennial and triennial cycles also strongly supported a biennial cycle.

This counter intuitive result appears to be due to the gradual attainment of maternal status as a function of length in sandbar sharks. $L'_{5}$ and $L'_{95}$ occur at 146cm and 176cm, respectively, a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013]. It therefore appears that the shape of the maternity curve is not just due to individuals differing in their size at maturity, but can also be attributed to smaller females reproducing less frequently. Larger females, in contrast, appear capable of reproducing biennially. This finding has important management implications that would not otherwise be apparent if using a maturity curve to approximate maternity data. Not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently as well. This implies removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

## Advantages and future directions

Despite the simplicity of the 3PLF-free method, it nonetheless represents a conceptual shift for chondrichthyan reproductive and population biology. Until now reproductive periodicity has largely been determined qualitatively from observations of female reproductive biology or inferred based on circumstantial evidence. Using the 3PLF-free method to estimate $P_{Max}$ within a maternity function transforms it from a nuisance parameter to one of direct inferential interest that can be seen as a valuable output of the modelling process itself. In addition to the potential benefits of using this method to improve estimation of maternal parameters, there are also a range of other advantages that arise from estimating $P_{Max}$.

One major benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to $P_{Max}$. Such assumptions can have important implications for population modelling. For example, to account for uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, *Urolophus paucimaculatus*, Trinnie *et al* (2014) estimated maternity parameters for ten plausible scenarios. In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle.

More commonly, practitioners have chosen fixed values of the form $P_{Max} = 1/n$, where *n* is the apparent duration of the reproductive cycle in years. This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently that the larger population. Long-term study of offspring from genetically profiled lemon sharks, *Negaprion brevirostris*, confirmed that most individuals reproduced biennially, but also revealed some cases of triennial reproduction [@feldheim_two_2014]. Statistical estimation of $P_{Max}$ from data avoids these decisions and enables uncertainty in this parameter to be included in subsequent population models.

Estimation of $P_{Max}$ using the 3PLF-free method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans. White sharks and whale shark are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016]. In both species lack of data on reproductive frequency is a major impediment to population modelling and the development of management strategies [@huveneers_future_2018; @bowlby_implications_2020]. With many of these populations now the focus of ongoing monitoring, non-lethal methods for assessing maternal state may eventually provide a way of collecting sufficient data for estimation of $P_{Max}$ [@sulikowski_seasonal_2016].

A logical progression from estimating $P_{Max}$ is the consideration of alternative functional forms for this parameter such as time- or space-varying $P_{Max}$ or more complex length- or age-dependent forms. Despite the diverse range of reproductive modes found in chondrichthyan fishes, the current, limited knowledge of maternal investment has typically investigated variables such as litter size and pup condition [@hussey_maternal_2010]. Little is known about how reproductive frequency varies as a function of size or age, in response to environmental effects, or at changing population densities. Better understanding of these factors has the potential to improve understanding reproductive strategies and is, many cases, of direct relevance to management.

# Acknowledgements

John Carlson

# References

### 2023-11-14 rmds/notes.Rmd
refresh

---
title: "Notes for maternal output paper"
output: html_notebook
csl: "../references/cjfas.csl"
bibliography: "../references/maternal-output.bib"
---

# Reading list
Walker 2005
Monteleagre-Quijano (2014)
Blue sharks (2 papers)

### Huveneers et al [-@huveneers_reproductive_2007]
- 3 species of Wobbegong, ovarian follicle development supported a three year reproductive cycle. Females found concurrently sat three different stages of development.

### Braccini et al [-@braccini_determining_2006]
- Piked spurdog
- Used both maternity ogives and pregnancy ogives
- Pregnancy ogives had $P_{Max}$ of 1 because there was around a two year gestation period
- Had to make some assumptions about follicle growth rates because of asynchronous reproduction, but interesting paper and good data

### Baremore and Passerotti [-@baremore_reproduction_2013]
- Limbatus
- Only pregnant females included...may need to check this
- Ogive multiplied by 0.5

### Baremore and Passerotti [-@baremore_reproduction_2012]
- Plumbeus
- Need to check with Ivy about the criteria used to determine maternal status, but seems to be pregnant females
- Possible 2 or 3 year reproductive cycle

### Taylor et al [-@taylor_living_2016]
- R. taylori
- Some individuals maturing at a large length caused the maternity ogive to be right-shifted for this species, highlighting importance even for annually reproducing species

### Harry et al [-@harry_age_2013]
- C. sorrah and C. tilstoni
- In C. sorrah maturity occurs at two but females don't begin reproducing to the following season / year. This is captured in the maternity ogive and is important given the species is relatively short-lived and fast maturing
- Some evidence of non-annual reproduction incorporated into the maternity ogive for C. tilstoni

## Walker [-@walker_spatial_2007]
- Gummy shark

## Mejia-Falla et al 2012

### References

### 2023-11-14 rmds/tables-supp.Rmd
refresh

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load results
source(here::here("code", "setup.R"))

data <- read_rds(here("data", "simulations-summarised.Rds"))
```

## Tables

Table S1.

```{r}
# Tidy up data, save, reorder manually, then reload
data %>% 
select(species, repro_freq, Nsamples, mesh_name_long, everything()) %>% 
  select(-sd_re, - sd_ae, -int_coverage, -mesh_name, -ntotal, -prop_pregnant, -n_maternal, -n_immature) %>% 
  filter(!par_name %in% "m95") %>% 
  pivot_wider(names_from = c("method_long", "par_name"), values_from = c("rel_error", "abs_error", "conv_success")) %>% 
  arrange(mesh_name_long, species, desc(repro_freq)) %>% 
  write_csv(here("tables", "table-s1.csv"))
```

Table S2. 
```{r}
data %>% 
  group_by(species, repro_freq, mesh_name, Nsamples, ntotal) %>% 
  nest() %>% 
  select(-data) %>% 
  ungroup() %>% 
  mutate(n = 250 - ntotal) %>% 
  filter(n > 0) %>% 
  write_csv(here("tables", "table-s2.csv"))

```

### 2023-11-14 rmds/tables.Rmd
refresh

---
output: 
  pdf_document:
    latex_engine: "pdflatex"
geometry: 
    - bottom=1.5cm
    - top=1.5cm
    - left=1.5cm
    - right=1.5cm
header-includes:
    - \fontsize{8}{20}
    - \usepackage{lscape}
    - \pagenumbering{gobble} 
tables: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("code", "setup.R"))

# Load data
empirical <- read_rds(here("data", "empirical-summary.Rds"))
```

\
\
```{r tbl1, results = 'asis'}
t1 <- readxl::read_excel(here("data", "table-1.xlsx")) %>% 
  mutate(`School shark` = as.character(`School shark`)) %>% 
  mutate(`Gummy shark` = as.character(`Gummy shark`)) %>%  
  xtable(caption = "Descriptions and values of parameters used in a simulation study to evaluate the performance of alternative methods for estimating maternity parameters \\newline")

print(t1, type = "latex", caption.placement = "top", comment = F, include.rownames = F, 
      sanitize.text.function = identity, booktabs = TRUE)
```

\newpage

```{r tbl2, results = 'asis'}
t2 <- empirical %>% 
  rename(`$\\hat{P_{\\text{Max}}}$` = PMAX) %>% 
  rename(`$\\hat{L'_{50}}\\text{(cm)}$` = L50) %>% 
  rename(`$\\hat{L'_{95}}\\text{(cm)}$` = L95) %>% 
  rename(`$\\Delta_{i}$` = delta_AIC) %>% 
  rename(Method = method) %>%
  rename(`$w_i$` = w_AIC)

t2 <- xtable(t2, caption = "Parameter estimates and model selection criteria for three parameter logistic functions fit to empirical data for western North Atlantic sandbar sharks")

print(t2, type = "latex", caption.placement = "top", comment = F, include.rownames = F,
      sanitize.text.function = identity, booktabs = TRUE)
```

### 2023-11-14 code/empirical.R

@@ -1,0 +1,297 @@
+# Housekeeping ----------------------------------------------------------------
+
+source(here::here("code", "setup.R"))
+source(here::here("code", "functions.R"))
+
+# Sandbar shark analysis ------------------------------------------------------
+
+compile("code/logistic3.cpp")
+dyn.load(dynlib("code/logistic3"))
+
+# Load data
+# Complete dataset from Ivy, includes some smaller females that were missing from previous spreadsheet
+data_complete <- readxl::read_xlsx(here("data", "plumbeus-unprocessed.xlsx"))
+
+# Incomplete dataset that has full repro details
+data_incomplete <- readxl::read_xls(here("data", "plumbeus-extra-unprocessed.xls"), sheet = 6) %>%
+  select(
+    ID = `SBAR .`, `OVARY L (mm)`, `OVARY WIDTH`, `OVARY WEIGHT (g)`, `TOTAL OOCYTES`, `. VIT`,
+    `DIAMETER LARGEST OO`, `UTERUS WIDTH`, `NIDAMENTAL WIDTH`, `PUPS?`, `. LEFT`, `. RIGHT`
+  ) %>%
+  arrange(ID) %>%
+  # Remove #358 as it is a duplicate, doesn't seem to have been included in final analysis
+  filter(!row_number() %in% 281)
+
+# Load data for pups
+pups <- read_csv(here("data", "plumbeus-pups-unprocessed.csv")) %>% 
+  rename_all(tolower) %>%
+  select(ID = maternal.id, puplen = `stl.cm`) %>% 
+  mutate(puplen = as.numeric(puplen))
+
+# Treat July as beginning of a new year / season
+
+# SBAR1242 stage 4 just ovulated and pregnant, but June so count as resting for this year
+# SBAR1248 stage 4 just ovulated and pregnant, but June so count as resting for this year
+# SBAR1244 stage 4 just ovulated and pregnant, but June so count as resting for this year
+# SBAR471 non-pregnant, perhaps pupped late?
+# SBAR353 stage 6 postpartum but month is September so would have been maternal last season
+# SBAR303 possible stage 6 but definitely not in maternal condition
[…]

### 2023-11-14 code/functions.R

@@ -1,0 +1,635 @@
+# Utility functions -----------------------------------------------------------
+
+## Parameter transformation
+inv_logit <- function(x) {
+  1 / (1 + exp(-x))
+}
+logit <- function(x) {
+  -log((1 - x) / x)
+}
+
+## Save data
+save_sim_data <- function(df, id) {
+  dir_out <- paste0(here(), "/data/simulations/", id)
+  if (!dir.exists(dir_out)) {
+    dir.create(dir_out)
+  }
+  write_rds(df, paste0(dir_out, "/population-structure.rds"))
+}
+
+## The continuous time Lotka equation (Xiao and Walker 2000)
+continuous_lotka <- function(a, Lambda, df) {
+
+  # Length at age function
+  len <- function(a) {
+    df$Linf * (1 - exp(-df$K * (a - df$t0)))
+  }
+
+  # Maternity at age function
+  maternity <- function(a) {
+    (df$c / (1 + exp(-log(19) * ((len(a) - df$m50) / (df$m95 - df$m50)))))
+  }
+
+  # Fecundity at age function
+  fecundity <- function(a) {
+    if (df$species == "School shark") {
+      df$f + df$g * len(a)
+    } else {
+      df$f * exp(df$g * len(a))
+    }
+  }
+
+  # Birth function
+  birth <- function(a) {
+    if (df$species == "School shark") {
+      ifelse(a >= 11.2404, fecundity(a) * maternity(a) * (1 / 2), 0)
+    } else {
+      ifelse(a >= 3.9694, fecundity(a) * maternity(a) * (1 / 2), 0)
+    }
+  }
+
+  # Intrinsic rate of population decrease with age
+  LAMBDA <- function(a) {
+    rep(Lambda, length(a))
+  }
+
+  # Sum to optimize
+  birth(a) * exp(sapply(a, function(x) {
+    integrate(LAMBDA, 0, x)$value
+  }))
+}
+
+## Function for solving Lotka equation
+solve_lotka <- function(df, Lambda = 0) {
+  Integral <- integrate(continuous_lotka, 0, Inf, Lambda = Lambda, df = df)$value
+  dif <- abs(1 - Integral)
+  return(dif)
+}
+
+## Function for optimizing Lotka equation
+optimize_lotka <- function(df, lower = -2, upper = 0) {
[…]

### 2023-11-14 code/logistic2.cpp

@@ -1,0 +1,31 @@
+#include <TMB.hpp>
+
+/* Parameter transform */
+template <class Type>
+Type inv_logit(Type x){return 1/(1 + exp(- x));}
+
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(y);
+  DATA_VECTOR(x);
+  
+  PARAMETER(l50);
+  PARAMETER(l95);
+  
+  vector<Type> X = log(19) * ((x - l50) / (l95 - l50));
+  vector<Type> p = inv_logit(X);
+  
+  
+  Type nll=-sum(dbinom(y,Type(1.0),p,true));
+  
+  SIMULATE {
+    y = rbinom(Type(1.0), p);
+    REPORT(y);
+  }
+
+  
+  return nll;
+}
+

### 2023-11-14 code/logistic2maternal.cpp

@@ -1,0 +1,31 @@
+#include <TMB.hpp>
+
+/* Parameter transform */
+template <class Type>
+Type inv_logit(Type x){return 1/(1 + exp(- x));}
+
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(z);
+  DATA_VECTOR(x);
+  
+  PARAMETER(m50);
+  PARAMETER(m95);
+  
+  vector<Type> X = log(19) * ((x - m50) / (m95 - m50));
+  vector<Type> p = inv_logit(X);
+  
+  
+  Type nll=-sum(dbinom(z,Type(1.0),p,true));
+  
+  SIMULATE {
+    z = rbinom(Type(1.0), p);
+    REPORT(z);
+  }
+  
+  
+  return nll;
+}
+

### 2023-11-14 code/logistic3.cpp

@@ -1,0 +1,31 @@
+#include <TMB.hpp>
+
+/* Parameter transform */
+template <class Type>
+Type inv_logit(Type x){return 1/(1 + exp(- x));}
+
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(z);
+  DATA_VECTOR(x);
+
+  PARAMETER(m50);
+  PARAMETER(m95);
+  PARAMETER(c);
+  
+  vector<Type> X = log(19) * ((x - m50) / (m95 - m50));
+  vector<Type> p = c * inv_logit(X);
+    
+    
+  Type nll=-sum(dbinom(z,Type(1.0),p,true));
+  
+  SIMULATE {
+    z = rbinom(Type(1.0), p);
+    REPORT(z);
+  }
+  
+  return nll;
+}
+

### 2023-11-14 code/selectivity.R

@@ -1,0 +1,32 @@
+
+
+sel_pars_gummy <- list(mesh = seq(4, 9, 0.01),
+                 
+                 theta1 = 184.3,
+                 theta2 = 29739,
+                 m50 = 1129,
+                 m95 = 1344)
+
+sel_pars_school <- list(mesh = seq(4, 9, 0.01),
+                       
+                        theta1 = 192,
+                        theta2 = 67595,
+                        m50 = 1421,
+                        m95 = 1488)
+
+peak_sel <- function(len, df){
+  1 - (len / (df$rho * df$eta))^df$rho * exp(df$rho - len / df$eta)
+}
+
+sel_df <- cross_df(sel_pars_gummy) %>% 
+  mutate(eta = -0.5 * (theta1 * mesh - sqrt(theta1^2 * mesh^2 + 4 * theta2))) %>% 
+  mutate(rho = (theta1 * mesh) / eta) %>% 
+  mutate(id = row_number()) %>%
+  group_by(id) %>% nest(.key = "df")  %>% 
+  mutate(peak_sel = map_dbl(df, ~ optimize(peak_sel, lower = 0, upper = 2500, df = .)$minimum)) %>% 
+  unnest() %>% 
+  mutate(peak_sel = round(peak_sel, 0)) %>% 
+  mutate(P = (1 / (1 + exp(-log(19) * ((peak_sel - m50) / (m95 - m50)))))) %>% 
+  mutate(P = round(P, 2))
+
+# Selectivity parameters (Kirkwood and Walker 1986)

### 2023-11-14 code/setup.R

@@ -1,0 +1,18 @@
+# Setup
+
+# Knitr options
+knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = FALSE, dev = "pdf")
+
+# Housekeeping
+set.seed(3)
+
+# Load libraries
+libs<-c("MASS", "tidyverse", "knitcitations", "bibtex", "lubridate",
+        "forcats", "here", "TMB", "broom", "rsample", "cowplot",
+        "grid", "gridExtra", "xtable")
+
+lapply(libs, library, character.only = TRUE)
+
+# Bibliography
+cleanbib()
+cite_options(citation_format = "pandoc", check.entries = FALSE)

### 2023-11-14 code/simulation.R

@@ -1,0 +1,176 @@
+# Housekeeping ----------------------------------------------------------------
+
+source(here::here("code", "setup.R"))
+source(here::here("code", "functions.R"))
+
+compile("code/logistic3.cpp")
+dyn.load(dynlib("code/logistic3"))
+compile("code/logistic2.cpp")
+dyn.load(dynlib("code/logistic2"))
+compile("code/logistic2maternal.cpp")
+dyn.load(dynlib("code/logistic2maternal"))
+
+# Parameter section  ----------------------------------------------------------
+
+sim_pars_galeus <- list(
+  species = "School shark",
+  
+  # Maximum age
+  AMAX = 54, 
+  # Natural mortality (Grant et al 1979)
+  M = 0.1075,
+  # Fishing mortality (Nominally set to 25 % of M)
+  F_mort = 0.25 * 0.1075,
+  # von Bertalanffy growth parameters (Grant et al 1979)
+  K = 0.1600,
+  Linf = 1618.3,
+  t0 = -1.2818,
+  cv_l = 0.05,
+  # Maternity parameters (Walker 2005)
+  m50 = 1421,
+  m95 = 1488,
+  c = c(1, 3 / 4, 2 / 3, 1 / 2, 1 / 3, 1 / 4), # Proportion maternal
+  # Maturity parameters (Walker 2005)
+  l50 = 1349,
+  l95 = 1502,
+  # Uterine fecundity as function of length (Walker 2005)
+  f = -46.0,
+  g = 0.0491,
+  # Selectivity parameters (Punt and Walker 1998)
+  # Correspond to peak sel at 25% and 75% length maternity
+  mesh = c(7.27, 7.53),
+  theta1 = 192,
+  theta2 = 67595,
+  
+  # Number of samples
+  Nsamples = c(50, 100, 250, 500, 1000, 2500)
+)
+
+sim_pars_mustelus <- list(
+  species = "Gummy shark",
+  
+  # Maximum age
+  AMAX = 20, 
+  # Natural mortality (Max age reported in Moulton et al 1992, Walker 1992 M)
+  M = 0.1970,
+  # Fishing mortality (Nominally set to 25 % of M)
+  F_mort = 0.25 * 0.1970,
+  # von Bertalanffy growth parameters (Moulton et al 1992, BS 1973-1976)
+  K = 0.123,
+  Linf = 2019,
+  t0 = -1.55,
+  cv_l = 0.05,
+  # Maternity parameters (Walker 2007 EKI 73-76)
[…]

### 2023-11-14 code/summarise_results.R

@@ -1,0 +1,97 @@
+# Housekeeping ----------------------------------------------------------------
+
+source(here::here("code", "setup.R"))
+source(here::here("code", "functions.R"))
+
+# Summarise simulation study --------------------------------------------------
+data <- read_rds(here("data", "simulations.Rds")) %>% 
+  group_by(id) %>% nest() %>% 
+  # How many iterations in each scenario had zero pregnant females and were excluded? 
+  mutate(zero_maternal = map_dbl(data, ~ sum(.$no_maternal))) %>% 
+  unnest(cols = c(data)) %>% 
+  ungroup() %>% 
+  filter(no_maternal == 0) %>% select(-no_maternal) %>% 
+  # How many other iterations in each scenario failed due to lack of pregnant females? 
+  mutate(few_maternal = ifelse(is.na(method), 1, 0)) %>% 
+  group_by(id) %>% 
+  nest() %>% 
+  mutate(low_maternal = map_dbl(data, ~ sum(.$few_maternal))) %>% 
+  unnest(cols = c(data)) %>% 
+  ungroup() %>% 
+  filter(!is.na(method)) %>% 
+  # Calculate total number of iterations that failed to start due to lack of data
+  mutate(maternal_data_fail = low_maternal + zero_maternal) %>% 
+  # Calculate total number of iterations done for each simulation
+  mutate(ntotal = 250 - maternal_data_fail) %>% 
+  select(-low_maternal, -few_maternal, -zero_maternal) %>% 
+  # Now tally parameter boundary failures and non-convergence failures
+  group_by(id, iteration, method) %>%
+  nest() %>% 
+  # Remove iterations where boundary was hit
+  mutate(hit_boundary = map_lgl(data, par_check)) %>%
+  filter(hit_boundary %in% FALSE) %>% 
+  select(-hit_boundary) %>% 
+  unnest(cols = c(data)) %>% 
+  # Remove iterations where convergence failed
+  filter(convergence == 0) %>% 
+  mutate(method_long = case_when(
+    method %in% "2" ~ "2PLF - maturity",
+    method %in% "2a" ~ "2PLF - maternity",
+    method %in% "3" ~ "3PLF - free",
+    method %in% "3a" ~ "3PLF - fixed"
+  )) %>%
[…]

### 2023-11-14 run_all.R

@@ -1,0 +1,37 @@
+## Run all
+
+# Load libraries
+library(knitr)
+library(rmarkdown)
+
+# Data analysis
+source(here::here("code", "simulation.R")) # 48 hour run time on my computer
+source(here::here("code", "summarise-results.R"))
+source(here::here("code", "empirical.R"))
+
+
+# Generate files
+render("rmds/manuscript.Rmd", output_format = "all", output_dir = "reports", envir = new.env())
+render("rmds/figures.RmD", output_format = "all", output_dir = "reports", envir = new.env())
+render("rmds/figures-supp.RmD", output_format = "all", output_dir = "reports", envir = new.env())
+render("rmds/tables.RmD", output_format = "all", output_dir = "reports", envir = new.env())
+
+## Zotero citation key format
+# [auth:lower]_[Title:lower:skipwords:select,1,1]_[year]
+
+# Rmarkdown template
+# http://labrtorian.com/2019/08/26/rmarkdown-template-that-manages-academic-affiliations/
+
+# TODO 
+# - Bubley spiny dogfish - inconsistent definition (any with pups or large developing follicles)
+# - Cotton Squalus mitsikuri uses pregnancy ogive as maternity ogive
+# - Fujimani blue shark
+# - Rochowski paper
+# - Corro-Espinosa rhizoprionodon
+# - mustelus henlei (file:///Users/avh/Downloads/2805-Article%20Text-420420528-1-10-20180628.pdf)
+# - Squalus blainvillei (https://www-publish-csiro-au.libproxy.murdoch.edu.au/mf/pdf/MF19372)
+
+# Notes
+# - https://www.jstor.org/stable/2531207
+# - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5978598/
+# - https://www.researchgate.net/profile/Gunter_Maris/publication/237079179_On_Interpreting_the_Model_Parameters_for_the_Three_Parameter_Logistic_Model/links/5661554508aebae678aa7970.pdf

### 2023-11-16 code/simulation.R
simulations

@@ -105,7 +105,7 @@ data <-
   ungroup() %>% 
   rename(pars = data) %>%
   # Create and number n repliactes of each row
-  mutate(reps = 6) %>% 
+  mutate(reps = 300) %>% 
   uncount(reps) %>%
   group_by(id) %>%
   mutate(iteration = row_number()) %>%
@@ -118,6 +118,7 @@ data <- read_rds("data/generated_data.rds")
 
 
 data <- data %>%
+  slice(1:10800) %>% 
   # Simulate maternal data
   mutate(data = map2(pars, data, generate_maternal_data)) %>% # How many maternal females are there in the simulated data
   mutate(n_maternal = map_dbl(data, ~ sum(.$z))) %>%
@@ -162,15 +163,16 @@ data <- data %>%
     mutate(int_coverage = ifelse(par_true >= lower & par_true <= upper, TRUE, FALSE)) %>%
     select(-xmin, -xmax, -data)
 
+write_rds(data, here("data", "simulations-gummy-1.Rds"))
 
 write_rds(data, here("data", "simulations-revised-all.Rds"))
 
-data <- read_rds("data/simulations-revised-gummy.rds")
+data <- read_rds("data/simulations-gummy.rds")
 data1 <- read_rds("data/simulations-revised-school-1.rds")
 data2 <- read_rds("data/simulations-revised-school-2.rds")
 data3 <- read_rds("data/simulations-revised-school-3.rds")
 
 data <- rbind(data, data1) %>% rbind(data2) %>% rbind(data3)
 
-write_rds(data, here("data", "simulations.Rds"))
+
 

### 2024-01-22 rmds/figures-supp.Rmd
updates

@@
 ```{r fig.width = 12, fig.height=7}
-ggplot(data %>% filter(par_name %in% "m50", method_long %in% c("3PLF - free", "3PLF - fixed"))) +
+ggplot(data %>% filter(par_name %in% "m50", method_long %in% c("3PLF - estimated", "3PLF - fixed"))) +
   geom_hline(yintercept = 50, linetype = "dashed") +
@@
 ```{r fig.width = 12, fig.height=5}
-ggplot(data %>% filter(par_name %in% "c", method_long %in% c("3PLF - free"))) +
+ggplot(data %>% filter(par_name %in% "c", method_long %in% c("3PLF - estimated"))) +
   geom_hline(yintercept = 50, linetype = "dashed") +
@@
 ```{r, fig.width = 11, fig.height = 8.5}
-filter(data, method_long %in% c("3PLF - free", par_name %in% "c")) %>% 
+filter(data, method_long %in% c("3PLF - estimated", par_name %in% "c")) %>% 
   ggplot() + 
@@
 ```{r, fig.width = 11, fig.height = 8.5}
-filter(data, method_long %in% c("3PLF - free", par_name %in% "c")) %>% 
+filter(data, method_long %in% c("3PLF - estimated", par_name %in% "c")) %>% 
 ggplot() + 

### 2024-01-22 rmds/manuscript.Rmd
updates

@@
 - name: Ivy Baremore
-  affiliation: MarAlliance
+  affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
   number: 2
@@
 Applying logistic models to a combination of simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-At sample sizes < 250 individuals a fixed-value for $P_{Max}$ was more effective at reducing bias.
+At sample sizes < 250 individuals a fixed-value for $P_{Max}$ was more effective at reducing bias in simulated data for gummy sharks, *Mustelus antarcticus*.
 Maturity parameters could be estimated with greater precision however substituting them in place of maternity parameters overestimated lifetime reproductive output.
@@
 
-In chondrichthyan fishes (sharks, rays, skates, and chimaeras), the process of maturation can likewise be modeled with a simple 2PLF.
-However for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
-Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015].
+In chondrichthyan fishes (sharks, rays, skates, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
+Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015].
 Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017].
 This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014].
-Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stock-recruitment_2013].
+Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stockrecruitment_2013].
 This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output.
@@
 
-Numerous authors have adopted Walker's [-@walker_reproduction_2005] maternity function, including for species with biennial or longer reproductive cycles [@huveneers_reproductive_2007; @rochowski_reproductive_2015; @trinnie_reproductive_2016] as well as those with annual or shorter reproductive cycles [@trinnie_reproductive_2009; @mejia-falla_reproductive_2012; @harry_age_2013; @taylor_living_2016].
-Techniques have been described for determining maternal output in asynchronously reproducing species [@braccini_determining_2006; @colonello_reproductive_2016], and maternity functions are also increasingly being used directly in shark and ray population assessments [@sedar_sedar_2012; @sedar_sedar_2017].
+Numerous authors have adopted Walker's [-@walker_reproduction_2005] maternity function, including for species with biennial or longer reproductive cycles [@huveneers_reproductive_2007; @rochowski_reproductive_2015; @trinnie_reproductive_2016] as well as those with annual or shorter reproductive cycles [@trinnie_reproductive_2009; @mejiafalla_reproductive_2012; @harry_age_2013; @taylor_living_2016].
+Techniques have been described for determining maternal output in asynchronously reproducing species [@braccini_determining_2006; @colonello_reproductive_2016], and maternity functions are also increasingly being used directly in shark and ray population assessments [@sedar_sedar_2012; @sedar_sedar_2017; @sedar_sedar_2023].
 
-Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions, potentially due to a range of factors.
+Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions.
 Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles.
-Such data can be difficult to collect for sharks and rays, which are often data-poor and sampled opportunistically.
+Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically.
 The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.
@@
 
-To assess the feasibility of estimating maternity parameters a simulation study was done using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1).
+To assess the feasibility of estimating maternity parameters a simulation study was carried out using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1).
 Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs.
@@
 Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency.
-All simulation modelling and data analysis was conducted using the R language [@rcoreteam_r:_2018] and the Template Model Builder (TMB) R package [@kristensen_tmb:_2016].
+All simulation modelling and data analysis was conducted using the R language [@r_core_team_r:_2018] and the Template Model Builder (TMB) R package [@kristensen_tmb:_2016].
 
@@
 
-Catch at length data were simulated using a female-only, length-based catch curve model [@L3Assess].
+Catch at length data were simulated using a female-only, length- and age-based equilibrium population model [@hesp_l3assess_2023].
 The approach involved calculating expected survival and catches per recruit subject to specified life history parameters, gear selectivity, and total mortality.
@@
 Recruitment into the population was assumed to occur at age 0 with length conforming to a normal distribution.
-The proportion of fish in length class, *l*, at age, $a = 0$, was calculated as $$\theta_k = \int_{l^-_k}^{l^+_k}f_{a=0}(l)dl$$ where $l^-$ and $l^+$ are the lower and upper limits of each 1cm length class, *k*.
+The proportion of fish of length, *l*, at age, $a = 0$, was calculated as $$\Theta_k = \int_{l^-_k}^{l^+_k}f_{a=0}(l)dl$$ where $l^-$ and $l^+$ are the lower and upper limits of each 1cm length class, *k*.
 $f_{a=0}(l)$ is the value of the normal probability density function at age $a = 0$ for a given length, *l*,$$f_{a=0}(l) = \frac{1}{\sigma\sqrt{2{\pi}}}\exp\bigg[\frac{(l-\mu)^2}{2\sigma^2}\bigg]
@@
 
-Growth in subsequent age classes was modelled using a length transition matrix [@punt1997; @hall2000], $G = g_{j,k}$ that represents the probability that a fish in length class, $j$, will grow into length class, $k$, over a specified time interval $$ G = \begin{bmatrix}
+Growth in subsequent age classes was modelled using a length transition matrix [@punt_estimating_1997; @hall_modelling_2000], $G = g_{j,k}$ that represents the probability that a fish in length class, $j$, will grow into length class, $k$, over a specified time interval $$ G = \begin{bmatrix}
  g_{1,1}&0& 0 &0\\ 
@@
 
-where the elements of the matrix followed the general form of Punt *et al* [-@punt1997] and are given by$$ g_{k,j} = \left\{
+where the elements of the matrix followed the general form of Punt *et al* [-@punt_estimating_1997] and are given by$$ g_{k,j} = \left\{
     \begin{array}{ll}
@@
 Two selectivity scenarios were simulated corresponding to low and high selectivity of the maternal component of the population.
-For consistency between species, gillnet mesh sizes were chosen that would result in peak relative selectivity occurring at the lengths equal to 5% and 95% of the population in maternal condition (Table 1).
+For consistency between species, gillnet mesh sizes were chosen that would result in peak relative selectivity occurring at the lengths equal to 25% and 75% of the population in maternal condition (Table 1).
 
@@
 
-$$ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=1<a<A \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A \\    \end{array}\right.$$
+$$ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=1<a<A_{Max} \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A_{Max} \\    \end{array}\right.$$
 
-The estimated numbers of fish caught in length class, l, at age, a, was calculated using the Baranov catch equation as
+where $A_{Max}$ is the maximum age (Table 1). The estimated numbers of fish caught in length class, l, at age, a, was calculated using the Baranov catch equation as
 
@@
 
-Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This processes was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l(a) - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 250 iterations were run for each of the 144 unique variables combinations.
+Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This processe was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.
 
@@
 
-Two approaches to estimating maternity parameters were compared, the 3PLF itself (3PLF-free) and the 3PLF function with a fixed asymptote (3PLF-fixed).
+Two approaches to estimating maternity parameters were compared: the 3PLF itself (3PLF-estimated) and the 3PLF function with a fixed asymptote (3PLF-fixed).
 Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF maturity).
-The 3PLF-free method was examined to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016].
+The 3PLF-estimated method was undertaken to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016].
 The 3PLF-fixed is the method described by Walker [-@walker_reproduction_2005] and most commonly used in practice.
@@
 $P_{Max}$ fixed at 1).
-The 2PLF-maturity method is tacitly assumed if using mature females to approximate maternal ones.
-The resulting maturity curve can then by weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially).
+The 2PLF-maturity method is commonly used to approximate a maternity curve, and tacitly assumed to be similar.
+The resulting maturity curve can then be weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially [@sedar_sedar_2023]).
 
-Only the 3PLF-free method involves statistically estimating $P_{Max}$, which must be subjectively chosen in the case of the 3PLF-fixed and the 2PLF-maturity.
+Only the 3PLF-estimated method involves statistically estimating $P_{Max}$, which must be subjectively chosen in the case of the 3PLF-fixed and the 2PLF-maturity.
 $P_{Max}$ is ideally chosen based on detailed study of the ovarian and uterine cycles [@walker_reproduction_2005], or alternatively based on the proportion of mature females observed to be in maternal condition during sampling [@baremore_reproduction_2012; @harry_age_2013; @trinnie_reproductive_2016].
@@
 $P_{Max}$ was chosen as the proportion of females in maternal condition above a the length at which $99\%$ of females were mature, $L > \Psi(L_{99\%})$.
-If this failed (e.g.
-zero pregnant females), $L_{95}$ was used, followed by $L_{50}$, and finally the proportion of all mature females in the sample that were in maternal condition.
+In some cases this procedure failed due to there being no maternal females in the simulated data set meeting this criteria.
+If this occurred, $L_{95}$ was used, followed by $L_{50}$, and finally the proportion of all mature females in the sample that were in maternal condition.
 Statistical estimation of all non-fixed parameters in each of the methods was undertaken using maximum likelihood.
@@
 Data from two studies of the reproductive biology of the sandbar shark, *Carcharhinus plumbeus*, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods.
-The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{yr}^{-1}$) that was thought to have a biennial reproductive cycle [@springer_natural_1960].
+The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{yr}^{-1}$) with a protracted reproductive cycle lasting longer than a year [@springer_natural_1960].
 The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s [@sminkey_age_1995] and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes [@baremore_reproduction_2012; @piercy_histological_2016].
@@
 
-Parameter estimation was generally straightforward; with sample sizes $> 100$ a success rate of $\geq 95\%$ was achieved in all scenarios except the 2PLF-maternity method for gummy shark (Figure S1, Table S1).
+Parameter estimation was generally straightforward; with sample sizes $> 50$ a success rate of $\geq 95\%$ was achieved in all scenarios (Figure S1, Table S1).
 Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where $P_{Max}$ was always fixed at 1.
-In addition to convergence failures, 7 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.
+In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.
 
@@
 For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3).
-Fixing the asymptote was effective at reducing bias at sample sizes $\leq$ 250, although above this neither method was clearly preferable.
+In some scenarios, particularly with sample sizes $\leq$ 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3).
 The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5).
-For $L'_{50}$ gains in precision tended to diminish at sample sizes $\geq$ 500.
-$P_{Max}$ was comparatively more challenging to estimate accurately, with precision continuing to improve noticeably at the largest sample sizes (Figure 3).
+Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved precision with the 3PLF methods, although at larger sample sizes better precision was achieved by estimating $P_{Max}$ (Figure 3).
+Overall, $P_{Max}$ was comparatively more challenging to estimate accurately than $L'_{50}$ (Figure S4, Figure S5).
 
@@
   filter(par_name %in% "c") %>% 
-  filter(method_long %in% c("3PLF - free", "3PLF - fixed")) %>% 
+  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
   filter(mesh_name %in% "high") %>% 
@@
 For most simulations, interval coverage for $L'_{50}$ was well below the expected level of 50%, irrespective of sample size.
-In contrast, for the 3PLF-free method interval coverage oscillated around 50% for both $L'_{50}$ and $P_{Max}$.
+In contrast, for the 3PLF-estimated method interval coverage oscillated around 50% for both $L'_{50}$ and $P_{Max}$.
 
-```{r fig.width = 7, fig.height= 5.5, fig.cap = "Interval coverage for $\\hat{L'_{50}}$ for 3PLF methods.
+```{r fig.width = 7, fig.height= 5.5, fig.cap = "Interval coverage for $\\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
 Figure shows the percentage of simulations where the true parameter value fell within the 50% bootstrap confidence interval.
@@
 data %>% 
-  filter(par_name %in% "m50", method_long %in% c("3PLF - free", "3PLF - fixed")) %>% 
+  filter(par_name %in% "m50", method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
   filter(mesh_name %in% "high") %>% 
@@
 Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in low numbers of females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
+Certain combinations of variables in the simulated data resulted in captures of few females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
 These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure.
-For example, in high selectivity scenarios for gummy sharks \~10% of individuals were immature (Figure S9) compared to \~50% in the corresponding scenarios for school sharks.
-As $P_{Max}$ decreased the proportion of both maternal females and immature females in the analysis also decreased.
-The latter effect was due to the right shift in length structure associated with the proportionally fewer juveniles in the population.
+For example, in low selectivity scenarios for gummy sharks \~60% of individuals were immature  compared to \~50% in the corresponding scenarios for school sharks (Figure S9).
+As $P_{Max}$ decreased the proportion of maternal females in the analysis also decreased.
 
 Overall, the 2PLF-maternity method performed the worst.
-Ignoring $P_{Max}$ was equivalent to introducing a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3).
+Ignoring $P_{Max}$ introduced a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3).
 This manifested in an overestimation of $L'_{50}$ that was exacerbated by gear selectivity effects (Figure S2).
@@
 $L_{50}$ could usually be estimated with a higher precision than $L'_{50}$.
-For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2% and exceeding the performance of the 3PLF-methods at sample sizes \< 250.
+For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2% and exceeding the performance of the 3PLF-methods at most sample sizes.
 
-The effect of the different methods in ultimately calculating $R_0$ varied considerably and was strongly influenced by life history and selectivity characteristics.
-Greater precision was achieved with the 3PLF methods, with the 3PLF-free method generally performing best (Figure 5, Figure S11).
-For gummy sharks the 3PLF-fixed method was the preferred model at lower sample sizes and lower values of $P_{Max}$, consistent with decreasing performance of 3PLF methods as $P_{Max}$ decreased (Figure 2, Figure 3).
+The effect of the different methods in ultimately calculating $R_0$ varied considerably and was also influenced by life history and selectivity characteristics.
+For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios for the school shark (Figure 5, Figure S11).
+In contrast, using a fixed $P_{Max}$ or substituting maturity parameters often led to better performance for the gummy shark (Figure 5, Figure S11).
+
+```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in minimising bias in calculations of $R_0$.
+The preferred method was that which minimised bias, $|\\text{relative error}|$ across 300 simulated datasets.
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}
+
+read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs(rel_error)) %>% 
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_viridis_d() + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
+```
+
+Greater precision in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method  performing best in most scenarios tested (Figure 6, Figure S11).
+For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$, consistent with decreasing performance of 3PLF methods as $P_{Max}$ decreased (Figure 2, Figure 3).
 
@@
 
-Calculations of $R_0$ for the 3PLF methods were generally unbiased with sample sizes $\geq 250$, and neither method clearly outperformed the other (Figure S10, Figure 6).
-As anticipated, the 2PLF-maternity method performed the worst overall and ignoring $P_{Max}$ consistently led to $R_0$ being overestimated.
-This effect was particularly pronounced for gummy sharks under low selectivity scenarios where derived $R_0$ was more than double the true value.
-Using maturity data to approximate maternal data (2PLF-maturity method) also consistently overestimated $R_0$, as expected since $L_{50}$ was less than $L'_{50}$.
-However, the magnitude of bias was relatively small in some cases.
-For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, this method frequently outperformed the 3PLF methods (Figure 6).
-
-```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in minimising bias in calculations of $R_0$.
-The preferred method was that which minimised bias, $|\\text{relative error}|$ across 250 simulated datasets.
-Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}
-
-read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
-  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
-  slice_min(abs(rel_error)) %>% 
-  ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_viridis_d() + 
-  guides(fill = guide_legend(title = "Method")) + 
-  facet_wrap(~species, nrow = 1) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90)) + 
-  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
-```
-
 ## Empirical case study
@@
 Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$Stage > 2))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$Stage > 2))))`% were in maternal condition.
-Using the 3PLF-free method the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
+Using the 3PLF-estimated method the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
 Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence.
-Comparison of 3PLF-fixed models with values of $P_{Max} = 0.33$ and $P_{Max} = 0.5$ showed much greater support for a biennial cycle ($\Delta_{AIC}$ = 9.4).
-The model with $P_{Max} = 0.5$ also outperformed the 3PLF-free method ($\Delta_{AIC}$ = 1.66).
+Comparison of 3PLF-fixed models with values of $P_{Max} = 0.33$ and $P_{Max} = 0.5$ showed much greater support for a biennial cycle ($\Delta_{AIC}$ = 14.15).
+The model with $P_{Max} = 0.5$ also outperformed the 3PLF-estimated method ($\Delta_{AIC}$ = 1.86).
 Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data [@burnham_model_2002].
 
-```{r fig.width = 4.5, fig.height= 5, fig.cap= "Comparison of 3PLF-free and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C.
+```{r fig.width = 4.5, fig.height= 5, fig.cap= "Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C.
 plumbeus*, in the Gulf of Mexico and Western North Atlantic.
@@
 
-preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - free") %>% 
-  mutate(label2 = "3PLF - free")
+preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - estimated") %>% 
+  mutate(label2 = "3PLF - estimated")
 
@@
 Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that precision, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated.
-Using a fixed value for $P_{Max}$ resulted in lower bias at low sample sizes.
-This study also demonstrated that recruitment was overestimated when maturity data were used to approximate maternal data and if $P_{Max}$ was ignored when estimating maternity parameters.
+Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
+This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and if $P_{Max}$ was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 
-Using simulated data to compare the relative performance of the 3PLF-free and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
-In most scenarios sample sizes of at least 250 were needed for the 3PLF-free method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below 250 it is therefore advisable to fix $P_{Max}$.
-Due to the decline in performance associated with longer reproductive cycles, sample sizes of at least 500 are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle.
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified some situations where it may be preferable to fix $P_{Max}$.
+In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
+At sample sizes below this is therefore advisable to fix $P_{Max}$.
+Due to the decline in performance associated with longer reproductive cycles, larger sample sizes are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle.
 
 While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
-Outcomes of the simulation study suggested that samples sizes of $\geq1000$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species.
+Outcomes of the simulation study suggested that samples sizes of $\geq500$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species.
 This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999].
@@
 This study also confirms that fitting a 2PLF to maternal data when $P_{Max} < 1$ (effectively ignoring $P_{Max}$) is likely to result in biased parameters and can also overestimate recruitment.
-While not widely used, this approach has been undertaken in several studies [@mejia-falla_reproductive_2012; @baremore_reproduction_2012; @baremore_reproduction_2013; @rambahiniarison_life_2018].
+While not widely used, this approach has been undertaken in several studies [@mejiafalla_reproductive_2012; @baremore_reproduction_2012; @baremore_reproduction_2013; @rambahiniarison_life_2018].
 
-The success of any analysis using the 3PLF-free method, and its ability to outperform the 3PLF-fixed method, ultimately depends on the analyst choosing a suitable value for $P_{Max}$.
+The success of any analysis using the 3PLF-estimated method, and its ability to outperform the 3PLF-fixed method, ultimately depends on the analyst choosing a suitable value for $P_{Max}$.
 In this study, the value for $P_{Max}$ used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data.
@@
 
-The empirical analysis carried out using data for the sandbar shark illustrates the potential for the 3PLF-free method to provide novel insights from maternal data.
+The empirical analysis carried out using data for the sandbar shark illustrates the potential for the 3PLF-estimated method to provide novel insights from maternal data.
 Detailed reproductive studies of the western North Atlantic population by Baremore and Hale [-@baremore_reproduction_2012] and Piercy *et al* [-@piercy_histological_2016] were ambiguous about the frequency of reproduction.
-In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or longer at the population level.
-Reanalysis of these combined data sets using the 3PLF-free method, however, resulted in a maximum likelihood estimate of $P_{Max} = 0.48$, most consistent with biennial cycle.
-Using AIC values to compare the relative plausibility of 3PLF-fixed models corresponding to biennial and triennial cycles also strongly supported a biennial cycle.
+In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
+Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of $P_{Max} = 0.48$.
+Similarly, based on AIC values, a fixed value of $P_{Max} = 0.5$ had higher relative plausibility than a value of 0.33 (Table 2).
 
-This counter intuitive result appears to be due to the gradual attainment of maternal status as a function of length in sandbar sharks.
-$L'_{5}$ and $L'_{95}$ occur at 146cm and 176cm, respectively, a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013].
-It therefore appears that the shape of the maternity curve is not just due to individuals differing in their size at maturity, but can also be attributed to smaller females reproducing less frequently.
-Larger females, in contrast, appear capable of reproducing biennially.
-This finding has important management implications that would not otherwise be apparent if using a maturity curve to approximate maternity data.
-Not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently as well.
-This implies removal of larger and older female sharks could have a disproportionately greater impact on population productivity.
+This somewhat surprising result may stem from the gradual attainment of asymptotic maternal status as a function of length.
+Maturity in sandbar sharks appears to occur over an extended period; $L_{5}$ and $L_{95}$ are 146cm and 176cm, a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013].
+With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, with the largest females ultimately capable of reproducing biennially.
+Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management of this stock; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
+This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.
 
@@
 
-Despite the simplicity of the 3PLF-free method, it nonetheless represents a conceptual shift for chondrichthyan reproductive and population biology.
+Despite the simplicity of the 3PLF-estimated method, it nonetheless represents a conceptual shift for chondrichthyan reproductive and population biology.
 Until now reproductive periodicity has largely been determined qualitatively from observations of female reproductive biology or inferred based on circumstantial evidence.
-Using the 3PLF-free method to estimate $P_{Max}$ within a maternity function transforms it from a nuisance parameter to one of direct inferential interest that can be seen as a valuable output of the modelling process itself.
+Using the 3PLF-estimated method to estimate $P_{Max}$ within a maternity function transforms it from a nuisance parameter to one of direct inferential interest that can be seen as a valuable output of the modelling process itself.
 In addition to the potential benefits of using this method to improve estimation of maternal parameters, there are also a range of other advantages that arise from estimating $P_{Max}$.
 
-One major benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to $P_{Max}$.
+One benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to $P_{Max}$.
 Such assumptions can have important implications for population modelling.
 For example, to account for uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, *Urolophus paucimaculatus*, Trinnie *et al* (2014) estimated maternity parameters for ten plausible scenarios.
-In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle.
+In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle [@sedar_sedar_2017].
 
@@
 
-Estimation of $P_{Max}$ using the 3PLF-free method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans.
-White sharks and whale shark are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
+Estimation of $P_{Max}$ using the 3PLF-estimated method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans.
+White sharks, *Carcharodon carcharias*, and whale shark, *Rhinchodon typus*,are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
 In both species lack of data on reproductive frequency is a major impediment to population modelling and the development of management strategies [@huveneers_future_2018; @bowlby_implications_2020].
-With many of these populations now the focus of ongoing monitoring, non-lethal methods for assessing maternal state may eventually provide a way of collecting sufficient data for estimation of $P_{Max}$ [@sulikowski_seasonal_2016].
+With many of these populations the focus of ongoing monitoring, non-lethal methods for assessing maternal state may eventually provide a way of collecting sufficient data for estimation of $P_{Max}$ [@sulikowski_seasonal_2016].
 

### 2024-01-28 rmds/figures-supp.Rmd
changes

@@
 ```
-Figure S1.
+Figure S1. Convergance success of alternative logistic regression models fit (n = 300) for school shark and gummy shark.
 
@@
   background_grid() +
-  labs(x = "Sample size", y = "Mean % relative error", title = "Bias - Length at 50% maternity") +
+  labs(x = "Sample size", y = "Mean % relative error", title = expression("Bias - Length at 50% maternity")) +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
@@
 ```
-Figure S2.
+Figure S2. Bias (per cent relative error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
 
@@
 ```
-Figure S3. Note 2PLF-maturity and 3PLF-fixed methods are identical
+Figure S3.
+Bias (per cent relative error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets.
+Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical.
 
@@
   background_grid() +
-  labs(x = "Sample size", y = "Mean % absolute relative error", title = "Precision - Length at 50% maternity") +
+  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - Length at 50% maternity")) +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
-Figure S4.
+Figure S4. Precision (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
 
@@
 Figure S5.
+Precision (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets.
+Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical.
 \newpage
@@
   background_grid() +
-  labs(x = "Sample size", y = "% interval coverage", title = "Interval coverage - Length at 50% maternity") +
+  labs(x = "Sample size", y = "% interval coverage", title = expression("Interval coverage - Length at 50% maternity")) +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
@@
 Figure S6.
+Confidence interval coverage for $L'_{50}$ for 3PLF methods.
+Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
 \newpage
@@
 Figure S7.
+Confidence interval coverage for $P_{Max}$ for the 3PLF-estimated method.
+Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
 \newpage
@@
 Median number of females in maternal condition in simulated data that were used to estimate maternity parameters subject to varying sample size, gear selectivity, reproductive frequency.
-250 iterations were carried out for each simulation.
+300 iterations were carried out for each simulation.
 \newpage
@@
 Median number of immature females in simulated data that were used to estimate maternity parameters subject to varying sample size, gear selectivity, reproductive frequency.
-250 iterations were carried out for each simulation.
+300 iterations were carried out for each simulation.
 \newpage
@@
 ```
+Figure S10.
+Bias (per cent relative error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets
 
@@
 ```
+Figure S11.
+Precision (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets

### 2024-01-28 rmds/figures.Rmd
changes

@@
 ---
-output: pdf_document
+output: 
+  pdf_document:
+    latex_engine: "pdflatex"
 ---
@@
 
-```{r out.width = '100%', fig.cap = "Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. "}
+```{r out.width = '100%'}
 knitr::include_graphics(here("img","figure1.png"))
@@
 
-\newpage
+Figure 1. Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. Illustrations © R.Swainston/www.anima.net.au
+
 ```{r fig.width = 7, fig.height= 5.5}
@@
 ```
-
 Figure 2.
-Per cent relative error (bias) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods.
-Each point represents parameter estimates one iteration of simulated data.
+Bias (per cent relative error) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods.
+Each point represents parameter estimates from one iteration of simulated data.
 Bias tended to increase as the true underlying $P_{Max}$ decreased.
+\
+\
 \newpage
 
-```{r}
-data <- read_rds(here("data", "simulations-slim.Rds"))
+```{r fig.width = 7, fig.height= 5.5}
+data <- read_rds(here("data", "simulations-summarised.Rds"))
 
-data <- data %>% 
-    mutate(method_long = case_when(
-    method %in% "2" ~ "2PLF - maturity",
-    method %in% "2a" ~ "2PLF - maternity",
-    method %in% "3" ~ "3PLF - free",
-    method %in% "3a" ~ "3PLF - fixed"
-  )) %>%
-  mutate(mesh_name_long = case_when(
-    mesh_name %in% "high" ~ "High selectivity",
-    mesh_name %in% "low" ~ "Low selectivity"
-  )) %>% 
-  filter(mesh_name %in% "high", par_name %in% "r_0") %>% 
-  filter(Nsamples %in% c(50, 250, 1000))
+data %>% 
+  filter(par_name %in% "c") %>% 
+  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
+  filter(mesh_name %in% "high") %>% 
+  ggplot() +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
+  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
+  facet_grid(species ~ method_long) + 
+  scale_color_brewer(palette = "Spectral") + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  theme_bw() + 
+  labs(x = "Sample size", y = "Mean % absolute relative error") +
+  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+```
+Figure 3.
+Precision (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods.
+Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased.
+Each point reflects a mean value from 300 simulated data sets.
+\
+\
+\newpage
 
-data <- data %>% 
-  mutate(method_long = fct_relevel(
-    method_long, "3PLF - free",
-    "3PLF - fixed",
-    "2PLF - maternity",
-    "2PLF - maturity"
-  ))
+```{r fig.width = 7, fig.height= 5.5}
+data %>% 
+  filter(par_name %in% "m50", method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
+  filter(mesh_name %in% "high") %>% 
+  ggplot() +
+  geom_hline(yintercept = 50, linetype = "dashed") +
+  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
+  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
+  facet_grid(species ~ method_long, scales = "free_y") + 
+  scale_color_brewer(palette = "Spectral") + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  theme_bw() + 
+  labs(x = "Sample size", y = "% interval coverage") +
+  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+```
+Figure 4.
+Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
+Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
+\
+\
+\newpage
 
-ggplot(data) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_boxplot(aes(x = Nsamples, y = abs_error, fill = repro_freq), outlier.shape = NA) +
-  facet_grid(method_long ~ species) + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(fill = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
-  cowplot::theme_half_open() +
-  panel_border() +
-  background_grid() +
-  labs(x = "Sample size", y = "Median % absolute error") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
-  ylim(0, 80)
+```{r fig.width = 9, fig.height= 3.5}
+
+read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs(rel_error)) %>% 
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_viridis_d() + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
 ```
+Figure 5.
+Performance of alternative maternity functions in minimising bias in calculations of $R_0$.
+The preferred method was that which minimised bias, $|\text{relative error}|$ across 300 simulated datasets.
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
+\
+\
+\newpage
 
+```{r fig.width = 9, fig.height= 3.5}
+
+read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs_error) %>%
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_viridis_d() + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
+```
+Figure 6.
+Performance of alternative maternity functions in accurately calculating $R_0$.
+The best performing method was that which minimised mean absolute error across 300 simulated datasets.
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
+\
 \newpage
-```{r fig.width = 10, fig.height= 7}
+
+```{r fig.width = 4.5, fig.height= 5}
 # Load data - Sandbar
@@
 # Bin data into length categories for plotting
-brks = seq(40, 220, 8)
+brks = seq(40, 220, 7.5)
 data_binned <- data %>%
@@
 
-preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - free") %>% 
-  mutate(label2 = "italic('C. plumbeus ')~`- 3PLF  free`")
+preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - estimated") %>% 
+  mutate(label2 = "3PLF - estimated")
 
@@
   ylim(0, 1) +
-  theme_cowplot() + 
+  theme_bw() + 
   labs(x = NULL, y = NULL) + 
-  facet_wrap(~label2, labeller = label_parsed) + 
+  facet_wrap(~label2) + 
   theme(strip.text = element_text(hjust = 0.5))
@@
 preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
-  mutate(label2 = "italic('C. plumbeus')~` - 3PLF fixed`")
+  mutate(label2 = "3PLF - fixed")
 
@@
   ylim(0, 1) +
-  theme_cowplot() + 
+  theme_bw() + 
   labs(x = NULL, y = NULL) + 
-  facet_wrap(~label2, labeller = label_parsed) + 
+  facet_wrap(~label2) + 
   theme(strip.text = element_text(hjust = 0.5))
 
-# Load data - blacktip
-data <- read_rds(here("data", "empirical-limbatus.Rds"))
+grid.arrange(p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 12, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 12, font = 8)))
+```
 
-# Bin data into length categories for plotting
-brks = seq(40, 160, 8)
-data_binned <- data %>%
-  mutate(x_bin = findInterval(x, brks)) %>% 
-  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
-  group_by(x_bin) %>%
-  summarise(p = sum(z)/ n())
-
-preds3 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - free") %>%
-  mutate(label2 = "italic('C. limbatus')~` - 3PLF free`")
-
-# Plot
-p3 <- ggplot()  + 
-  geom_ribbon(data = preds3, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
-  geom_line(data = preds3, aes(x = x, y = y)) +
-  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
-  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
-  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
-  ylim(0, 1) +
-  theme_cowplot() + 
-  labs(x = NULL, y = NULL) + 
-  facet_wrap(~label2, labeller = label_parsed) + 
-  theme(strip.text = element_text(hjust = 0.5))
-
-preds4 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
-  mutate(label2 = "italic('C. limbatus')~` - 3PLF fixed`")
-
-p4 <- ggplot()  + 
-  geom_ribbon(data = preds4, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
-  geom_line(data = preds4, aes(x = x, y = y)) +
-  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
-  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
-  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
-  ylim(0, 1) +
-  theme_cowplot() + 
-  labs(x = NULL, y = NULL)  + 
-  facet_wrap(~label2, labeller = label_parsed) + 
-  theme(strip.text = element_text(hjust = 0.5))
-
-grid.arrange(p3, p4, p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 16, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 16, font = 8)))
-```
-*Figure 3.
-Comparison of 3PLF-free and -fixed methods used to estimate maternal parameters for blacktip shark, *C.
-limbatus*, and sandbar shark, *C.
-plumbeus* in the Gulf of Mexico and Western North Atlantic.
-Sold line is the expected proportion in maternal condition at fork length.
+Figure 7.
+Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C.
+plumbeus*, in the Gulf of Mexico and Western North Atlantic.
+Solid line is the expected proportion in maternal condition at length, $\Psi'(L)$.
 The grey shaded region denotes 95% confidence intervals based on bootstrap resampling.
-A fixed value of $P_{Max} = 0.5$ was used for both species.*
+$P_{Max}$ was fixed at 0.5 in the lower panel

### 2024-01-28 rmds/manuscript.Rmd
changes

@@
 - name: Andrew Piercy
-  affiliation: Valencia College
+  affiliation: College of Southern Maryland, La Plata MD 20646, United States
   number: 3
@@
 output:
-  word_document: default
   bookdown::pdf_document2:
@@
 Under such an approach all mature females are tacitly assumed to reproduce in each breeding season [@cortes_demographic_1998].
-Non-annual reproduction is accounted for by weighting fecundity or the maturity function by the assumed (but often unknown) duration of the reproductive cycle [-@walker_reproduction_2005].
+Non-annual reproduction is accounted for by weighting fecundity or the maturity function by the assumed (but often unknown) duration of the reproductive cycle [@walker_reproduction_2005].
 In some cases, calculations of reproductive output may use the age-at-first-reproduction, defined as the mean age at maturity plus the gestation period [@mollet_comparative_2002].
@@
 
-```{r out.width = '100%', fig.cap = "Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. "}
-knitr::include_graphics(here("img","figure1.png"))
-```
-
 ## Simulation
@@
 
-where mean length, $\mu$, was caclulated from the growth curve.
-The standard deviation, $\sigma$ was not available for either species so was assumed to be directly proportional to length, with a constant of proportionality of 5% (Table 1), i.e.
-$\sigma = 0.05 \mu$.
+where mean length, $\mu$, was calculated from the growth curve.
+The standard deviation, $\sigma$ was not available for either species so was assumed to be directly proportional to length, with a constant of proportionality of 5% i.e.
+$\sigma = 0.05 \mu$ (Table 1).
 
@@
 
-where the elements of the matrix followed the general form of Punt *et al* [-@punt_estimating_1997] and are given by$$ g_{k,j} = \left\{
+where the elements of the matrix followed the general form of Punt et. al. [-@punt_estimating_1997] and are given by$$ g_{k,j} = \left\{
     \begin{array}{ll}
@@
 
-Fishing mortality, *F*, in length class *l*, was calculated as $$F(l) = S(l) \cdot F$$where $F$ was assumed to be constant and nominally set to 25% of natural mortality, $M$ (i.e.
+Fishing mortality, *F*, in length class $l$, was calculated as $$F(l) = S(l) \cdot F$$where $F$ was assumed to be constant and nominally set to 25% of natural mortality, $M$ (i.e.
 $0.25M$).
 The effects of varying levels of $F$ were not investigated further in this study.
-Total mortality, *Z,* in length class, *l*, was further given by $$Z(l) = F(l)\cdot{M} $$
+Total mortality, *Z,* in length class, $l$, was further given by $$Z(l) = F(l)\cdot{M} $$
 
-The number of fish, $N$, per recruit in length class, *l,* that survived to age, *a*, was calculated as
+The number of fish, $N$, per recruit in length class, $l$, that survived to age, $a$, was calculated as
 
-$$ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=1<a<A_{Max} \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A_{Max} \\    \end{array}\right.$$
+$$ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=0<a<A_{Max} \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A_{Max} \\    \end{array}\right.$$
 
-where $A_{Max}$ is the maximum age (Table 1). The estimated numbers of fish caught in length class, l, at age, a, was calculated using the Baranov catch equation as
+where $A_{Max}$ is the maximum age (Table 1). The estimated numbers of fish caught in length class, $l$, at age, $a$, was calculated using the Baranov catch equation as
 
@@
 
-Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This processe was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.
+Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This process was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.
 
@@
 With lower values of $P_{Max}$ it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying $P_{Max}$ decreased there was a greater potential for positive bias in $\hat{P_{Max}}$ (Figure 2).
-
-```{r fig.width = 7, fig.height= 5.5, fig.cap= "Bias (per cent relative error) in parameter estimates for $\\hat{L'_{50}}$ and $\\hat{P_{Max}}$ for 3PLF methods.
-Each point represents parameter estimates from one iteration of simulated data.
-Bias tended to increase as the true underlying $P_{Max}$ decreased."}
-data <- read_rds(here("data", "simulations-pivot.Rds"))
-
-ggplot() + 
-  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
-  scale_color_viridis_d(option = "D", direction = 1) + 
-  facet_grid(species~method_long) + 
-  ylim(-30, 50) +
-  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
-  theme_bw() +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
-  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
-  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
-```
+As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$ (Figure 2).
 
@@
 
-```{r fig.width = 7, fig.height= 5.5, fig.cap = "Precision (per cent absolute error) in parameter estimates  of $\\hat{P_{Max}}$ for 3PLF methods.
-Large sample sizes were needed to accurately estimate $\\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased "}
-data <- read_rds(here("data", "simulations-summarised.Rds"))
-
-data %>% 
-  filter(par_name %in% "c") %>% 
-  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
-  filter(mesh_name %in% "high") %>% 
-  ggplot() +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
-  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
-  facet_grid(species ~ method_long) + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
-  theme_bw() + 
-  labs(x = "Sample size", y = "Mean % absolute relative error") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
-```
-
 While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6).
-For most simulations, interval coverage for $L'_{50}$ was well below the expected level of 50%, irrespective of sample size.
-In contrast, for the 3PLF-estimated method interval coverage oscillated around 50% for both $L'_{50}$ and $P_{Max}$.
-
-```{r fig.width = 7, fig.height= 5.5, fig.cap = "Interval coverage for $\\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
-Figure shows the percentage of simulations where the true parameter value fell within the 50% bootstrap confidence interval.
-"}
-data %>% 
-  filter(par_name %in% "m50", method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
-  filter(mesh_name %in% "high") %>% 
-  ggplot() +
-  geom_hline(yintercept = 50, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
-  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
-  facet_grid(species ~ method_long, scales = "free_y") + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
-  theme_bw() + 
-  labs(x = "Sample size", y = "% interval coverage") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
-```
+For most simulations, interval coverage for $\hat{L'_{50}}$ was well below the expected level of 50%, irrespective of sample size.
+In contrast, for the 3PLF-estimated method interval coverage oscillated around 50% for both $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
 
 Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in captures of few females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
+Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
 These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure.
@@
 
-Overall, the 2PLF-maternity method performed the worst.
+Overall, the 2PLF-maternity method performed the poorest.
 Ignoring $P_{Max}$ introduced a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3).
@@
 $L_{50}$ could usually be estimated with a higher precision than $L'_{50}$.
-For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2% and exceeding the performance of the 3PLF-methods at most sample sizes.
+For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2%,  exceeding the performance of the 3PLF-methods at most sample sizes.
 
-The effect of the different methods in ultimately calculating $R_0$ varied considerably and was also influenced by life history and selectivity characteristics.
-For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios for the school shark (Figure 5, Figure S11).
-In contrast, using a fixed $P_{Max}$ or substituting maturity parameters often led to better performance for the gummy shark (Figure 5, Figure S11).
-
-```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in minimising bias in calculations of $R_0$.
-The preferred method was that which minimised bias, $|\\text{relative error}|$ across 300 simulated datasets.
-Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}
-
-read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
-  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
-  slice_min(abs(rel_error)) %>% 
-  ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_viridis_d() + 
-  guides(fill = guide_legend(title = "Method")) + 
-  facet_wrap(~species, nrow = 1) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90)) + 
-  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
-```
-
-Greater precision in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method  performing best in most scenarios tested (Figure 6, Figure S11).
-For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$, consistent with decreasing performance of 3PLF methods as $P_{Max}$ decreased (Figure 2, Figure 3).
-
-```{r fig.width = 9, fig.height= 3.5, fig.cap = "Performance of alternative maternity functions in accurately calculating $R_0$.
-The best performing method was that which minimised mean absolute error across 250 simulated datasets.
-Note 2PLF-maternity (Annual) scenarios were excluded for this comparison."}
-
-read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
-  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
-  slice_min(abs_error) %>%
-  ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_viridis_d() + 
-  guides(fill = guide_legend(title = "Method")) + 
-  facet_wrap(~species, nrow = 1) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90)) + 
-  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
-```
+The effect of the different methods in ultimately calculating $R_0$ varied considerably across the various scenarios.
+For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10).
+Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of $R_0$ by 3.4% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1%.
+(Table S1, Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10).
+Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10).
+For both species, better precision in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$.
 
@@
 Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$Stage > 2))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$Stage > 2))))`% were in maternal condition.
-Using the 3PLF-estimated method the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
+Using the 3PLF-estimated method, the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
 Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence.
-Comparison of 3PLF-fixed models with values of $P_{Max} = 0.33$ and $P_{Max} = 0.5$ showed much greater support for a biennial cycle ($\Delta_{AIC}$ = 14.15).
+For 3PLF-fixed models a value of $P_{Max} = 0.5$ was strongly supported ($\Delta_{AIC}$ = 14.15) over the alternative fixed value of $P_{Max} = 0.33$.
 The model with $P_{Max} = 0.5$ also outperformed the 3PLF-estimated method ($\Delta_{AIC}$ = 1.86).
 Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data [@burnham_model_2002].
-
-```{r fig.width = 4.5, fig.height= 5, fig.cap= "Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C.
-plumbeus*, in the Gulf of Mexico and Western North Atlantic.
-Solid line is the expected proportion in maternal condition at length, $\\Psi'(L)$.
-The grey shaded region denotes 95% confidence intervals based on bootstrap resampling.
-$P_{Max}$ was fixed at 0.5 in the lower panel"}
-# Load data - Sandbar
-data <- read_rds(here("data", "empirical-plumbeus.Rds"))
-preds <- read_rds(here("data", "empirical-predictions.Rds"))
-
-# Bin data into length categories for plotting
-brks = seq(40, 220, 7.5)
-data_binned <- data %>%
-  mutate(x_bin = findInterval(x, brks)) %>% 
-  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
-  group_by(x_bin) %>%
-  summarise(p = sum(z)/ n())
-
-preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - estimated") %>% 
-  mutate(label2 = "3PLF - estimated")
-
-p1 <- ggplot()  + 
-  geom_ribbon(data = preds1, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
-  geom_line(data = preds1, aes(x = x, y = y)) +
-  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
-  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
-  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
-  ylim(0, 1) +
-  theme_bw() + 
-  labs(x = NULL, y = NULL) + 
-  facet_wrap(~label2) + 
-  theme(strip.text = element_text(hjust = 0.5))
-
-preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
-  mutate(label2 = "3PLF - fixed")
-
-# Plot
-p2 <- ggplot()  + 
-  geom_ribbon(data = preds2, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
-  geom_line(data = preds2, aes(x = x, y = y)) +
-  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
-  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
-  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
-  ylim(0, 1) +
-  theme_bw() + 
-  labs(x = NULL, y = NULL) + 
-  facet_wrap(~label2) + 
-  theme(strip.text = element_text(hjust = 0.5))
-
-grid.arrange(p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 12, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 12, font = 8)))
-```
 
@@
 Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
-This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and if $P_{Max}$ was ignored when estimating maternity parameters.
+This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified some situations where it may be preferable to fix $P_{Max}$.
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
 In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this is therefore advisable to fix $P_{Max}$.
-Due to the decline in performance associated with longer reproductive cycles, larger sample sizes are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle.
+At sample sizes below this it is therefore advisable to fix $P_{Max}$.
+Larger sample sizes are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
 
@@
 
-The success of any analysis using the 3PLF-estimated method, and its ability to outperform the 3PLF-fixed method, ultimately depends on the analyst choosing a suitable value for $P_{Max}$.
+The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for $P_{Max}$.
 In this study, the value for $P_{Max}$ used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data.
@@
 
-The empirical analysis carried out using data for the sandbar shark illustrates the potential for the 3PLF-estimated method to provide novel insights from maternal data.
+The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating $P_{Max}$.
 Detailed reproductive studies of the western North Atlantic population by Baremore and Hale [-@baremore_reproduction_2012] and Piercy *et al* [-@piercy_histological_2016] were ambiguous about the frequency of reproduction.
 In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
-Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of $P_{Max} = 0.48$.
-Similarly, based on AIC values, a fixed value of $P_{Max} = 0.5$ had higher relative plausibility than a value of 0.33 (Table 2).
+Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of $\hat{P_{Max}} = 0.48$, considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale [-@baremore_reproduction_2012].
+Similarly, based on AIC values, a fixed value of $P_{Max} = 0.5$ had a much higher relative plausibility than a value of 0.33 (Table 2).
 
-This somewhat surprising result may stem from the gradual attainment of asymptotic maternal status as a function of length.
-Maturity in sandbar sharks appears to occur over an extended period; $L_{5}$ and $L_{95}$ are 146cm and 176cm, a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013].
-With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, with the largest females ultimately capable of reproducing biennially.
-Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management of this stock; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
+This unexpected result may be related to the gradual attainment of asymptotic maternal status.
+Maturity in sandbar sharks appears to occur over an extended size range.
+$L_{5}$ and $L_{95}$ are 146cm and 176cm – a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013].
+With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
+The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size [@hale_age_2013].
+Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
 This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.
@@
 
-John Carlson
+We thank John Carlson and all NMFS staff who were involved in collecting the sandbar shark reproductive data used in this study.
+We also thank Alex Hesp for valuable comments on an early draft of this paper and generosity in sharing R code.
 

### 2024-01-28 run_all.R

@@ -11,10 +11,10 @@ source(here::here("code", "empirical.R"))
 
 
 # Generate files
-render("rmds/manuscript.Rmd", output_format = "all", output_dir = "reports", envir = new.env())
-render("rmds/figures.RmD", output_format = "all", output_dir = "reports", envir = new.env())
-render("rmds/figures-supp.RmD", output_format = "all", output_dir = "reports", envir = new.env())
-render("rmds/tables.RmD", output_format = "all", output_dir = "reports", envir = new.env())
+render("rmds/manuscript.Rmd", output_format = "all", output_file = "main-text.pdf", output_dir = "reports", envir = new.env())
+render("rmds/figures.RmD", output_format = "all", output_file = "figs.pdf", output_dir = "reports", envir = new.env())
+render("rmds/figures-supp.RmD", output_format = "all", output_file = "fig-supp.pdf", output_dir = "reports", envir = new.env())
+render("rmds/tables.RmD", output_format = "all", output_file = "tabs.pdf", output_dir = "reports", envir = new.env())
 
 ## Zotero citation key format
 # [auth:lower]_[Title:lower:skipwords:select,1,1]_[year]

### 2024-02-02 rmds/manuscript.Rmd
manuscript

@@
   number: 1
-- name: Ivy Baremore
+- name: Ivy E Baremore
   affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
   number: 2
-- name: Andrew Piercy
+- name: Andrew N Piercy
   affiliation: College of Southern Maryland, La Plata MD 20646, United States
   number: 3
-abstract: |
-  For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to calculate the average probability of a female giving birth or laying eggs in a season.
-Although fundamental to calculating the reproductive capacity of a population, relatively few studies report maternity functions.
-Instead, maturity functions have typically been used as a proxy, despite evidence of a delay between attainment of functional maturity and pregnancy in some species.
-This study examined the relative performance of alternative approaches for quantifying maternal reproductive output.
-Applying logistic models to a combination of simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-At sample sizes < 250 individuals a fixed-value for $P_{Max}$ was more effective at reducing bias in simulated data for gummy sharks, *Mustelus antarcticus*.
-Maturity parameters could be estimated with greater precision however substituting them in place of maternity parameters overestimated lifetime reproductive output.
-The extent to which maturity functions can provide a good approximation for maternity functions may be species specific and requires futher research.
-Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
-We proposed that estimation of $P_{Max}$ is generally preferable to using a fixed value.
-In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
-
 toc: false
 output:
+  word_document: default
   bookdown::pdf_document2:
@@
 
+**Abstract**
+
+For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season.
+Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy.
+Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output.
+Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
+In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
+This study suggests that for sample sizes \> 100 estimation of $P_{Max}$ is generally preferable to using a fixed value.
+
+**Key words** Logistic regression; reproductive biology; maturity ogive; fisheries management
+
+\newpage
+
 # Introduction
 
-Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass [@quinn_quantitative_1999].
+Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass [@hilborn1992; @quinn_quantitative_1999].
 Calculating reproductive capacity requires data on the sexually mature proportion of the population contributing to reproduction at a given size or age.
@@
 
-In chondrichthyan fishes (sharks, rays, skates, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
+In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
 Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015].
@@
 
-# Methods
+# Materials and methods
 
@@
 
-A logical progression from estimating $P_{Max}$ is the consideration of alternative functional forms for this parameter such as time- or space-varying $P_{Max}$ or more complex length- or age-dependent forms.
+A logical progression from estimating $P_{Max}$ is the consideration of alternative functional forms for this parameter such as time- or space-varying $P_{Max}$ or more complex length- or age-dependent forms [@winton2014].
 Despite the diverse range of reproductive modes found in chondrichthyan fishes, the current, limited knowledge of maternal investment has typically investigated variables such as litter size and pup condition [@hussey_maternal_2010].
@@
 
+# Competing interests
+
+The authors declare there are no competing interests.
+
+# Data availability
+
+Data generated or analyzed during this study are available from the corresponding author upon reasonable request.
+
 # References

### 2024-05-01 README.qmd
README

---
title: "Maternity Function"
format: gfm
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-05-01 code/simulation.R

@@ -118,7 +118,6 @@ data <- read_rds("data/generated_data.rds")
 
 
 data <- data %>%
-  slice(41401:43200) %>% 
   # Simulate maternal data
   mutate(data = map2(pars, data, generate_maternal_data)) %>% # How many maternal females are there in the simulated data
   mutate(n_maternal = map_dbl(data, ~ sum(.$z))) %>%

### 2024-05-01 run_all.R

@@ -5,7 +5,7 @@ library(knitr)
 library(rmarkdown)
 
 # Data analysis
-source(here::here("code", "simulation.R")) # 48 hour run time on my computer
+source(here::here("code", "simulation.R"))
 source(here::here("code", "summarise-results.R"))
 source(here::here("code", "empirical.R"))
 

### 2024-05-19 README.qmd
readme
revisions

@@
 ---
-title: "Maternity Function"
+title: "Quantifying maternal reproductive output of chondrichthyan fishes"
 format: gfm
-editor: visual
 ---
-
-## Quarto
 
@@
 
-## Running Code
+## Example
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+Load data for sandbar shark maternity data set. Rows are data from individuals sharks x is fork length (cm) and z is maternal status (0 = non maternal, 1 = maternal).
 
-```{r}
-1 + 1
+```{r message=FALSE, warning=FALSE}
+library(tidyverse)
+
+data <- read_rds(here::here("data", "empirical-plumbeus.Rds"))
 ```
 
-You can add options to executable code like this
+## Bin data into suitable length categories for plotting
 
-```{r}
-#| echo: false
-2 * 2
+```{r message=FALSE, warning=FALSE}
+brks = seq(40, 220, 7.5)
+data_binned <- data %>%
+  mutate(x_bin = findInterval(x, brks)) %>% 
+  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
+  group_by(x_bin) %>%
+  summarise(p = sum(z)/ n())
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+Compile 3 parameter logistic model
+
+```{r message=FALSE, warning=FALSE}
+library(TMB)
+compile("code/logistic3.cpp")
+dyn.load(dynlib("code/logistic3"))
+```

### 2024-05-19 reports/main-text.tex
revisions

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Quantifying maternal reproductive output of chondrichthyan fishes},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\fontsize{12}{20}
\usepackage{lineno}
\linenumbers
\linenumberfont
\renewcommand\linenumberfont{\normalfont\small\sffamily}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[L]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}
\title{Quantifying maternal reproductive output of chondrichthyan fishes}
\usepackage{authblk}
                                        \author[1]{Alastair V Harry}
                                                            \affil{Fisheries \& Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia \thanks{\href{mailto:alastair.harry@gmail.com}{\nolinkurl{alastair.harry@gmail.com}}}}
                                                                                \author[2]{Ivy E Baremore}
                                                            \affil{MarAlliance, Roatán, Islas de la Bahía, Honduras}
                                                                                \author[3]{Andrew N Piercy}
                                                            \affil{College of Southern Maryland, La Plata MD 20646, United States}
                                            \date{}

\begin{document}
\maketitle

\setstretch{2}
\newpage

\textbf{Abstract}

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic `maternity' function with a variable upper asymptote, \(P_{Max}\), can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \textgreater{} 100, estimation of \(P_{Max}\) is generally preferable to using a fixed value.

\textbf{Key words} Logistic regression; reproductive biology; maturity ogive; fisheries management

\newpage

\section{Introduction}\label{introduction}

Many ecological and evolutionary applications need quantitative data on the reproductive capacity of a population and in fisheries science this information forms the basis of widely-used management reference points such as spawning stock biomass (Hilborn and Walters 1992, Quinn and Deriso 1999). Calculating reproductive capacity requires data on the sexually mature proportion of the population contributing to reproduction at a given size or age. For most populations of teleost fishes a simple two-parameter logistic regression function (2PLF) is sufficient to model this process using dichotomous sexual maturity stage data (immature / mature) from a sample of individuals (Jennings et al. 2001, King 2007). While the mathematical formulae and methods of statistically estimating parameters for a 2PLF vary, the estimated values are similar and can be obtained with the built in generalised linear model (GLM) programs in most modern statistical software.

In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one (Frisk et al. 2005, Dulvy et al. 2008, Rigby and Simpfendorfer 2015). Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing (Harry et al. 2013, Fujinami et al. 2017). This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity (Braccini et al. 2006, Walker 2007, Montealegre-Quijano et al. 2014). Finally, because female fecundity is low (usually \(\ll\) 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability (Taylor et al. 2013). This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

\subsection{Maternity function}\label{maternity-function}

Walker (2005) recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: \[Y_i \sim B(x_i; n_i, P_i)\] \[E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x_i)}}\] where the proportion of individuals in a given length or age category, \(x_i\), that are in maternal condition, \(Y_i\), is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, \(P_i\), multiplied by the total number of individuals in the category, \(n_i\). If \(x\) is a continuous rather than categorical variable (i.e.~\(n = 1\)), \(Y_i\) reduces to a Bernoulli random variable (Zuur et al. 2007). \(\alpha\) is the upper asymptote of the curve, controlling the maximum value of \(Y\) as \(x\) approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when \(\alpha = 1\), as is appropriate for most teleost and chondrichthyan populations if modeling maturity (Quinn and Deriso 1999, Roa et al. 1999). When \(\alpha = 1\), the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows \(Y\) to be modeled as a linear function of \(x_i\), where \(b_0\) and \(b_1\) are the intercept and slope, respectively (Zuur et al. 2007). Alternatively, the model can be expressed in more biologically relevant terms as: \[E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} \] where \(x_{50}\) and \(x_{95}\) are the sizes or ages at which 50\% and 95\% of the maximum proportion of individuals (\(P_{Max}\)) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

Walker (2005) defines individual females as being in maternal condition if they would have given birth or laid eggs by the end of a given year such that they contribute to annual recruitment (age 0+ cohort) at the beginning of the next year. Pregnancy is a necessary, but not sufficient, condition for maternity, and the two conditions are not always synonymous. For example, reproduction in the Australian population of school shark, \emph{Galeorhinus galeus}, occurs triennially, such that approximately one third of mature females give birth annually and \(P_{Max} \approx 1/3\) (Walker 2005). Gestation lasts 20 months, and includes a protracted period of ovulation. Newly pregnant females are not considered to be in maternal condition until the year of pre-recruitment. Walker's (2005) definition is specific to species that reproduce during a single, well-defined period of the year, but the concept can be generalised to accommodate asynchronously reproducing species or those that reproduce more than once a year.

Numerous authors have adopted Walker's (2005) maternity function, including for species with biennial or longer reproductive cycles (Huveneers et al. 2007, Rochowski et al. 2015, Trinnie et al. 2016) as well as those with annual or shorter reproductive cycles (Trinnie et al. 2009, Mejía‐Falla et al. 2012, Harry et al. 2013, Taylor et al. 2016). Techniques have been described for determining maternal output in asynchronously reproducing species (Braccini et al. 2006, Colonello et al. 2016), and maternity functions are also increasingly being used directly in shark and ray population assessments (SEDAR 2012, 2017, 2023).

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles. Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

For practitioners that are aware of maternity functions, lack of information on implementation may also have discouraged use. Walker's (2005) approach to estimating parameters in the maternity function is difficult to reproduce, apparently due to the constraints of the proprietary statistical program first used to implement the analysis. Specifically, the method as described involves adjusting the raw data prior to parameter estimation and then weighting it during analysis (Walker 2005). Additionally, the use of a fixed value for \(P_{Max}\), as in the original analysis, has the potential to bias parameter estimates and reduce their standard errors (Motulsky and Christopoulos 2004). Alternative approaches to estimating maternal parameters have also arisen in the literature, indicative of a lack of guidance on implementation. For example, some authors have obtained maternity parameters by fitting a 2PLF to maternity data (Baremore and Hale 2012, Baremore and Passerotti 2013).

A consequence of the low uptake in use of maternity functions is that most practitioners undertaking population assessments invariably take an \emph{ad hoc} approach to quantifying maternal reproductive output. This typically involves approximating maternal output using a maturity function. Under such an approach all mature females are tacitly assumed to reproduce in each breeding season (Cortes 1998). Non-annual reproduction is accounted for by weighting fecundity or the maturity function by the assumed (but often unknown) duration of the reproductive cycle (Walker 2005). In some cases, calculations of reproductive output may use the age-at-first-reproduction, defined as the mean age at maturity plus the gestation period (Mollet and Cailliet 2002). This accounts for the protracted gestation period of many chondrichthyans, but still assumes that all individuals begin reproducing immediately after maturity. Measures of annual reproductive output that follow this approach are hence built upon a series of assumptions that are rarely tested in practice.

\subsection{Revisiting the maternity function}\label{revisiting-the-maternity-function}

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions (Cortés et al. 2012). It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.~annual, biennial), despite observations of plasticity in this trait within discrete populations (Higgs et al. 2020). Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles. To date only a single study on spiny dogfish, \emph{Squalus acanthias}, has statistically estimated \(P_{Max}\) from data (Colonello et al. 2016), and the feasibility of doing so for a wider range of species has not been investigated.

This study revisits the use of maternity functions with the objectives of providing guidance on implementation and appropriate use. To address this objective a combination of simulated and empirical data were used to 1) evaluate the performance of two alternative methods for estimating maternity parameters, subject to varying reproductive frequency, sample size, and gear selectivity, 2) illustrate the effect of not using using maternity functions on calculations of lifetime reproductive output, and 3) outline strategies for making inferences on reproductive frequency from maternity data.

\section{Materials and methods}\label{materials-and-methods}

\subsection{Approach}\label{approach}

To assess the feasibility of estimating maternity parameters a simulation study was carried out using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1). Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs. The performance of three methods in estimating `true' maternity parameters from simulated data was tested by examining precision, bias, and interval coverage. To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters. Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency. All simulation modelling and data analysis was conducted using the R language {[}Version 4.2.0{]} (R Core Team 2022) and the Template Model Builder (TMB) R package {[}Version 1.9.1{]} (Kristensen et al. 2016).

\subsection{Simulation}\label{simulation}

\subsubsection{Data generation}\label{data-generation}

Data were generated for the gummy shark, \emph{Mustelus antarcticus}, and school shark, \emph{Galeorhinus galeus}, using parameters from studies conducted on southern Australian populations (Table 1) (Grant et al. 1979, Kirkwood and Walker 1986, Moulton et al. 1992, Walker 1992, Punt and Walker 1998, Walker 2005, 2007). Both species are well-studied and were chosen to be representative of fast and slow chondrichthyan life history styles, respectively (Stevens 1999).

Catch at length data were simulated using a female-only, length- and age-based equilibrium population model (Hesp 2023). The approach involved calculating expected survival and catches per recruit subject to specified life history parameters, gear selectivity, and total mortality. Growth was modelled using a von Bertalanffy growth function with length \emph{l}, of a fish of age, \emph{a}, calculated as \[l(a) = L_\infty (1- e^{-K(a-a_0)})\]

where \(L_\infty\) is asymptotic length, \emph{K} is the growth coefficient and \(a_0\) is the hypothetical age at zero length. Recruitment into the population was assumed to occur at age 0 with length conforming to a normal distribution. The proportion of fish of length, \emph{l}, at age, \(a = 0\), was calculated as \[\Theta_k = \int_{l^-_k}^{l^+_k}f_{a=0}(l)dl\] where \(l^-\) and \(l^+\) are the lower and upper limits of each 1cm length class, \emph{k}. \(f_{a=0}(l)\) is the value of the normal probability density function at age \(a = 0\) for a given length, \emph{l},\[f_{a=0}(l) = \frac{1}{\sigma\sqrt{2{\pi}}}\exp\bigg[\frac{(l-\mu)^2}{2\sigma^2}\bigg]
\]

where mean length, \(\mu\), was calculated from the growth curve. The standard deviation, \(\sigma\) was not available for either species so was assumed to be directly proportional to length, with a constant of proportionality of 5\% i.e.~\(\sigma = 0.05 \mu\) (Table 1).

Growth in subsequent age classes was modelled using a length transition matrix (Punt et al. 1997, Hall et al. 2000), \(G = g_{j,k}\) that represents the probability that a fish in length class, \(j\), will grow into length class, \(k\), over a specified time interval \[ G = \begin{bmatrix}
 g_{1,1}&0& 0 &0\\ 
 g_{2,1}& g_{2,2} & 0  &0 \\ 
 \cdots&\cdots  & \ddots  & 0\\ 
 g_{n,1}& g_{n,2} & \cdots & g_{n,n}
\end{bmatrix}\]

where the elements of the matrix followed the general form of Punt et. al. (1997) and are given by\[ g_{k,j} = \left\{
    \begin{array}{ll}
        \int_{{\infty}^-}^{\mathrm{l_{k}^+}}f[\phi(l,j)]dl & \mbox{if }k=1 \\
        \int_{l_{k}^-}^{\mathrm{l_{k}^+}}f[\phi(l,j)]dl &\mbox{if }k=1<k<n \\
\int_{l_{k}^-}^{\mathrm{\infty^+}}f[\phi(l,j)]dl &\mbox{if }k=n \\
    \end{array}
\right.
\]

where \emph{l} is mean length, \emph{f} is the specified normal distribution, \(\mathrm{l_{k}^-}\) and \(\mathrm{l_{k}^+}\) are the lower and upper limits of length class \emph{j}, and \(\phi\) is a vector of parameters.

Selectivity in the model was assumed to follow a gamma function based on gillnet selectivity experiments (Kirkwood and Walker 1986, Punt and Walker 1998), where the selectivity of an individual in length class, \(l\) was given by \[S(l) = \left [ \frac{l}{\theta_1 \theta_2} \right ]^{\theta_1 }exp(\theta_1-\frac{l}{\theta_2}) \] where \(\theta_1\) and \(\theta_2\) are estimated parameters (Kirkwood and Walker 1986). Two selectivity scenarios were simulated corresponding to low and high selectivity of the maternal component of the population. For consistency between species, gillnet mesh sizes were chosen that would result in peak relative selectivity occurring at the lengths equal to 25\% and 75\% of the population in maternal condition (Table 1).

Fishing mortality, \emph{F}, in length class \(l\), was calculated as \[F(l) = S(l) \cdot F\]where \(F\) was assumed to be constant and nominally set to 25\% of natural mortality, \(M\) (i.e.~\(0.25M\)). The effects of varying levels of \(F\) were not investigated further in this study. Total mortality, \emph{Z,} in length class, \(l\), was further given by \[Z(l) = F(l)\cdot{M} \]

The number of fish, \(N\), per recruit in length class, \(l\), that survived to age, \(a\), was calculated as

\[ N_{l,a} = \left\{    \begin{array}{ll}        1 & \mbox{if }a=0 \\        N_{l,a-1}\exp(-Z_{l}) &\mbox{if }a=0<a<A_{Max} \\        N_{l,a-1}\exp(-Z_{l})/(1-\exp(-Z_{l})) &\mbox{if }a=A_{Max} \\    \end{array}\right.\]

where \(A_{Max}\) is the maximum age (Table 1). The estimated numbers of fish caught in length class, \(l\), at age, \(a\), was calculated using the Baranov catch equation as

\[
\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
\]

To examine the effect of varying sample size, six scenarios were conducted with catch ranging from 50 to 2500 individuals, reflecting a gradient from data poor to rich.

Next, dichotomous maturity-at-length data, \(\psi(l)\), were randomly generated by simulating from a Bernoulli distribution, \(\Psi(l) \sim B(P)\), where \(P\) was given by the 2PLF: \[P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}\] This process was then repeated to generate maternity-at-length data, \(\Psi'(l)\), using the 3PLF: \[\Psi' \sim B(P)\] \[P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}\] Finally, 300 iterations were run for each of the 144 unique variables combinations.

\subsubsection{Estimation approaches}\label{estimation-approaches}

Two approaches to estimating maternity parameters were compared: the 3PLF itself (3PLF-estimated) and the 3PLF function with a fixed asymptote (3PLF-fixed). Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF-maturity). The 3PLF-estimated method was undertaken to validate the utility of this model, which has so far been used in only a single study (Colonello et al. 2016). The 3PLF-fixed is the method described by Walker (2005) and most commonly used in practice. The 2PLF-maternity method was used to examine what effect simply ignoring the upper asmyptote had (i.e.~\(P_{Max}\) fixed at 1). The 2PLF-maturity method is commonly used to approximate a maternity curve, and tacitly assumed to be similar. The resulting maturity curve can then be weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially (SEDAR 2023)).

Only the 3PLF-estimated method involves statistically estimating \(P_{Max}\), which must be subjectively chosen in the case of the 3PLF-fixed and the 2PLF-maturity. \(P_{Max}\) is ideally chosen based on detailed study of the ovarian and uterine cycles (Walker 2005), or alternatively based on the proportion of mature females observed to be in maternal condition during sampling (Baremore and Hale 2012, Harry et al. 2013, Trinnie et al. 2016). For the purposes of the simulation, the timing and duration of the uterine and ovarian cycles were assumed to be unknown. The following procedure was used to `guess' the fixed value of \(P_{Max}\) in each simulation. \(P_{Max}\) was chosen as the proportion of females in maternal condition above a the length at which \(99\%\) of females were mature, \(L > \Psi(L_{99\%})\). In some cases this procedure failed due to there being no maternal females in the simulated data set meeting this criteria. If this occurred, \(L_{95}\) was used, followed by \(L_{50}\), and finally the proportion of all mature females in the sample that were in maternal condition. Statistical estimation of all non-fixed parameters in each of the methods was undertaken using maximum likelihood. Nonparametric bootstrapping was used to calculate approximate 50\% confidence intervals for the best-fit parameters from 250 resampled data sets.

\subsubsection{Reproductive output}\label{reproductive-output}

For each iteration, the per-generation rate of multiplication, \(R_0\), or lifetime female pup production, was calculated as: \[R_0 =\int_{0}^{\infty}R\beta(a)e^{\int_{0}^{a}M ds}da \] where \(R\) and \(M\) are the embryonic sex ratio and natural mortality rate (Table 1) and \(\beta(a)\) is the age-specific reproductive rate which incorporates \(\Psi'(L)\) (Xiao and Walker 2000).

\subsubsection{Performance}\label{performance}

Model performance was evaluated in each iteration where 1) the fitted model successfully converged as indicated by a positive definite Hessian matrix, and 2) the estimated parameters \(L_{50}'\) and \(L_{95}'\) remained within specified bounds (\(0 - 10^4\)). An iteration was also not attempted if there were no maternal females in the randomly generated sample. Model performance was evaluated by quantifying precision, bias, and interval coverage for \(L'_{50}\), \(P_{Max}\) and \(R_0\). Bias was quantified using the mean relative error, \(\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]\) and precision using the mean absolute relative error, \(\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|\). Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50\% confidence intervals (Rudd and Thorson 2018). For a well-performing model, approximately 50\% of parameters would be expected to fall withing the nominal 50\% confidence intervals.

\subsubsection{Empirical case study}\label{empirical-case-study}

An empirical case study was also undertaken to show how maternity functions can be used in practice. Data from two studies of the reproductive biology of the sandbar shark, \emph{Carcharhinus plumbeus}, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods. The sandbar shark is a relatively long-lived (\(A_{Max}\) = 33 years) and slow growing species (\(K = 0.12 \text{ yr}^{-1}\)) with a protracted reproductive cycle lasting longer than a year (Springer 1960). The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s (Sminkey and Musick 1995) and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes (Baremore and Hale 2012, Piercy et al. 2016). In their study, Baremore and Hale (2012) estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle. Piercy \emph{et al.} (2016) did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles. To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with \(P_{Max}\) estimated and fixed at 0.5 and 0.33, respectively.

\section{Results}\label{results}

\subsection{Simulation study}\label{simulation-study}

Parameter estimation was generally straightforward; with sample sizes \(> 50\) a success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1. In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.

Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased. The nature of this problem can be seen in the positive correlation between the \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) parameters whereby \(P_{Max}\) affects the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\) (Figure 2).

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved precision with the 3PLF methods, although at larger sample sizes better precision was achieved by estimating \(P_{Max}\) (Figure 3). Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for \(\hat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size. In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\).

Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9). As \(P_{Max}\) decreased the proportion of maternal females in the analysis also decreased.

Overall, the 2PLF-maternity method performed the poorest. Ignoring \(P_{Max}\) introduced a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3). This manifested in an overestimation of \(L'_{50}\) that was exacerbated by gear selectivity effects (Figure S2). In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance. \(L_{50}\) could usually be estimated with a higher precision than \(L'_{50}\). For the gummy shark, where \(L_{50}\) and \(L'_{50}\) were relatively similar, using the 2PLF-maturity method led to a constant underestimate of \(L'_{50}\) of approximately 2\%, exceeding the performance of the 3PLF-methods at most sample sizes.

The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios. For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%. (Table S1, Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10). For both species, better precision in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).

\subsection{Empirical case study}\label{empirical-case-study-1}

Maternal data were re-analysed for 1087 sandbar sharks including 0 mature individuals of which \ensuremath{\infty{}}\% were in maternal condition. Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2). Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence. For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\). The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86). Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data (Burnham and Anderson 2002).

\section{Discussion}\label{discussion}

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, \(P_{Max}\). Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that precision, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated. Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

\subsection{Implementing maternity functions}\label{implementing-maternity-functions}

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\) from data, but also identified situations where it may be preferable to fix \(P_{Max}\). In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method. At sample sizes below this it is therefore advisable to fix \(P_{Max}\). Larger sample sizes are desirable before attempting to estimate \(P_{Max}\) in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as \(P_{Max}\) decreases.

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of \(\geq500\) would be needed to estimate \(P_{Max}\) with \textless10\% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al. 1999). In the case of the gummy shark this imprecision led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\). Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000. A similar sample size was used by Colonello \emph{et al}. (2016) to successfully estimate \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al. 2010, Tribuzio and Kruse 2012).

In light of the data requirements suggested by this study, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close \(L_{50}\) is to \(L'_{50}\). While several studies have shown these parameters to be similar (Walker 2007, Soto-López et al. 2018), \(L'_{50}\) is more frequently shifted to the right of \(L_{50}\) (Braccini et al. 2006, Montealegre-Quijano et al. 2014, Colonello et al. 2016, Palacios-Hernández et al. 2020), and this length difference can equate to one or more years. For example, Harry \emph{et al.} (2013) found both spot-tail shark, \emph{C. sorrah}, and Australian blacktip shark, \emph{C. tilstoni}, began reproducing the year after reaching sexual maturity. Similarly, Fujiyama found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.

More research is needed to understand the relationship between size at maturity and maternity and variability within and among taxa. In general, this study confirms Walker's (2005) assertion that the common practice of weighting the maturity curve by the frequency of parturition overestimates recruitment. This study also confirms that fitting a 2PLF to maternal data when \(P_{Max} < 1\) (effectively ignoring \(P_{Max}\)) is likely to result in biased parameters and can also overestimate recruitment. While not widely used, this approach has been undertaken in several studies (Baremore and Hale 2012, Mejía‐Falla et al. 2012, Baremore and Passerotti 2013, Rambahiniarison et al. 2018).

The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for \(P_{Max}\). In this study, the value for \(P_{Max}\) used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data. The potential to introduce a greater level of bias in the analysis through the incorrect selection of \(P_{Max}\) should also be considered if using this approach.

\subsection{Empirical study}\label{empirical-study}

The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\). Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy \emph{et al.} (2016) were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level. Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\hat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012). Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013). Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

\subsection{Advantages and future directions}\label{advantages-and-future-directions}

Despite the simplicity of the 3PLF-estimated method, it nonetheless represents a conceptual shift for chondrichthyan reproductive and population biology. Until now reproductive periodicity has largely been determined qualitatively from observations of female reproductive biology or inferred based on circumstantial evidence. Using the 3PLF-estimated method to estimate \(P_{Max}\) within a maternity function transforms it from a nuisance parameter to one of direct inferential interest that can be seen as a valuable output of the modelling process itself. In addition to the potential benefits of using this method to improve estimation of maternal parameters, there are also a range of other advantages that arise from estimating \(P_{Max}\).

One benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to \(P_{Max}\). Such assumptions can have important implications for population modelling. For example, to account for uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, \emph{Urolophus paucimaculatus}, Trinnie \emph{et al}. (Trinnie et al. 2014) estimated maternity parameters for ten plausible scenarios. In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle (SEDAR 2017).

More commonly, practitioners have chosen fixed values of the form \(P_{Max} = 1/n\), where \emph{n} is the apparent duration of the reproductive cycle in years. This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently that the larger population. Long-term study of offspring from genetically profiled lemon sharks, \emph{Negaprion brevirostris}, confirmed that most individuals reproduced biennially, but also revealed some cases of triennial reproduction (Feldheim et al. 2014). Statistical estimation of \(P_{Max}\) from data avoids these decisions and enables uncertainty in this parameter to be included in subsequent population models.

Estimation of \(P_{Max}\) using the 3PLF-estimated method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans. White sharks, \emph{Carcharodon carcharias}, and whale shark, \emph{Rhinchodon typus}, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist (Joung et al. 1996, Sato et al. 2016). In both species lack of data on reproductive frequency is a major impediment to population modelling and the development of management strategies (Huveneers et al. 2018, Bowlby and Gibson 2020). With many of these populations the focus of ongoing monitoring, non-lethal methods for assessing maternal state may eventually provide a way of collecting sufficient data for estimation of \(P_{Max}\) (Sulikowski et al. 2016).

A logical progression from estimating \(P_{Max}\) is the consideration of alternative functional forms for this parameter such as time- or space-varying \(P_{Max}\) or more complex length- or age-dependent forms (Winton et al. 2014). Despite the diverse range of reproductive modes found in chondrichthyan fishes, the current, limited knowledge of maternal investment has typically investigated variables such as litter size and pup condition (Hussey et al. 2010). Little is known about how reproductive frequency varies as a function of size or age, in response to environmental effects, or at changing population densities. Better understanding of these factors has the potential to improve understanding reproductive strategies and is, in many cases, of direct relevance to management.

\section{Acknowledgements}\label{acknowledgements}

We thank John Carlson and all NMFS staff who were involved in collecting the sandbar shark reproductive data used in this study. We also thank Alex Hesp for valuable comments on an early draft of this paper and generosity in sharing R code.

\section{Competing interests}\label{competing-interests}

The authors declare there are no competing interests.

\section{Data availability}\label{data-availability}

Data generated or analyzed during this study are available from the corresponding author upon reasonable request.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-baremore_reproduction_2012}
Baremore, I.E., and Hale, L.F. 2012. Reproduction of the {Sandbar} {Shark} in the {Western} {North} {Atlantic} {Ocean} and {Gulf} of {Mexico}. Marine and Coastal Fisheries \textbf{4}(1): 560--572. doi: \href{https://doi.org/10.1080/19425120.2012.700904}{10.1080/19425120.2012.700904}.

\bibitem[\citeproctext]{ref-baremore_reproduction_2013}
Baremore, I.E., and Passerotti, M.S. 2013. Reproduction of the {Blacktip} {Shark} in the {Gulf} of {Mexico}. Marine and Coastal Fisheries \textbf{5}(1): 127--138. doi: \href{https://doi.org/10.1080/19425120.2012.758204}{10.1080/19425120.2012.758204}.

\bibitem[\citeproctext]{ref-bowlby_implications_2020}
Bowlby, H.D., and Gibson, A.J.F. 2020. Implications of life history uncertainty when evaluating status in the {Northwest} {Atlantic} population of white shark (\emph{{Carcharodon} carcharias)}. Ecology and Evolution \textbf{10}(11): 4990--5000. doi: \href{https://doi.org/10.1002/ece3.6252}{10.1002/ece3.6252}.

\bibitem[\citeproctext]{ref-braccini_determining_2006}
Braccini, J.M., Gillanders, B.M., and Walker, T.I. 2006. Determining reproductive parameters for population assessments of chondrichthyan species with asynchronous ovulation and parturition: Piked spurdog (\emph{{Squalus} megalops}) as a case study. Mar. Freshwater Res. \textbf{57}: 105--119. doi: \href{https://doi.org/doi:\%2010.1071/Mf05076}{doi: 10.1071/Mf05076}.

\bibitem[\citeproctext]{ref-burnham_model_2002}
Burnham, K.P., and Anderson, D.R. 2002. Model selection and multimodel inference: A practical information-theoretic approach. 2nd edn. Springer, New York, NY.

\bibitem[\citeproctext]{ref-colonello_reproductive_2016}
Colonello, J.H., Cortés, F., Belleggia, M., and Massa, A.M. 2016. Reproductive and population parameters of spiny dogfish \emph{{Squalus} acanthias} in the south-western {Atlantic} {Ocean}. J. Fish Biol. \textbf{88}(5): 1758--1775. doi: \href{https://doi.org/10.1111/jfb.12940}{10.1111/jfb.12940}.

\bibitem[\citeproctext]{ref-cortes_demographic_1998}
Cortes, E. 1998. Demographic analysis as an aid in shark stock assessment and management. Fish. Res. \textbf{39}: 199--208. doi: \href{https://doi.org/10.1016/S0165-7836(98)00183-0}{10.1016/S0165-7836(98)00183-0}.

\bibitem[\citeproctext]{ref-cortes_population_2012}
Cortés, E., Brooks, E.N., and Gedamke, T. 2012. Population dynamics, demography, and stock assessment. \emph{In} Biology of {Sharks} and {Their} {Relatives}, {Second} {Edition}. \emph{Edited by} J.C. Carrier, J.A. Musick, and M.R. Heithaus. CRC Press, New York. pp. 453--86.

\bibitem[\citeproctext]{ref-dulvy_you_2008}
Dulvy, N.K., Baum, J.K., Clarke, S., Compagno, L.J.V., Cortes, E., Domingo, A., Fordham, S., Fowler, S., Francis, M.P., Gibson, C., Martínez, J., Musick, J.A., Soldo, A., Stevens, J.D., and Valenti, S. 2008. You can swim but you can't hide: The global status and conservation of oceanic pelagic sharks and rays. Aquatic Conservation: Marine and Freshwater Ecosystems.

\bibitem[\citeproctext]{ref-feldheim_two_2014}
Feldheim, K.A., Gruber, S.H., DiBattista, J.D., Babcock, E.A., Kessel, S.T., Hendry, A.P., Pikitch, E.K., Ashley, M.V., and Chapman, D.D. 2014. Two decades of genetic profiling yields first evidence of natal philopatry and long-term fidelity to parturition sites in sharks. Mol. Ecol. \textbf{23}(1): 110--117. doi: \href{https://doi.org/10.1111/mec.12583}{10.1111/mec.12583}.

\bibitem[\citeproctext]{ref-frisk_life_2005}
Frisk, M.G., Miller, T.J., and Dulvy, N.K. 2005. Life histories and vulnerability to exploitation of elasmobranchs: {Inferences} from elasticity, pertubation and phylogenetic analyses. Journal of Northwest Atlantic Fishery Science \textbf{35}: 27--45. doi: \href{https://doi.org/10.2960/J.v35.m514}{10.2960/J.v35.m514}.

\bibitem[\citeproctext]{ref-fujinami_reproductive_2017}
Fujinami, Y., Semba, Y., Okamoto, H., Ohshimo, S., and Tanaka, S. 2017. Reproductive biology of the blue shark (\emph{{Prionace} glauca}) in the western {North} {Pacific} {Ocean}. Mar. Freshwater Res. \textbf{68}(11): 2018--2027. Available from \url{https://doi.org/10.1071/MF16101}.

\bibitem[\citeproctext]{ref-grant_estimation_1979}
Grant, C., Sandland, R., and Olsen, A. 1979. Estimation of growth, mortality and yeild per recruit of the {Australian} school shark, \emph{{Galeorhinus} australis} ({Macleay}), from tag recoveries. Mar. Freshwater Res. \textbf{30}(5): 625--637. doi: \href{https://doi.org/10.1071/MF9790625}{10.1071/MF9790625}.

\bibitem[\citeproctext]{ref-hale_age_2013}
Hale, L.F., and Baremore, I.E. 2013. Age and {Growth} of the {Sandbar} {Shark} (\emph{{Carcharhinus} plumbeus}) from the {Northern} {Gulf} of {Mexico} and the {Western} {North} {Atlantic} {Ocean}. Gulf of Mexico Science \textbf{1}(2): 28--39.

\bibitem[\citeproctext]{ref-hall_modelling_2000}
Hall, N., Cao, L., Chubb, C., Caputi, N., Cheng, H., Melville-Smith, R., and Shanks, S. 2000. Modelling to explore management strategies to optimise the value of the {Rock} {Lobster} {Fishery} of {Western} {Australia}. Fisheries Research Development Corporation Final Report Project (97/104).

\bibitem[\citeproctext]{ref-harry_age_2013}
Harry, A.V., Tobin, A.J., and Simpfendorfer, C.A. 2013. Age, growth and reproductive biology of the spot-tail shark,\emph{{Carcharhinus} sorrah}, and the {Australian} blacktip shark, \emph{{Carcharhinus} tilstoni}, from the {Great} {Barrier} {Reef} {World} {Heritage} {Area}, north-eastern {Australia}. Mar. Freshwater Res. \textbf{64}: 277--293. doi: \href{https://doi.org/10.1071/MF12142}{10.1071/MF12142}.

\bibitem[\citeproctext]{ref-hesp_l3assess_2023}
Hesp, A. 2023. {L3Assess}: {Catch} curve and per recruit analyses. R package version 0.1.0.

\bibitem[\citeproctext]{ref-higgs_reproductive_2020}
Higgs, J.M., Hoffmayer, E.R., Sulikowski, J.A., Driggers, W.B., Stiller, D.A., and Hendon, J.M. 2020. Reproductive biology of the finetooth shark (\emph{{Carcharhinus} isodon}) in the northern {Gulf} of {Mexico}, with evidence of both annual and biennial reproduction. Mar. Freshwater Res. Available from \url{https://doi.org/10.1071/MF20120}.

\bibitem[\citeproctext]{ref-hilborn1992}
Hilborn, R., and Walters, C.J. 1992. Quantitative fisheries stock assessment. Choice, dynamics and uncertainty. Springer, New York, NY.

\bibitem[\citeproctext]{ref-hussey_maternal_2010}
Hussey, N.E., Wintner, S.P., Dudley, S.F.J., Cliff, G., Cocks, D.T., and MacNeil, M.A. 2010. Maternal investment and size-specific reproductive output in carcharhinid sharks. J. Anim. Ecol. \textbf{79}: 184--193. doi: \href{https://doi.org/doi:\%2010.1111/j.1365-2656.2009.01623.x}{doi: 10.1111/j.1365-2656.2009.01623.x}.

\bibitem[\citeproctext]{ref-huveneers_future_2018}
Huveneers, C., Apps, K., Becerril-García, E.E., Bruce, B., Butcher, P.A., Carlisle, A.B., Chapple, T.K., Christiansen, H.M., Cliff, G., Curtis, T.H., Daly-Engel, T.S., Dewar, H., Dicken, M.L., Domeier, M.L., Duffy, C.A.J., Ford, R., Francis, M.P., French, G.C.A., Galván-Magaña, F., García-Rodríguez, E., Gennari, E., Graham, B., Hayden, B., Hoyos-Padilla, E.M., Hussey, N.E., Jewell, O.J.D., Jorgensen, S.J., Kock, A.A., Lowe, C.G., Lyons, K., Meyer, L., Oelofse, G., Oñate-González, E.C., Oosthuizen, H., O'Sullivan, J.B., Ramm, K., Skomal, G., Sloan, S., Smale, M.J., Sosa-Nishizaki, O., Sperone, E., Tamburin, E., Towner, A.V., Wcisel, M.A., Weng, K.C., and Werry, J.M. 2018. Future research directions on the {``{Elusive}''} white shark. Frontiers in Marine Science \textbf{5}: 455. doi: \href{https://doi.org/10.3389/fmars.2018.00455}{10.3389/fmars.2018.00455}.

\bibitem[\citeproctext]{ref-huveneers_reproductive_2007}
Huveneers, C., Walker, T.I., Otway, N.M., and Harcourt, R.G. 2007. Reproductive synchrony of three sympatric species of wobbegong shark (genus \emph{orectolobus}) in {New} {South} {Wales}, {Australia}: Reproductive parameter estimates necessary for population modelling. Mar. Freshwater Res. \textbf{58}: 765--777. doi: \href{https://doi.org/10.1071/MF06187}{10.1071/MF06187}.

\bibitem[\citeproctext]{ref-jennings_marine_2001}
Jennings, S., Kaiser, M., and Reynolds, J.D. 2001. Marine fisheries ecology. John Wiley \& Sons, Oxford.

\bibitem[\citeproctext]{ref-joung_whale_1996}
Joung, S.-J., Chen, C.-T., Clark, E., Uchida, S., and Huang, W.Y.P. 1996. The whale shark, \emph{{Rhincodon} typus}, is a livebearer: 300 embryos found in one {``megamamma''} supreme. Environ. Biol. Fish. \textbf{46}(3): 219--223. doi: \href{https://doi.org/10.1007/BF00004997}{10.1007/BF00004997}.

\bibitem[\citeproctext]{ref-king_fisheries_2007}
King, M. 2007. Fisheries biology, assessment and management. Blackwell Publishing, Oxford.

\bibitem[\citeproctext]{ref-kirkwood_gill_1986}
Kirkwood, G.P., and Walker, T.I. 1986. Gill net mesh selectivities for gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, taken in southeastern {Australian} waters. Aust. J. Mar. Fresh. Res. \textbf{37}: 689--697.

\bibitem[\citeproctext]{ref-kristensen_tmb:_2016}
Kristensen, K., Nielsen, A., Berg, C.W., Skaug, H., and Bell, B.M. 2016. {TMB}: {Automatic} {Differentiation} and {Laplace} {Approximation}. J. Stat. Softw. \textbf{70}(5): 1--21. doi: \href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.

\bibitem[\citeproctext]{ref-mejiafalla_reproductive_2012}
Mejía‐Falla, P.A., Navia, A.F., and Cortés, E. 2012. Reproductive variables of \emph{{Urotrygon} rogersi} ({Batoidea}: {Urotrygonidae}): A species with a triannual reproductive cycle in the eastern tropical {Pacific} {Ocean}. J. Fish Biol. \textbf{80}(5): 1246--1266. doi: \href{https://doi.org/10.1111/j.1095-8649.2012.03237.x}{10.1111/j.1095-8649.2012.03237.x}.

\bibitem[\citeproctext]{ref-mollet_comparative_2002}
Mollet, H.F., and Cailliet, G.M. 2002. Comparative population demography of elasmobranchs using life history tables, {Leslie} matrices and stage-based matrix models. Mar. Freshwater Res. \textbf{53}: 503--516.

\bibitem[\citeproctext]{ref-montealegre-quijano_sexual_2014}
Montealegre-Quijano, S., Cardoso, A.T.C., Silva, R.Z., Kinas, P.G., and Vooren, C.M. 2014. Sexual development, size at maturity, size at maternity and fecundity of the blue shark \emph{{Prionace} glauca} ({Linnaeus}, 1758) in the {Southwest} {Atlantic}. Fish. Res. \textbf{160}: 18--32. doi: \url{https://doi.org/10.1016/j.fishres.2014.03.003}.

\bibitem[\citeproctext]{ref-motulsky_fitting_2004}
Motulsky, H., and Christopoulos, A. 2004. Fitting models to biological data using linear and nonlinear regression: A practical guide to curve fitting. Oxford University Press, San Diego.

\bibitem[\citeproctext]{ref-moulton_age_1992}
Moulton, P.L., Walker, T.I., and Saddlier, S.R. 1992. Age and growth-studies of gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, and school shark, \emph{{Galeorhinus} galeus}({Linnaeus}), from {Southern} {Australian} waters. Aust. J. Mar. Fresh. Res. \textbf{43}: 1241--1267. doi: \href{https://doi.org/10.1071/MF9921241}{10.1071/MF9921241}.

\bibitem[\citeproctext]{ref-oddone_size_2010}
Oddone, M.C., Paesch, L., and Norbis, W. 2010. Size structure, abundance and preliminary information on the reproductive parameters of the shortspine spurdog (\emph{{Squalus} mitsukurii}) in the argentinean-uruguayan common fishing zone from the mid-1990s. J. Northw. Atl. Fish. Sci. \textbf{43}: 13--26. doi: \href{https://doi.org/10.2960/j.v.43.m662}{10.2960/j.v.43.m662}.

\bibitem[\citeproctext]{ref-palacios-hernandez_temporal_2020}
Palacios-Hernández, D., Castillo-Géniz, J.L., Méndez-Loeza, I., and Pérez-Jiménez, J.C. 2020. Temporal and latitudinal comparisons of reproductive parameters in a heavily exploited shark, the bonnethead, \emph{{Sphyrna} tiburo} ({L}. 1758), in the southern {Gulf} of {Mexico}. J. Fish Biol. \textbf{97}(1): 100--112. doi: \href{https://doi.org/10.1111/jfb.14330}{10.1111/jfb.14330}.

\bibitem[\citeproctext]{ref-piercy_histological_2016}
Piercy, A.N., Murie, D.J., and Gelsleichter, J.J. 2016. Histological and morphological aspects of reproduction in the sandbar shark \emph{{Carcharhinus} plumbeus} in the {U}.{S}. South-eastern {Atlantic} {Ocean} and {Gulf} of {Mexico}. J. Fish Biol. \textbf{88}(5): 1708--1730. doi: \href{https://doi.org/10.1111/jfb.12945}{10.1111/jfb.12945}.

\bibitem[\citeproctext]{ref-punt_estimating_1997}
Punt, A.E., Kennedy, R.B., and Frusher, S.D. 1997. Estimating the size-transition matrix for {Tasmanian} rock lobster, \emph{{Jasus} edwardsii}. Mar. Freshwater Res. \textbf{48}(8): 981--992. Available from \url{https://doi.org/10.1071/MF97017}.

\bibitem[\citeproctext]{ref-punt_stock_1998}
Punt, A.E., and Walker, T.I. 1998. Stock assessment and risk analysis for the school shark (\emph{{Galeorhinus} galeus)} off southern {Australia}. Mar. Freshwater Res. \textbf{49}: 719--731. doi: \href{https://doi.org/10.1071/MF96101}{10.1071/MF96101}.

\bibitem[\citeproctext]{ref-quinn_quantitative_1999}
Quinn, I.I., and Deriso, R.B. 1999. Quantitative fish dynamics. Oxford University Press, New York, NY.

\bibitem[\citeproctext]{ref-r_core_team_r:_2022}
R Core Team. 2022. R: {A} {Language} and {Environment} for {Statistical} {Computing}. Vienna, Austria. Available from \url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-rambahiniarison_life_2018}
Rambahiniarison, J.M., Lamoste, M.J., Rohner, C.A., Murray, R., Snow, S., Labaja, J., Araujo, G., and Ponzo, A. 2018. Life history, growth, and reproductive biology of four mobulid species in the {Bohol} {Sea}, {Philippines}. Frontiers in Marine Science \textbf{5}: 269. doi: \href{https://doi.org/10.3389/fmars.2018.00269}{10.3389/fmars.2018.00269}.

\bibitem[\citeproctext]{ref-rigby_patterns_2015}
Rigby, C., and Simpfendorfer, C.A. 2015. Patterns in life history traits of deep-water chondrichthyans. Deep Sea Research Part II: Topical Studies in Oceanography \textbf{115}: 30--40. doi: \url{https://doi.org/10.1016/j.dsr2.2013.09.004}.

\bibitem[\citeproctext]{ref-roa_estimation_1999}
Roa, R., Ernst, B., and Tapia, F. 1999. Estimation of size at sexual maturity: An evaluation of analytical and resampling procedures. Fish. B-noaa. \textbf{97}: 570--580.

\bibitem[\citeproctext]{ref-rochowski_reproductive_2015}
Rochowski, B.E.A., Graham, K.J., Day, R.W., and Walker, T.I. 2015. Reproductive biology of the greeneye spurdog \emph{{Squalus} chloroculus} ({Squaliformes}, {Squalidae}). J. Fish Biol. \textbf{86}(2): 734--754. doi: \href{https://doi.org/10.1111/jfb.12593}{10.1111/jfb.12593}.

\bibitem[\citeproctext]{ref-rudd_accounting_2018}
Rudd, M.B., and Thorson, J.T. 2018. Accounting for variable recruitment and fishing mortality in length-based stock assessments for data-limited fisheries. Can. J. Fish. Aquat. Sci. \textbf{75}(7): 1019--1035.

\bibitem[\citeproctext]{ref-sato_how_2016}
Sato, K., Nakamura, M., Tomita, T., Toda, M., Miyamoto, K., and Nozu, R. 2016. How great white sharks nourish their embryos to a large size: Evidence of lipid histotrophy in lamnoid shark reproduction. Biology Open \textbf{5}(9): 1211--1215. doi: \href{https://doi.org/10.1242/bio.017939}{10.1242/bio.017939}.

\bibitem[\citeproctext]{ref-sedar_sedar_2012}
SEDAR. 2012. {SEDAR} 29 stock assessment report: {HMS} {Gulf} of {Mexico} blacktip shark. North Charleston, SC.

\bibitem[\citeproctext]{ref-sedar_sedar_2017}
SEDAR. 2017. {SEDAR} 54 stock assessment report: {HMS} sandbar shark. North Charleston, SC.

\bibitem[\citeproctext]{ref-sedar_sedar_2023}
SEDAR. 2023. {HMS} {Hammerhead} {Sharks}: {Scalloped} {Hammerhead} {Shark}. {Section} {III}: {Assessment} {Report}. North Charleston, SC.

\bibitem[\citeproctext]{ref-sminkey_age_1995}
Sminkey, T.R., and Musick, J.A. 1995. Age and growth of the sandbar shark, \emph{{Carcharhinus} plumbeus}, before and after population depletion. Copeia: 871--883.

\bibitem[\citeproctext]{ref-soto-lopez_reproductive_2018}
Soto-López, K., Ochoa-Báez, R.I., Tovar-Ávila, J., and Galván-Magaña, F. 2018. Reproductive biology of the brown smooth-hound shark, \emph{{Mustelus} henlei} ({Chondrichthyes}: {Triakidae}), off northwestern {Mexico} based on macroscopic and histological analyses. Cienc. Mar. \textbf{44}(2): 125--139.

\bibitem[\citeproctext]{ref-springer_natural_1960}
Springer, S. 1960. Natural history of the sandbar shark, \emph{{Eulamia} milberti}. US Fish Wildl. Serv. Fish. Bull. \textbf{61}: 1--38.

\bibitem[\citeproctext]{ref-stevens_variable_1999}
Stevens, J.D. 1999. Variable resilience to fishing pressure in two sharks: The significance of different ecological and life history parameters. Am. Fish. S. S. \textbf{23}: 11--15.

\bibitem[\citeproctext]{ref-sulikowski_seasonal_2016}
Sulikowski, J.A., Wheeler, C.R., Gallagher, A.J., Prohaska, B.K., Langan, J.A., and Hammerschlag, N. 2016. Seasonal and life-stage variation in the reproductive ecology of a marine apex predator, the tiger shark \emph{{Galeocerdo} cuvier}, at a protected female-dominated site. Aquat. Biol. \textbf{24}(3): 175--184.

\bibitem[\citeproctext]{ref-taylor_stockrecruitment_2013}
Taylor, I.G., Gertseva, V., Methot, R.D., and Maunder, M.N. 2013. A stock--recruitment relationship based on pre-recruit survival, illustrated with application to spiny dogfish shark. Fish. Res. \textbf{142}: 15--21. doi: \url{https://doi.org/10.1016/j.fishres.2012.04.018}.

\bibitem[\citeproctext]{ref-taylor_living_2016}
Taylor, S.M., Harry, A.V., and Bennett, M.B. 2016. Living on the edge: Latitudinal variations in the reproductive biology of two coastal species of sharks. J. Fish Biol. \textbf{89}(5): 2399--2418. doi: \href{https://doi.org/10.1111/jfb.13126}{10.1111/jfb.13126}.

\bibitem[\citeproctext]{ref-tribuzio_life_2012}
Tribuzio, C.A., and Kruse, G.H. 2012. Life history characteristics of a lightly exploited stock of \emph{{Squalus} suckleyi}. J. Fish Biol. \textbf{80}(5): 1159--1180. doi: \href{https://doi.org/10.1111/j.1095-8649.2012.03241.x}{10.1111/j.1095-8649.2012.03241.x}.

\bibitem[\citeproctext]{ref-trinnie_reproductive_2009}
Trinnie, F.I., Walker, T.I., Jones, P.L., and Laurenson, L.J. 2009. Reproductive biology of the eastern shovelnose stingaree \emph{{Trygonoptera} imitata} from south-eastern {Australia}. Mar. Freshwater Res. \textbf{60}: 845--860.

\bibitem[\citeproctext]{ref-trinnie2014}
Trinnie, F.I., Walker, T.I., Jones, P.L., and Laurenson, L.J. 2014. Regional differences in the reproductive parameters of the sparsely-spotted stingaree, {\emph{Urolophus paucimaculatus}}, from south-eastern Australia. Mar. Freshwater Res. \textbf{65}(11): 943. doi: \href{https://doi.org/10.1071/mf13275}{10.1071/mf13275}.

\bibitem[\citeproctext]{ref-trinnie_reproductive_2016}
Trinnie, F.I., Walker, T.I., Jones, P.L., and Laurenson, L.J. 2016. Reproductive cycle of \emph{{Urolophus} cruciatus} in south-eastern {Australia}: {Does} the species exhibit obligate or facultative diapause? Mar. Biol. \textbf{163}(11): 226. doi: \href{https://doi.org/10.1007/s00227-016-2976-x}{10.1007/s00227-016-2976-x}.

\bibitem[\citeproctext]{ref-walker_fishery_1992}
Walker, T. 1992. Fishery simulation model for sharks applied to the {Gummy} {Shark}, \emph{{Mustelus} antarcticus} {Gunther}, from {Southern} {Australian} waters. Mar. Freshwater Res. \textbf{43}(1): 195--212. Available from \url{https://doi.org/10.1071/MF9920195}.

\bibitem[\citeproctext]{ref-walker_reproduction_2005}
Walker, T.I. 2005. Reproduction in fisheries science. \emph{In} Reproductive biology and phylogeny of {Chondrichthyans}: Sharks, batoids, and chimaeras. \emph{Edited by} W.C. Hamlett. Science Publishers Inc., Enfield. pp. 81--127.

\bibitem[\citeproctext]{ref-walker_spatial_2007}
Walker, T.I. 2007. Spatial and temporal variation in the reproductive biology of gummy shark \emph{{Mustelus} antarcticus} ({Chondrichthyes} : {Triakidae}) harvested off southern {Australia}. Mar. Freshwater Res. \textbf{58}: 67--97. doi: \href{https://doi.org/10.1071/MF06074}{10.1071/MF06074}.

\bibitem[\citeproctext]{ref-winton2014}
Winton, M.V., Wuenschel, M.J., and McBride, R.S. 2014. Investigating spatial variation and temperature effects on maturity of female winter flounder ({\emph{Pseudopleuronectes americanus}}) using generalized additive models. Can. J. Fish. Aquat. Sci. \textbf{71}(9): 1279--1290. doi: \href{https://doi.org/10.1139/cjfas-2013-0617}{10.1139/cjfas-2013-0617}.

\bibitem[\citeproctext]{ref-xiao_demographic_2000}
Xiao, Y.S., and Walker, T.I. 2000. Demographic analysis of gummy shark (\emph{{Mustelus} antarcticus}) and school shark (\emph{{Galeorhinus} galeus}) off southern {Australia} by applying a generalized {Lotka} equation and its dual equation. Can. J. Fish. Aquat. Sci. \textbf{57}: 214--222. doi: \href{https://doi.org/10.1139/f99-224}{10.1139/f99-224}.

\bibitem[\citeproctext]{ref-zuur_analyzing_2007}
Zuur, A., Ieno, E.N., and Smith, G.M. 2007. Analyzing ecological data. New York, NY.

\end{CSLReferences}

\end{document}

### 2024-05-19 rmds/figures-supp.Rmd
revisions

@@
 data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples)))
+  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples)))
 ```
@@
   background_grid() +
-  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - Length at 50% maternity")) +
+  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - Length at 50% maternity")) +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
-Figure S4. Precision (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
+Figure S4. Accuracy (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
 
@@
   background_grid() +
-  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - "~italic(P)[Max])) +
+  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - "~italic(P)[Max])) +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
@@
 Figure S5.
-Precision (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models.
+Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models.
 Each point reflects a mean value from 300 simulated data sets.
@@
   background_grid() +
-  labs(x = "Sample size", y = "Mean % relative absolute error", title = "Precision - Reproductive Output") +
+  labs(x = "Sample size", y = "Mean % relative absolute error", title = "Accuracy - Reproductive Output") +
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
@@
 Figure S11.
-Precision (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
+Accuracy (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
 Each point reflects a mean value from 300 simulated data sets

### 2024-05-19 rmds/figures.Rmd
revisions

@@
 Figure 3.
-Precision (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods.
-Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased.
+Accuracy (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods.
+Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased.
 Each point reflects a mean value from 300 simulated data sets.
@@
   filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
+  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
   filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
@@
 # Bin data into length categories for plotting
-brks = seq(40, 220, 7.5)
+brks = seq(40, 220, 10)
 data_binned <- data %>%
@@
 The grey shaded region denotes 95% confidence intervals based on bootstrap resampling.
-$P_{Max}$ was fixed at 0.5 in the lower panel
+Black points show proportion in maternal condition in 10cm length intervals.
+Marginal rug plots denote raw data that models were fit to.
+$P_{Max}$ was fixed at 0.5 in the lower panel.

### 2024-05-19 rmds/manuscript.Rmd
revisions

@@
 Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy.
-Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output.
+Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
 In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
-This study suggests that for sample sizes \> 100 estimation of $P_{Max}$ is generally preferable to using a fixed value.
+This study suggests that for sample sizes \> 100, estimation of $P_{Max}$ is generally preferable to using a fixed value.
 
@@
 In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
-Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015].
+Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of mature females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015].
 Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017].
@@
 Walker [-@walker_reproduction_2005] recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans.
-For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: $$Y_i \sim B(x_i; n_i, P_i)$$ $$E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x)}}$$ where the proportion of individuals in a given length or age category, $x_i$, that are in maternal condition, $Y_i$, is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, $P_i$, multiplied by the total number of individuals in the category, $n_i$.
+For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: $$Y_i \sim B(x_i; n_i, P_i)$$ $$E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x_i)}}$$ where the proportion of individuals in a given length or age category, $x_i$, that are in maternal condition, $Y_i$, is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, $P_i$, multiplied by the total number of individuals in the category, $n_i$.
 If $x$ is a continuous rather than categorical variable (i.e.
@@
 Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions.
-Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles.
+Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if a there is a resting period between pregnancies.
 Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically.
@@
 Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing $P_{Max}$ in maternity functions prevents statistical inference from being made on these data.
-Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles.
+Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity.
 To date only a single study on spiny dogfish, *Squalus acanthias*, has statistically estimated $P_{Max}$ from data [@colonello_reproductive_2016], and the feasibility of doing so for a wider range of species has not been investigated.
@@
 Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs.
-The performance of three methods in estimating 'true' maternity parameters from simulated data was tested by examining precision, bias, and interval coverage.
+The performance of three methods in estimating 'true' maternity parameters from simulated data was tested by examining accuracy, bias, and interval coverage.
 To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters.
 Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency.
-All simulation modelling and data analysis was conducted using the R language [@r_core_team_r:_2018] and the Template Model Builder (TMB) R package [@kristensen_tmb:_2016].
+All simulation modelling and data analysis was conducted using the R language [Version 4.2.0] [@r_core_team_r:_2022] and the Template Model Builder (TMB) R package [Version 1.9.1] [@kristensen_tmb:_2016].
 
@@
 Data were generated for the gummy shark, *Mustelus antarcticus*, and school shark, *Galeorhinus galeus*, using parameters from studies conducted on southern Australian populations (Table 1) [@grant_estimation_1979; @kirkwood_gill_1986; @moulton_age_1992; @walker_fishery_1992; @punt_stock_1998; @walker_reproduction_2005; @walker_spatial_2007].
-Both species are well-studied and were chosen to be representative of r- and K-selected chondrichthyan life history styles, respectively [@stevens_variable_1999].
+Both species are well-studied and were chosen to be representative of fast and slow chondrichthyan life history styles, respectively [@stevens_variable_1999].
 
@@
 
-Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This process was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.
+Next, dichotomous maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This process was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.
 
@@
 Two approaches to estimating maternity parameters were compared: the 3PLF itself (3PLF-estimated) and the 3PLF function with a fixed asymptote (3PLF-fixed).
-Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF maturity).
+Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF-maturity).
 The 3PLF-estimated method was undertaken to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016].
@@
 An iteration was also not attempted if there were no maternal females in the randomly generated sample.
-Model performance was evaluated by quantifying precision, bias, and interval coverage for $L'_{50}$, $P_{Max}$ and $R_0$.
-Bias was quantified using the mean relative error, $\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]$ and precision using the mean absolute relative error, $\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|$.
+Model performance was evaluated by quantifying accuracy, bias, and interval coverage for $L'_{50}$, $P_{Max}$ and $R_0$.
+Bias was quantified using the mean relative error, $\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]$ and accuracy using the mean absolute relative error, $\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|$.
 Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50% confidence intervals [@rudd_accounting_2018].
@@
 Data from two studies of the reproductive biology of the sandbar shark, *Carcharhinus plumbeus*, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods.
-The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{yr}^{-1}$) with a protracted reproductive cycle lasting longer than a year [@springer_natural_1960].
+The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{ yr}^{-1}$) with a protracted reproductive cycle lasting longer than a year [@springer_natural_1960].
 The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s [@sminkey_age_1995] and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes [@baremore_reproduction_2012; @piercy_histological_2016].
 In their study, Baremore and Hale [-@baremore_reproduction_2012] estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle.
-Piercy *et al* [-@piercy_histological_2016] did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles.
+Piercy et al.
+[-@piercy_histological_2016] did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles.
 To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with $P_{Max}$ estimated and fixed at 0.5 and 0.33, respectively.
@@
 The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5).
-Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved precision with the 3PLF methods, although at larger sample sizes better precision was achieved by estimating $P_{Max}$ (Figure 3).
+Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating $P_{Max}$ (Figure 3).
 Overall, $P_{Max}$ was comparatively more challenging to estimate accurately than $L'_{50}$ (Figure S4, Figure S5).
@@
 In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance.
-$L_{50}$ could usually be estimated with a higher precision than $L'_{50}$.
+$L_{50}$ could usually be estimated with a higher accuracy than $L'_{50}$.
 For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2%, exceeding the performance of the 3PLF-methods at most sample sizes.
@@
 Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10).
-For both species, better precision in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For both species, better accuracy in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
 For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$.
@@
 Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity.
-Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that precision, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated.
+Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that accuracy, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated.
 Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
@@
 
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
-In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this it is therefore advisable to fix $P_{Max}$.
-Larger sample sizes are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
-
 While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
@@
 This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999].
-In the case of the gummy shark this imprecision led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$.
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$.
 Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a sample size of \> 1000.
-A similar sample size was used by Colonello *et al* [-@colonello_reproductive_2016] to successfully estimate $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish.
+A similar sample size was used by Colonello *et al*.
+[-@colonello_reproductive_2016] to estimate plausible values of $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish.
 These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].
 
-In light of the data requirements suggested by this study, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+In light of the data requirements suggested by this study, placing a penalty function on $P_{Max}$ or using a the use of a prior distribution
+
+For some species, and data on pregnant females are particularly sparse, 
+the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 While several studies have shown these parameters to be similar [@walker_spatial_2007; @soto-lopez_reproductive_2018], $L'_{50}$ is more frequently shifted to the right of $L_{50}$ [@braccini_determining_2006; @palacios-hernandez_temporal_2020; @montealegre-quijano_sexual_2014; @colonello_reproductive_2016], and this length difference can equate to one or more years.
-For example, Harry *et al* [-@harry_age_2013] found both spot-tail shark, *C.
+For example, Harry et al.
+[-@harry_age_2013] found both spot-tail shark, *C.
 sorrah*, and Australian blacktip shark, *C.
 tilstoni*, began reproducing the year after reaching sexual maturity.
-Similarly, Fujiyama found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.
+Similarly, Fujinami et al.
+[-@fujinami_reproductive_2017] found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.
 
@@
 While not widely used, this approach has been undertaken in several studies [@mejiafalla_reproductive_2012; @baremore_reproduction_2012; @baremore_reproduction_2013; @rambahiniarison_life_2018].
+
+Another approach to overcoming the
+
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
+In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
+At sample sizes below this it is therefore advisable to fix $P_{Max}$.
+Larger sample sizes are necessary before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
 
@@
 The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating $P_{Max}$.
-Detailed reproductive studies of the western North Atlantic population by Baremore and Hale [-@baremore_reproduction_2012] and Piercy *et al* [-@piercy_histological_2016] were ambiguous about the frequency of reproduction.
+Detailed reproductive studies of the western North Atlantic population by Baremore and Hale [-@baremore_reproduction_2012] and Piercy et al.
+[-@piercy_histological_2016] were ambiguous about the frequency of reproduction.
 In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
@@
 
-One benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to $P_{Max}$.
+One benefit of this approach is that it reduces the need for subjective modelling assumptions relating to $P_{Max}$.
 Such assumptions can have important implications for population modelling.
-For example, to account for uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, *Urolophus paucimaculatus*, Trinnie *et al* (2014) estimated maternity parameters for ten plausible scenarios.
+For example, due to uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, *Urolophus paucimaculatus*, Trinnie *et al*.
+[@trinnie2014] estimated maternity parameters for ten plausible scenarios.
 In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle [@sedar_sedar_2017].
@@
 More commonly, practitioners have chosen fixed values of the form $P_{Max} = 1/n$, where *n* is the apparent duration of the reproductive cycle in years.
-This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently that the larger population.
+This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently than the larger population.
 Long-term study of offspring from genetically profiled lemon sharks, *Negaprion brevirostris*, confirmed that most individuals reproduced biennially, but also revealed some cases of triennial reproduction [@feldheim_two_2014].
@@
 Estimation of $P_{Max}$ using the 3PLF-estimated method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans.
-White sharks, *Carcharodon carcharias*, and whale shark, *Rhinchodon typus*,are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
+White sharks, *Carcharodon carcharias*, and whale shark, *Rhinchodon typus*, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
 In both species lack of data on reproductive frequency is a major impediment to population modelling and the development of management strategies [@huveneers_future_2018; @bowlby_implications_2020].
@@
 Little is known about how reproductive frequency varies as a function of size or age, in response to environmental effects, or at changing population densities.
-Better understanding of these factors has the potential to improve understanding reproductive strategies and is, many cases, of direct relevance to management.
+Better understanding of these factors has the potential to improve understanding reproductive strategies and is, in many cases, of direct relevance to management.
 

### 2024-05-19 code/empirical.R

@@ -91,6 +91,7 @@ data2 <- select(data_complete2, month, Stage, x, z) %>%
 data <- data %>% 
   ungroup() %>% select(month, Stage, x, z) %>%
   rbind(data2) %>% 
+  select(x, z) %>% 
   write_rds(here("data", "empirical-plumbeus.Rds"))
 
 # Run analyses ----------------------------------------------------------------

### 2024-05-19 code/selectivity.R

@@ -30,3 +30,137 @@ sel_df <- cross_df(sel_pars_gummy) %>%
   mutate(P = round(P, 2))
 
 # Selectivity parameters (Kirkwood and Walker 1986)
+
+# Housekeeping ----------------------------------------------------------------
+
+source(here::here("code", "setup.R"))
+source(here::here("code", "functions.R"))
+
+compile("code/logistic3.cpp")
+dyn.load(dynlib("code/logistic3"))
+compile("code/logistic2.cpp")
+dyn.load(dynlib("code/logistic2"))
+compile("code/logistic2maternal.cpp")
+dyn.load(dynlib("code/logistic2maternal"))
+
+# Parameter section  ----------------------------------------------------------
+
+sim_pars_mustelus <- list(
+  species = "Gummy shark",
+  
+  # Maximum age
+  AMAX = 20, 
+  # Natural mortality (Max age reported in Moulton et al 1992, Walker 1992 M)
+  M = 0.1970,
+  # Fishing mortality (Nominally set to 25 % of M)
+  F_mort = 0.25 * 0.1970,
+  # von Bertalanffy growth parameters (Moulton et al 1992, BS 1973-1976)
+  K = 0.123,
+  Linf = 2019,
+  t0 = -1.55,
+  cv_l = 0.05,
+  # Maternity parameters (Walker 2007 EKI 73-76)
+  m50 = 1129,
+  m95 = 1344,
+  c = c(1 / 2), # Proportion maternal
+  # Maturity parameters (Walker 2007 EKI 73-76)
+  l50 = 1105,
+  l95 = 1293,
+  # Uterine fecundity as function of length (Walker 2007) F = f*exp(g x length)
+  f = 0.2804,
+  g = 0.00286,
+  # Selectivity parameters (Kirkwood and Walker 1986)
+  # Correspond to peak sel at 25% and 75% length maternity
+  mesh = c(5.68, 6.57),
+  theta1 = 184.3,
+  theta2 = 29739,
+  
+  # Number of samples
+  Nsamples = c(1000)
+)
+
+
+
+
+data <-
+  # Generate data frame of all combinations of scenarios
+  cross_df(sim_pars_mustelus) %>%
+  # Create some additional variables
+  mutate(mesh_name = case_when(
+    mesh == 5.68 | mesh == 7.27 ~ "low",
[…]

### 2024-05-27 reports/main-text.tex
revisions

@@
 Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy.
-Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output.
+Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
@@
 In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts.
-Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one (Frisk et al.
+Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of mature females that will give birth in the population in a given season is less than one (Frisk et al.
 2005, Dulvy et al.
@@
 Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data.
-Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles.
+Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity.
 To date only a single study on spiny dogfish, \emph{Squalus acanthias}, has statistically estimated \(P_{Max}\) from data (Colonello et al.
@@
 Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs.
-The performance of three methods in estimating `true' maternity parameters from simulated data was tested by examining precision, bias, and interval coverage.
+The performance of three methods in estimating `true' maternity parameters from simulated data was tested by examining accuracy, bias, and interval coverage.
 To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters.
@@
 The following procedure was used to `guess' the fixed value of \(P_{Max}\) in each simulation.
-\(P_{Max}\) was chosen as the proportion of females in maternal condition above a the length at which \(99\%\) of females were mature, \(L > \Psi(L_{99\%})\).
+\(P_{Max}\) was chosen as the proportion of females in maternal condition above the length at which \(99\%\) of females were mature, \(L > \Psi(L_{99\%})\).
 In some cases this procedure failed due to there being no maternal females in the simulated data set meeting this criteria.
@@
 An iteration was also not attempted if there were no maternal females in the randomly generated sample.
-Model performance was evaluated by quantifying precision, bias, and interval coverage for \(L'_{50}\), \(P_{Max}\) and \(R_0\).
-Bias was quantified using the mean relative error, \(\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]\) and precision using the mean absolute relative error, \(\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|\).
+Model performance was evaluated by quantifying accuracy, bias, and interval coverage for \(L'_{50}\), \(P_{Max}\) and \(R_0\).
+Bias was quantified using the mean relative error, \(\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]\) and accuracy using the mean absolute relative error, \(\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|\).
 Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50\% confidence intervals (Rudd and Thorson 2018).
@@
 In their study, Baremore and Hale (2012) estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle.
-Piercy \emph{et al.} (2016) did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles.
+Piercy et al.
+(2016) did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles.
 To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with \(P_{Max}\) estimated and fixed at 0.5 and 0.33, respectively.
@@
 
-Parameter estimation was generally straightforward; with sample sizes \(> 50\) a success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1).
+Parameter estimation was generally straightforward; with sample sizes \(> 50\) a convergence success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1).
 Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1.
@@
 The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5).
-Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved precision with the 3PLF methods, although at larger sample sizes better precision was achieved by estimating \(P_{Max}\) (Figure 3).
+Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3).
 Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).
@@
 In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance.
-\(L_{50}\) could usually be estimated with a higher precision than \(L'_{50}\).
+\(L_{50}\) could usually be estimated with a higher accuracy than \(L'_{50}\).
 For the gummy shark, where \(L_{50}\) and \(L'_{50}\) were relatively similar, using the 2PLF-maturity method led to a constant underestimate of \(L'_{50}\) of approximately 2\%, exceeding the performance of the 3PLF-methods at most sample sizes.
@@
 Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10).
-For both species, better precision in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
 For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).
@@
 Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity.
-Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that precision, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated.
+Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated.
 Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes.
@@
 
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\) from data, but also identified situations where it may be preferable to fix \(P_{Max}\).
-In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this it is therefore advisable to fix \(P_{Max}\).
-Larger sample sizes are desirable before attempting to estimate \(P_{Max}\) in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as \(P_{Max}\) decreases.
-
 While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
@@
 1999).
-In the case of the gummy shark this imprecision led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\).
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\).
 Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000.
 A similar sample size was used by Colonello \emph{et al}.
-(2016) to successfully estimate \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.
+(2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.
 These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
@@
 
-In light of the data requirements suggested by this study, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution
+
+For some species, and data on pregnant females are particularly sparse,
+the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 2020), and this length difference can equate to one or more years.
-For example, Harry \emph{et al.} (2013) found both spot-tail shark, \emph{C.
+For example, Harry et al.
+(2013) found both spot-tail shark, \emph{C.
 sorrah}, and Australian blacktip shark, \emph{C.
 tilstoni}, began reproducing the year after reaching sexual maturity.
-Similarly, Fujiyama found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.
+Similarly, Fujinami et al.
+(2017) found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.
 
@@
 
+Another approach to overcoming the
+
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\) from data, but also identified situations where it may be preferable to fix \(P_{Max}\).
+In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
+At sample sizes below this it is therefore advisable to fix \(P_{Max}\).
+Larger sample sizes are necessary before attempting to estimate \(P_{Max}\) in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as \(P_{Max}\) decreases.
+
 The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for \(P_{Max}\).
@@
 The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\).
-Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy \emph{et al.} (2016) were ambiguous about the frequency of reproduction.
+Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy et al.
+(2016) were ambiguous about the frequency of reproduction.
 In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
@@
 
-One benefit of this approach is that it that it reduces the need for subjective modelling assumptions relating to \(P_{Max}\).
+One benefit of this approach is that it reduces the need for subjective modelling assumptions relating to \(P_{Max}\).
 Such assumptions can have important implications for population modelling.
-For example, to account for uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, \emph{Urolophus paucimaculatus}, Trinnie \emph{et al}.
+For example, due to uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, \emph{Urolophus paucimaculatus}, Trinnie \emph{et al}.
 (Trinnie et al.
@@
 More commonly, practitioners have chosen fixed values of the form \(P_{Max} = 1/n\), where \emph{n} is the apparent duration of the reproductive cycle in years.
-This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently that the larger population.
+This process also has the potential to introduce bias if a small proportion of females reproduce more or less frequently than the larger population.
 Long-term study of offspring from genetically profiled lemon sharks, \emph{Negaprion brevirostris}, confirmed that most individuals reproduced biennially, but also revealed some cases of triennial reproduction (Feldheim et al.

### 2024-05-27 rmds/figures-supp.Rmd
revisions

@@
 ggplot(data %>% filter(par_name %in% "m50")) +
-  geom_point(aes(as.factor(Nsamples), conv_success, col = repro_freq, group = repro_freq)) + 
+  geom_point(aes(as.factor(Nsamples), conv_success, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_color_brewer(palette = "Spectral") + 
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), conv_success, col = repro_freq, group = repro_freq)) +
   facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) +
-  scale_color_brewer(palette = "Spectral") + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) + 
   cowplot::theme_half_open() +
@@
   labs(x = "Number of samples", y = "Convergence success (%)", title = "Model convergence") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+
 ```
@@
   geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) +
+  geom_point(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_color_brewer(palette = "Spectral") + 
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), rel_error, col = repro_freq, group = repro_freq)) + 
   facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) +  
   cowplot::theme_half_open() +
@@
   labs(x = "Sample size", y = "Mean % relative error", title = expression("Bias - Length at 50% maternity")) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 
@@
   labs(x = "Sample size", y = "Mean % relative error", title = expression("Bias - "~italic(P)[Max])) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 
@@
   geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
+  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_color_brewer(palette = "Spectral") + 
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
   facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) +  
   cowplot::theme_half_open() +
@@
   labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - Length at 50% maternity")) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+
 ```
@@
   labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - "~italic(P)[Max])) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+
 ```
@@
   geom_hline(yintercept = 50, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
+  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_color_brewer(palette = "Spectral") + 
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
   facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
-  scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) +  
   cowplot::theme_half_open() +
@@
   labs(x = "Sample size", y = "% interval coverage", title = expression("Interval coverage - Length at 50% maternity")) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
@@
   labs(x = "Sample size", y = "% interval coverage", title = expression("Interval coverage - "~italic(P)[Max])) +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+
 ```
@@
   labs(x = "Sample size", y = "Mean % relative error", title = "Bias - Reproductive Output") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
@@
   labs(x = "Sample size", y = "Mean % relative absolute error", title = "Accuracy - Reproductive Output") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+
 ```
@@
 Each point reflects a mean value from 300 simulated data sets
+
+```{r fig.width = 7, fig.height= 5.5}
+data <- read_rds(here("data", "simulations-pivot.Rds")) %>% 
+  filter(method_long %in% "3PLF - estimated")
+
+ggplot() + 
+  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, fill = repro_freq, size = n_maternal), pch = 21, alpha = 0.5) +
+  scale_size_continuous(breaks = c(2, 10, 50, 250, 1250)) + 
+  scale_fill_brewer(palette = "Spectral") + 
+  facet_grid( ~ species) + 
+  guides(fill = guide_legend(title = bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))), order = 1), size = guide_legend("No. maternal females")) + 
+  theme_bw() +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
+  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
+  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
+```
+
+Figure S12.
+Uncropped version of Figure 2.
+Bias (per cent relative error) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for the 3PLF maternity function with $P_{Max}$ estimated.
+Each point represents parameter estimates from one iteration of simulated data, including all combinations of variables.
+Simulations with longer reproductive cycles and fewer maternal females, were associated with higher bias in both parameters $\hat{L'_{50}}$ and $\hat{P_{Max}}$.

### 2024-05-27 rmds/figures.Rmd
revisions

@@
 ```{r fig.width = 7, fig.height= 5.5}
-data <- read_rds(here("data", "simulations-pivot.Rds"))
+data <- read_rds(here("data", "simulations-pivot.Rds")) %>% 
+  filter(method_long %in% "3PLF - estimated")
 
 ggplot() + 
-  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
-  scale_color_viridis_d(option = "D", direction = 1) + 
-  facet_grid(species~method_long) + 
-  ylim(-30, 50) +
-  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
+  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, fill = repro_freq, size = n_maternal), pch = 21, alpha = 0.5) +
+  scale_size_continuous(breaks = c(2, 10, 50, 250, 1250)) + 
+  scale_fill_brewer(palette = "Spectral") + 
+  facet_grid( ~ species) + 
+  ylim(-30,50) +
+  guides(fill = guide_legend(title = bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))), order = 1), size = guide_legend("No. maternal females")) + 
   theme_bw() +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
   xlab(expression("% relative error "~hat(italic(P)[Max]))) +
@@
 ```
+
 Figure 2.
-Bias (per cent relative error) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods.
-Each point represents parameter estimates from one iteration of simulated data.
-Bias tended to increase as the true underlying $P_{Max}$ decreased.
-\
+Bias (per cent relative error) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for the 3PLF maternity function with $P_{Max}$ estimated.
+Each point represents parameter estimates from one iteration of simulated data (n = 43,129), including all combinations of variables.
+Simulations with longer reproductive cycles and fewer maternal females were associated with higher bias in both parameters $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
+Note: 42 data points were cropped to aid with data visualization (see Figure S12 for uncropped figure).\
 \
@@
   geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
+  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
@@
   scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) + 
   theme_bw() + 
   labs(x = "Sample size", y = "Mean % absolute relative error") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
+
 Figure 3.
@@
 Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased.
-Each point reflects a mean value from 300 simulated data sets.
-\
+Each point reflects a mean value from 300 simulated data sets.\
 \
@@
 ```{r fig.width = 7, fig.height= 5.5}
+data <- read_rds(here("data", "simulations-summarised.Rds"))
+
 data %>% 
@@
   geom_hline(yintercept = 50, linetype = "dashed") +
-  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
+  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq, size = n_maternal), alpha = 0.5) +
+  scale_size_continuous(breaks = c(10, 50, 250, 1250)) + 
   geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
@@
   scale_color_brewer(palette = "Spectral") + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))), order = 1),
+         size = guide_legend("No. maternal females")) + 
   theme_bw() + 
   labs(x = "Sample size", y = "% interval coverage") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
+  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```
+
 Figure 4.
 Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
-Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
-\
+Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.\
 \
@@
 
-```{r fig.width = 9, fig.height= 3.5}
+```{r fig.width=7, fig.height= 4}
+(p <- ggplot(data = data %>% 
+  filter(method_long %in% c("3PLF - estimated"), par_name %in% "c"), aes(x = n_maternal, y = abs_error, col = repro_freq)) + 
+  geom_point() + 
+  scale_color_brewer(palette = "Spectral") +
+  scale_y_continuous(breaks = seq(0, 90, 10)) +
+  scale_x_continuous(breaks = seq(0, 1600, 200)) +
+  facet_grid(~ species) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  theme_bw() + 
+  xlab("Median number of maternal females") + 
+  ylab(expression("Mean % absolute error "~hat(italic(P)[Max]))) +
+  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center"))
+```
 
-read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
-  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
-  slice_min(abs(rel_error)) %>% 
-  ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_viridis_d() + 
-  guides(fill = guide_legend(title = "Method")) + 
-  facet_wrap(~species, nrow = 1) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90)) + 
-  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
-```
 Figure 5.
-Performance of alternative maternity functions in minimising bias in calculations of $R_0$.
-The preferred method was that which minimised bias, $|\text{relative error}|$ across 300 simulated datasets.
-Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
-\
-\
+Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ for the 3PLF - estimated method as a function of number of maternal females.
+Each point reflects a value from 300 simulated data sets.\
 \newpage
@@
   geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_viridis_d() + 
+  scale_fill_brewer(palette = "Spectral") + 
   guides(fill = guide_legend(title = "Method")) + 
@@
 ```
+
 Figure 6.
-Performance of alternative maternity functions in accurately calculating $R_0$.
+Performance of four alternative maternity functions in accurately calculating $R_0$.
 The best performing method was that which minimised mean absolute error across 300 simulated datasets.
-Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
-\
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.\
 \newpage

### 2024-05-27 rmds/manuscript.Rmd
revisions

@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
+Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
@@
 
-Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions [@cortes_population_2012].
+Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions [@cortes_population_2012; @swenson2024].
 It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles.
 Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.
-annual, biennial), despite observations of plasticity in this trait within discrete populations [@higgs_reproductive_2020].
+annual, biennial), despite observations of plasticity in this trait within discrete populations [@higgs_reproductive_2020; @nosal2021].
 Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing $P_{Max}$ in maternity functions prevents statistical inference from being made on these data.
-Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity.
+Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles.
 To date only a single study on spiny dogfish, *Squalus acanthias*, has statistically estimated $P_{Max}$ from data [@colonello_reproductive_2016], and the feasibility of doing so for a wider range of species has not been investigated.
@@
 To examine the effect of varying sample size, six scenarios were conducted with catch ranging from 50 to 2500 individuals, reflecting a gradient from data poor to rich.
+Length data were generated by randomly sampling, with replacement, from a vector of length classes, weighted by a vector of probabilities determined from the above catch equation.
 
@@
 The following procedure was used to 'guess' the fixed value of $P_{Max}$ in each simulation.
-$P_{Max}$ was chosen as the proportion of females in maternal condition above a the length at which $99\%$ of females were mature, $L > \Psi(L_{99\%})$.
+$P_{Max}$ was chosen as the proportion of females in maternal condition above the length at which $99\%$ of females were mature, $L > \Psi(L_{99\%})$.
 In some cases this procedure failed due to there being no maternal females in the simulated data set meeting this criteria.
@@
 
-Parameter estimation was generally straightforward; with sample sizes $> 50$ a success rate of $\geq 95\%$ was achieved in all scenarios (Figure S1, Table S1).
+Parameter estimation was generally straightforward; with sample sizes $> 50$ a convergence success rate of $\geq 95\%$ was achieved in all scenarios (Figure S1, Table S1).
 Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where $P_{Max}$ was always fixed at 1.
-In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.
+In addition to convergence failures, 12 simulations also generated zero maternal females, and were discarded.
 
 Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased.
-The nature of this problem can be seen in the positive correlation between the $\hat{L'_{50}}$ and $\hat{P_{Max}}$ parameters whereby $P_{Max}$ affects the magnitude and direction of bias in these parameters (Figure 2).
-With lower values of $P_{Max}$ it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$ (Figure 2).
+This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2).
+With lower values of $P_{Max}$ it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
+As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$.
+Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).
 
@@
 
-Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
+Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters.
+Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9).
+This was particularly the case for $P_{Max}$ which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5).
 These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure.
-For example, in low selectivity scenarios for gummy sharks \~60% of individuals were immature compared to \~50% in the corresponding scenarios for school sharks (Figure S9).
-As $P_{Max}$ decreased the proportion of maternal females in the analysis also decreased.
+For example, in biennial low selectivity scenarios for gummy sharks \~60% of individuals were immature compared to \~50% in the corresponding scenarios for school sharks (Figure S9).
 
@@
 The effect of the different methods in ultimately calculating $R_0$ varied considerably across the various scenarios.
-For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10).
+For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios (Figure S10).
 Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of $R_0$ by 3.4% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1%.
-(Table S1, Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10).
+(Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure S10).
 Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10).
@@
 
-Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$Stage > 2))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$Stage > 2))))`% were in maternal condition.
+Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$y == 1))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$y == 1))))`% were in maternal condition.
 Using the 3PLF-estimated method, the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
-Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence.
+Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence (Figure 7).
 For 3PLF-fixed models a value of $P_{Max} = 0.5$ was strongly supported ($\Delta_{AIC}$ = 14.15) over the alternative fixed value of $P_{Max} = 0.33$.
 The model with $P_{Max} = 0.5$ also outperformed the 3PLF-estimated method ($\Delta_{AIC}$ = 1.86).
-Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data [@burnham_model_2002].
+Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data [@burnham_model_2002].
 
@@
 
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
+Providing clear guidelines on this is difficult
+
+In most scenarios, sample sizes of upward of 100 maternal females were needed for the 3PLF-estimated method to consistently exceed the performance of the 3PLF-fixed method. 
+At sample sizes below this it is therefore advisable to fix $P_{Max}$.
+Larger sample sizes are necessary before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
+
 While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
@@
 
-For some species, and data on pregnant females are particularly sparse, 
-the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+For some species, and data on pregnant females are particularly sparse, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 Another approach to overcoming the
-
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
-In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this it is therefore advisable to fix $P_{Max}$.
-Larger sample sizes are necessary before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
 

### 2024-05-27 rmds/manuscript.tex
revisions

--- 2024-05-19 reports/main-text.tex
@@
 Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy.
-Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output.
+Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
@@
 Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions (Cortés et al.
-2012).
+2012, Swenson et al.
+2024).
 It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles.
 Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.~annual, biennial), despite observations of plasticity in this trait within discrete populations (Higgs et al.
-2020).
+2020, Nosal et al.
+2021).
 Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data.
@@
 
-Maternal data were re-analysed for 1087 sandbar sharks including 0 mature individuals of which \ensuremath{\infty{}}\% were in maternal condition.
+Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition.
 Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2).
-Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence.
+Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7).
 For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\).
 The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86).
-Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data (Burnham and Anderson 2002).
+Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data (Burnham and Anderson 2002).
 
@@
 
-In light of the data requirements suggested by this study, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution
+
+For some species, and data on pregnant females are particularly sparse, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 
+\bibitem[\citeproctext]{ref-nosal2021}
+Nosal, A.P., Cartamil, D.P., Ammann, A.J., Bellquist, L.F., Ben-Aderet, N.J., Blincow, K.M., Burns, E.S., Chapman, E.D., Freedman, R.M., Klimley, A.P., Logan, R.K., Lowe, C.G., Semmens, B.X., White, C.F., and Hastings, P.A.
+2021.
+Triennial migration and philopatry in the critically endangered soupfin shark {\emph{Galeorhinus galeus}}.
+J.
+Appl.
+Ecol.
+\textbf{58}(8): 1570--1582.
+doi: \href{https://doi.org/10.1111/1365-2664.13848}{10.1111/1365-2664.13848}.
+
 \bibitem[\citeproctext]{ref-oddone_size_2010}
@@
 \textbf{24}(3): 175--184.
+
+\bibitem[\citeproctext]{ref-swenson2024}
+Swenson, J.D., Brooks, E.N., Kacev, D., Boyd, C., Kinney, M.J., Marcy-Quay, B., Sévêque, A., Feldheim, K.A., and Komoroske, L.M.
+2024.
+Accounting for unobserved population dynamics and aging error in close{-}kin mark{-}recapture assessments.
+Ecology and Evolution \textbf{14}(2).
+doi: \href{https://doi.org/10.1002/ece3.10854}{10.1002/ece3.10854}.
 

### 2024-05-27 code/age-underestimation.R

@@ -1,0 +1,97 @@
+# Housekeeping ----------------------------------------------------------------
+
+source(here::here("code", "setup.R"))
+source(here::here("code", "functions.R"))
+
+compile("code/logistic3.cpp")
+dyn.load(dynlib("code/logistic3"))
+compile("code/logistic2.cpp")
+dyn.load(dynlib("code/logistic2"))
+compile("code/logistic2maternal.cpp")
+dyn.load(dynlib("code/logistic2maternal"))
+
+
+sim_pars_galeus <- list(
+  species = "School shark",
+  
+  # Maximum age
+  AMAX = c(27, 54), 
+  # Natural mortality (Grant et al 1979)
+  M = 0.1075,
+  # Fishing mortality (Nominally set to 25 % of M)
+  F_mort = 0.25 * 0.1075,
+  # von Bertalanffy growth parameters (Grant et al 1979)
+  K = 0.1600,
+  Linf = 1618.3,
+  t0 = -1.2818,
+  cv_l = 0.05,
+  # Maternity parameters (Walker 2005)
+  m50 = 1421,
+  m95 = 1488,
+  c = c(1 / 3), # Proportion maternal
+  # Maturity parameters (Walker 2005)
+  l50 = 1349,
+  l95 = 1502,
+  # Uterine fecundity as function of length (Walker 2005)
+  f = -46.0,
+  g = 0.0491,
+  # Selectivity parameters (Punt and Walker 1998)
+  # Correspond to peak sel at 25% and 75% length maternity
+  mesh = c(7.53),
+  theta1 = 192,
+  theta2 = 67595,
+  
+  # Number of samples
+  Nsamples = c(100, 1000, 2500)
+)
+
+data <-
+  # Generate data frame of all combinations of scenarios
+  cross_df(sim_pars_galeus) %>%
+  # Create some additional variables
+  mutate(mesh_name = case_when(
+    mesh == 5.68 | mesh == 7.27 ~ "low",
+    mesh == 6.57 | mesh == 7.53 ~ "high"
+  )) %>%
+  mutate(repro_freq = case_when(
+    c == 1 ~ "1 (Annual)",
+    c == 3 / 4 ~ "0.75 (Annual / biennial)",
+    c == 2 / 3 ~ "0.66 (Annual / biennial)",
+    c == 0.5 ~ "0.5 (Biennial)",
+    c == 1 / 3 ~ "0.33 (Triennial)",
+    c == 0.25 ~ "0.25 (Quadrennial)"
+  )) %>%
[…]

### 2024-05-27 code/empirical.R

@@ -91,7 +91,8 @@ data2 <- select(data_complete2, month, Stage, x, z) %>%
 data <- data %>% 
   ungroup() %>% select(month, Stage, x, z) %>%
   rbind(data2) %>% 
-  select(x, z) %>% 
+  mutate(y = ifelse(Stage > 2, 1, 0)) %>% 
+  select(x, y, z) %>% 
   write_rds(here("data", "empirical-plumbeus.Rds"))
 
 # Run analyses ----------------------------------------------------------------

### 2024-05-27 code/extra-plots.R

@@ -1,0 +1,68 @@
+data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  #filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  filter(method_long %in% c("3PLF - estimated")) %>% 
+  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "c") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs_error)
+
+data %>% 
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_brewer(palette = "Spectral") + 
+  #scale_fill_viridis_c() + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
+
+
+
+(p <- ggplot(data = data, aes(x = n_maternal, y = as.numeric(method_long))) + 
+  geom_point() + 
+  geom_smooth() + 
+  scale_x_continuous(breaks = seq(0,1600,200)) + 
+    facet_wrap(~species))
+
+
+(p <- ggplot(data = data, aes(x = n_maternal, y = abs_error, col = repro_freq)) + 
+  geom_point() + 
+  scale_color_brewer(palette = "Spectral") +
+  scale_y_continuous(breaks = seq(0, 90, 10)) +
+  facet_grid( ~ species) + 
+  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
+  theme_bw() + 
+  labs(x = "Median number of maternal females", y = "Mean % absolute relative error") +
+  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") +
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center"))
+
+
+data <- read_rds(here("data", "simulations-pivot.Rds")) %>% 
+  filter(method_long %in% "3PLF - estimated")
+
+(p <- ggplot() + 
[…]

### 2024-05-27 code/functions.R

@@ -128,13 +128,6 @@ generate_maternal_data <- function(pars, data){
     bind_cols("y" = data_mature$y)
 }
 
-# #
-# id = data$id
-# iteration <- data$iteration
-# n_maternal <- data$n_maternal
-# pars <- data$pars[[1]]
-# data <- data$data[[1]]
-
 
 ## Run simulations
 run_sims <- function(data, pars, n_maternal, id, iteration, ...) {

### 2024-05-27 code/summarise_results.R

@@ -5,25 +5,26 @@ source(here::here("code", "functions.R"))
 
 # Summarise simulation study --------------------------------------------------
 data <- read_rds(here("data", "simulations.Rds")) %>% 
-  group_by(id) %>% nest() %>% 
+  group_by(id) %>% 
+  nest() %>% 
   # How many iterations in each scenario had zero pregnant females and were excluded? 
   mutate(zero_maternal = map_dbl(data, ~ sum(.$no_maternal))) %>% 
   unnest(cols = c(data)) %>% 
   ungroup() %>% 
   filter(no_maternal == 0) %>% select(-no_maternal) %>% 
-  # How many other iterations in each scenario failed due to lack of pregnant females? 
-  mutate(few_maternal = ifelse(is.na(method), 1, 0)) %>% 
-  group_by(id) %>% 
-  nest() %>% 
-  mutate(low_maternal = map_dbl(data, ~ sum(.$few_maternal))) %>% 
-  unnest(cols = c(data)) %>% 
-  ungroup() %>% 
-  filter(!is.na(method)) %>% 
+  # # How many other iterations in each scenario failed due to lack of pregnant females? 
+  # mutate(few_maternal = ifelse(n_maternal == 1, 1, 0)) %>% 
+  # group_by(id) %>% 
+  # nest() %>% 
+  # mutate(low_maternal = map_dbl(data, ~ sum(.$few_maternal))) %>% 
+  # unnest(cols = c(data)) %>% 
+  # ungroup() %>% 
+  # filter(few_maternal == 0) %>% select(-few_maternal) %>% 
   # Calculate total number of iterations that failed to start due to lack of data
-  mutate(maternal_data_fail = low_maternal + zero_maternal) %>% 
+  mutate(maternal_data_fail =  zero_maternal) %>% 
   # Calculate total number of iterations done for each simulation
   mutate(ntotal = 300 - maternal_data_fail) %>% 
-  select(-low_maternal, -few_maternal, -zero_maternal) %>% 
+  select(-zero_maternal) %>% 
   # Now tally parameter boundary failures and non-convergence failures
   group_by(id, iteration, method) %>%
   nest() %>% 
@@ -72,13 +73,21 @@ write_rds(data, here("data", "simulations-summarised.Rds"))
 
 # Pivot table of simulation study --------------------------------------------------
 data <- read_rds(here("data", "simulations.Rds")) %>%
[…]

### 2024-05-28 rmds/figures.Rmd
revisions

@@
 Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased.
-Each point reflects a mean value from 300 simulated data sets.\
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
+\
 \
@@
 Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
-Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.\
+Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
+Point size denotes mean number of females in maternal condition at a given sample size.
+\
 \

### 2024-05-28 rmds/manuscript.Rmd
revisions

@@
 In some scenarios, particularly with sample sizes $\leq$ 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3).
-The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5).
+The accuracy of parameter estimates varied considerably across different combinations of variables used and among parameters (Figure S4, Figure S5).
 Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating $P_{Max}$ (Figure 3).
@@
 Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9).
+Certain combinations of variables in the simulated data resulted in few females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9).
 This was particularly the case for $P_{Max}$ which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5).
@@
 Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
+Overall, relatively large sample sizes were needed to estimated $P_{Max}$ accurately, particularly in species with longer reproductive cyles.
 This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
@@
 Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
-Providing clear guidelines on this is difficult
-
-In most scenarios, sample sizes of upward of 100 maternal females were needed for the 3PLF-estimated method to consistently exceed the performance of the 3PLF-fixed method. 
+Providing clear guidelines on this is difficult as model performance was influenced by life history traits and gear selectivity.
+In most scenarios, sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
 At sample sizes below this it is therefore advisable to fix $P_{Max}$.
@@
 While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
-Outcomes of the simulation study suggested that samples sizes of $\geq500$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species.
+Outcomes of the simulation study showed that for non-annual reproductive cycles $\geq100$ and $\geq200$ maternal females were typically needed to estimate $P_{Max}$ with \<10% MARE for school and gummy sharks, respectively.
 This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999].
 In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$.
-Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a sample size of \> 1000.
+Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition.
 A similar sample size was used by Colonello *et al*.
 [-@colonello_reproductive_2016] to estimate plausible values of $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish.
+
 These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].
+For smaller samples sizes, placing a prior or penalty on $P_{Max}$ may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species [@10.1371/journal.pone.0246734].
+Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with $L_{50}$ effectively providing a lower bound for $L'_{50}$.
 
-In light of the data requirements suggested by this study, placing a penalty function on $P_{Max}$ or using a the use of a prior distribution
-
-For some species, and data on pregnant females are particularly sparse, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+For particularly data-poor species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 While not widely used, this approach has been undertaken in several studies [@mejiafalla_reproductive_2012; @baremore_reproduction_2012; @baremore_reproduction_2013; @rambahiniarison_life_2018].
-
-Another approach to overcoming the
-
-The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for $P_{Max}$.
-In this study, the value for $P_{Max}$ used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data.
-The potential to introduce a greater level of bias in the analysis through the incorrect selection of $P_{Max}$ should also be considered if using this approach.
 

### 2024-05-28 code/extra-plots.R

@@ -1,15 +1,16 @@
 data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
   #filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  filter(method_long %in% c("3PLF - estimated")) %>% 
+  #filter(method_long %in% c("3PLF - estimated")) %>% 
   mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
-  filter(par_name %in% "c") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  mutate(n_maternal_cut = cut(n_maternal, breaks = seq(0, 1600, 20))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
   slice_min(abs_error)
 
 data %>% 
   ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  scale_fill_brewer(palette = "Spectral") + 
-  #scale_fill_viridis_c() + 
+  geom_tile(aes(x = as.factor(n_maternal_cut), y = repro_freq, fill = method_long)) + 
+  #scale_fill_brewer(palette = "Spectral") + 
+  scale_fill_viridis_d() + 
   guides(fill = guide_legend(title = "Method")) + 
   facet_wrap(~species, nrow = 1) +
   theme_bw() +
@@ -53,7 +54,9 @@ data <- read_rds(here("data", "simulations-pivot.Rds")) %>%
 
 
 data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  filter(method_long %in% c("3PLF - estimated"))
+  filter(method_long %in% c("3PLF - estimated")) %>% 
+  filter(abs_error <= 10) %>% 
+  filter(par_name %in% "c")
 
 (p <- ggplot(data = data, aes(x = n_maternal, y = abs_error, col = repro_freq)) + 
     geom_point() + 

### 2024-05-31 rmds/figures.Rmd
revisions

@@
 Figure 3.
-Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ for 3PLF methods.
+Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ for 3PLF methods with high maternal selectivity.
 Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased.
@@
 Figure 4.
-Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios).
+Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods with high maternal selectivity.
 Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.

### 2024-05-31 rmds/manuscript.Rmd
revisions

@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+However, sample sizes of 100 - 200 maternal females were required for accurate estimation of $P_{Max}$.
 Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
@@
 In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
-This study suggests that for sample sizes \> 100, estimation of $P_{Max}$ is generally preferable to using a fixed value.
 
@@
 Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions.
-Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if a there is a resting period between pregnancies.
+Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if there is a resting period between pregnancies.
 Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically.
@@
 
-Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased.
-This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2).
+When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased.
+This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby true underlying $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2).
 With lower values of $P_{Max}$ it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
@@
 Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
-Overall, relatively large sample sizes were needed to estimated $P_{Max}$ accurately, particularly in species with longer reproductive cyles.
-This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
+Overall, however, relatively large sample sizes were needed to accurately estimated $P_{Max}$, particularly in species with longer reproductive cycles.
+This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$.
-Providing clear guidelines on this is difficult as model performance was influenced by life history traits and gear selectivity.
-In most scenarios, sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this it is therefore advisable to fix $P_{Max}$.
-Larger sample sizes are necessary before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.
-
-While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
-Outcomes of the simulation study showed that for non-annual reproductive cycles $\geq100$ and $\geq200$ maternal females were typically needed to estimate $P_{Max}$ with \<10% MARE for school and gummy sharks, respectively.
-This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999].
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$, but was difficult to do so accurately.
+For non-annual reproductive cycles, $> 100$ school sharks and $> 200$ gummy sharks in maternal condition were typically needed to estimate $P_{Max}$ with \<10% MARE (Figure 5).
+This indicates that the both sample sizes and levels of uncertainty are likely to be larger than conventional logistic maturity analysis [@roa_estimation_1999].
 In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$.
@@
 
-These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].
+The simulation study also also identified some situations where it may be preferable to fix $P_{Max}$, particularly at lower sample sizes ($\le$ 100 individuals total).
+Providing clear guidelines on this is difficult as model performance was species-specific and influenced by life history and gear selectivity.
+The success of any analysis using a fixed value also ultimately depends on the analyst choosing a suitable value.
+The potential to introduce a greater level of bias through the incorrect selection of $P_{Max}$ should also be considered if using this approach [@natanson_presence_2019].
+
+Together, these results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].
 For smaller samples sizes, placing a prior or penalty on $P_{Max}$ may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species [@10.1371/journal.pone.0246734].
@@
 
-For particularly data-poor species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
+For the numerous data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size [@hale_age_2013].
+As such, the modal reproductive frequency is still at least triennial.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
@@
 We also thank Alex Hesp for valuable comments on an early draft of this paper and generosity in sharing R code.
+We are grateful to three anonymous reviewers for their detailed and constructive comments.
 

### 2024-05-31 rmds/manuscript.tex
revisions

@@
 
+\begin{itemize}
+\tightlist
+\item
+  A sentence with caveat about the need for large sample sizes to produce ``feasible'' results is warranted.
+\end{itemize}
+
 \textbf{Key words} Logistic regression; reproductive biology; maturity ogive; fisheries management
@@
 Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1.
-In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.
-
-Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased.
-The nature of this problem can be seen in the positive correlation between the \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) parameters whereby \(P_{Max}\) affects the magnitude and direction of bias in these parameters (Figure 2).
-With lower values of \(P_{Max}\) it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\) (Figure 2).
+In addition to convergence failures, 12 simulations also generated zero maternal females, and were discarded.
+
+When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased.
+This is seen in the positive correlation between \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2).
+With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
+As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\).
+Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).
 
@@
 In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3).
-The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5).
+The accuracy of parameter estimates varied considerably across different combinations of variables used and among parameters (Figure S4, Figure S5).
 Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3).
@@
 
-Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
+Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters.
+Certain combinations of variables in the simulated data resulted in few females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9).
+This was particularly the case for \(P_{Max}\) which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5).
 These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure.
-For example, in low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).
-As \(P_{Max}\) decreased the proportion of maternal females in the analysis also decreased.
+For example, in biennial low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).
 
@@
 The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios.
-For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10).
+For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure S10).
 Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%.
-(Table S1, Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10).
+(Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure S10).
 Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10).
@@
 Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes.
-This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
+Overall, however, relatively large sample sizes were needed to accurately estimated \(P_{Max}\), particularly in species with longer reproductive cycles.
+This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and was inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 
-While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
-Outcomes of the simulation study suggested that samples sizes of \(\geq500\) would be needed to estimate \(P_{Max}\) with \textless10\% MARE for a triennially reproducing species.
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but also identified situations where it may be preferable to fix \(P_{Max}\).
+While the estimation of maternity parameters in most simulations was often possible, it was difficult to do so accurately.
+Outcomes of the simulation study showed that, for non-annual reproductive cycles, \(\geq100\) and \(\geq200\) maternal females were typically needed to estimate \(P_{Max}\) with \textless10\% MARE for school and gummy sharks, respectively (Figure 5).
 This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al.
@@
 In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\).
-Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000.
+Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, of which 206 were in maternal condition.
 A similar sample size was used by Colonello \emph{et al}.
 (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.
+
+Providing clear guidelines on when to fix or estimate \(P_{Max}\) is difficult as model performance was species-specific and influenced by life history traits and gear selectivity.
+Most instances where the 3PLF-fixed method outperformed the 3PLF-estimated method typically occurred at lower sample sizes.
+At sample sizes \textless100 it is therefore advisable to fix \(P_{Max}\).
+It is also important to note that accurate estimation of the \(P_{Max}\) parameter required a sample size of 100 - 200 females in maternal condition in simulated data, and that larger sample sizes are needed with longer reproductive cycles.
+
 These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
 2010, Tribuzio and Kruse 2012).
-
-In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution
-
-For some species, and data on pregnant females are particularly sparse, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+For smaller samples sizes, placing a prior or penalty on \(P_{Max}\) may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species (Smart and Grammer 2021).
+Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with \(L_{50}\) effectively providing a lower bound for \(L'_{50}\).
+
+For the numerous data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 
-Another approach to overcoming the
-
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\) from data, but also identified situations where it may be preferable to fix \(P_{Max}\).
-In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method.
-At sample sizes below this it is therefore advisable to fix \(P_{Max}\).
-Larger sample sizes are necessary before attempting to estimate \(P_{Max}\) in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as \(P_{Max}\) decreases.
-
-The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for \(P_{Max}\).
-In this study, the value for \(P_{Max}\) used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data.
-The potential to introduce a greater level of bias in the analysis through the incorrect selection of \(P_{Max}\) should also be considered if using this approach.
-
 \subsection{Empirical study}\label{empirical-study}
@@
 The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013).
+As such, the modal reproductive frequency is still at least triennial.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently.
@@
 
+\bibitem[\citeproctext]{ref-10.1371ux2fjournal.pone.0246734}
+Smart, J.J., and Grammer, G.L.
+2021.
+Modernising fish and shark growth curves with bayesian length-at-age models.
+Plos One \textbf{16}(2): 1--21.
+doi: \href{https://doi.org/10.1371/journal.pone.0246734}{10.1371/journal.pone.0246734}.
+
 \bibitem[\citeproctext]{ref-sminkey_age_1995}

### 2024-05-31 code/extra-plots.R

@@ -1,6 +1,6 @@
 data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
   #filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  #filter(method_long %in% c("3PLF - estimated")) %>% 
+  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
   mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
   mutate(n_maternal_cut = cut(n_maternal, breaks = seq(0, 1600, 20))) %>% 
   filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
@@ -8,7 +8,7 @@ data <- read_rds(here("data", "simulations-summarised.Rds")) %>%
 
 data %>% 
   ggplot() + 
-  geom_tile(aes(x = as.factor(n_maternal_cut), y = repro_freq, fill = method_long)) + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
   #scale_fill_brewer(palette = "Spectral") + 
   scale_fill_viridis_d() + 
   guides(fill = guide_legend(title = "Method")) + 

### 2024-05-31 rmds/figures-supp.Rmd
revisions

@@
 ```
-Figure S1. Convergance success of alternative logistic regression models fit (n = 300) for school shark and gummy shark.
+Figure S1.
+Convergance success of alternative logistic regression models fit (n = 300) for school shark and gummy shark.
+Point size denotes mean number of females in maternal condition at a given sample size.
 
@@
 ```
-Figure S2. Bias (per cent relative error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
+Figure S2.
+Bias (per cent relative error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
 
@@
 Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical.
+Point size denotes mean number of females in maternal condition at a given sample size.
 
@@
 ```
-Figure S4. Accuracy (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.
+Figure S4.
+Accuracy (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models.
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
 
@@
 Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical.
+Point size denotes mean number of females in maternal condition at a given sample size.
 \newpage
@@
 Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
+Point size denotes mean number of females in maternal condition at a given sample size.
 \newpage
@@
 Bias (per cent relative error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
-Each point reflects a mean value from 300 simulated data sets
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
+
+```{r}
+read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs(rel_error)) %>%
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_brewer(palette = "Spectral") + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
+```
+
+Figure S11.
+Performance of alternative maternity functions in minimising bias in calculations of R0.
+The preferred method was that which minimised bias, |relative error| across 300 simulated datasets.
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
 
@@
 ```
-Figure S11.
+Figure S12.
 Accuracy (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
-Each point reflects a mean value from 300 simulated data sets
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
 
@@
 
-Figure S12.
+Figure S13.
 Uncropped version of Figure 2.

### 2024-05-31 rmds/figures.Rmd
revisions

@@
 Simulations with longer reproductive cycles and fewer maternal females were associated with higher bias in both parameters $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
-Note: 42 data points were cropped to aid with data visualization (see Figure S12 for uncropped figure).\
+Note: 42 data points were cropped to aid with data visualization (see Figure S13 for uncropped figure).\
 \
@@
 ```{r fig.width=7, fig.height= 4}
-(p <- ggplot(data = data %>% 
-  filter(method_long %in% c("3PLF - estimated"), par_name %in% "c"), aes(x = n_maternal, y = abs_error, col = repro_freq)) + 
-  geom_point() + 
+
+data %>% 
+  filter(method_long %in% c("3PLF - estimated"), par_name %in% "c") %>% 
+  ggplot() + 
+  geom_point(aes(x = n_maternal, y = abs_error, col = repro_freq)) +  
   scale_color_brewer(palette = "Spectral") +
@@
   theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") +
-  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center"))
+  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
 ```

### 2024-05-31 rmds/manuscript.Rmd
revisions

@@
 (Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure S11).
 Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10).
-For both species, better accuracy in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For both species, better accuracy in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S12).
 For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$.
@@
 Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that accuracy, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated.
-Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes.
-Overall, however, relatively large sample sizes were needed to accurately estimated $P_{Max}$, particularly in species with longer reproductive cycles.
+Using a fixed value for $P_{Max}$ in some cases reduced bias.
+Overall, however, relatively large sample sizes were needed to accurately estimate $P_{Max}$, particularly with longer reproductive cycles.
 This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
-We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.
+We conclude by discussing the advantages of adopting this approach and future directions.
 
@@
 For non-annual reproductive cycles, $> 100$ school sharks and $> 200$ gummy sharks in maternal condition were typically needed to estimate $P_{Max}$ with \<10% MARE (Figure 5).
-This indicates that the both sample sizes and levels of uncertainty are likely to be larger than conventional logistic maturity analysis [@roa_estimation_1999].
-In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$.
-Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition.
+This indicates that the both sample size requirements and levels of uncertainty are likely to be larger than conventional logistic maturity analysis [@roa_estimation_1999].
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in $R_0$ (Figure S11) Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition.
 A similar sample size was used by Colonello *et al*.
@@
 
-For the numerous data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
+For data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
-The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size [@hale_age_2013].
-As such, the modal reproductive frequency is still at least triennial.
+The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially.
+Given that few females appear to attain this size [@hale_age_2013].
+this suggests the modal reproductive frequency likely still at least triennial.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
@@
 For example, due to uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, *Urolophus paucimaculatus*, Trinnie *et al*.
-[@trinnie2014] estimated maternity parameters for ten plausible scenarios.
+[-@trinnie2014] estimated maternity parameters for ten plausible scenarios.
 In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle [@sedar_sedar_2017].

### 2024-05-31 rmds/manuscript.tex
revisions

@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
+However, sample sizes of 100 - 200 maternal females were required for accurate estimation of \(P_{Max}\).
 Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
@@
 In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
-This study suggests that for sample sizes \textgreater{} 100, estimation of \(P_{Max}\) is generally preferable to using a fixed value.
-
-\begin{itemize}
-\tightlist
-\item
-  A sentence with caveat about the need for large sample sizes to produce ``feasible'' results is warranted.
-\end{itemize}
 
@@
 (Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure S11).
 Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10).
-For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S12).
 For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).
@@
 Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated.
-Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes.
-Overall, however, relatively large sample sizes were needed to accurately estimated \(P_{Max}\), particularly in species with longer reproductive cycles.
-This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and was inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
+Using a fixed value for \(P_{Max}\) in some cases reduced bias.
+Overall, however, relatively large sample sizes were needed to accurately estimate \(P_{Max}\), particularly with longer reproductive cycles.
+This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
-We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.
+We conclude by discussing the advantages of adopting this approach and future directions.
 
@@
 
-Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but also identified situations where it may be preferable to fix \(P_{Max}\).
-While the estimation of maternity parameters in most simulations was often possible, it was difficult to do so accurately.
-Outcomes of the simulation study showed that, for non-annual reproductive cycles, \(\geq100\) and \(\geq200\) maternal females were typically needed to estimate \(P_{Max}\) with \textless10\% MARE for school and gummy sharks, respectively (Figure 5).
-This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al.
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but was difficult to do so accurately.
+For non-annual reproductive cycles, \(> 100\) school sharks and \(> 200\) gummy sharks in maternal condition were typically needed to estimate \(P_{Max}\) with \textless10\% MARE (Figure 5).
+This indicates that the both sample size requirements and levels of uncertainty are likely to be larger than conventional logistic maturity analysis (Roa et al.
 1999).
-In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\).
-Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, of which 206 were in maternal condition.
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in \(R_0\) (Figure S11) Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, of which 206 were in maternal condition.
 A similar sample size was used by Colonello \emph{et al}.
@@
 
-Providing clear guidelines on when to fix or estimate \(P_{Max}\) is difficult as model performance was species-specific and influenced by life history traits and gear selectivity.
-Most instances where the 3PLF-fixed method outperformed the 3PLF-estimated method typically occurred at lower sample sizes.
-At sample sizes \textless100 it is therefore advisable to fix \(P_{Max}\).
-It is also important to note that accurate estimation of the \(P_{Max}\) parameter required a sample size of 100 - 200 females in maternal condition in simulated data, and that larger sample sizes are needed with longer reproductive cycles.
-
-These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
+The simulation study also also identified some situations where it may be preferable to fix \(P_{Max}\), particularly at lower sample sizes (\(\le\) 100 individuals total).
+Providing clear guidelines on this is difficult as model performance was species-specific and influenced by life history and gear selectivity.
+The success of any analysis using a fixed value also ultimately depends on the analyst choosing a suitable value.
+The potential to introduce a greater level of bias through the incorrect selection of \(P_{Max}\) should also be considered if using this approach (Natanson et al.
+2019).
+
+Together, these results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
 2010, Tribuzio and Kruse 2012).
@@
 With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
-The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013).
-As such, the modal reproductive frequency is still at least triennial.
+The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially.
+Given that few females appear to attain this size (Hale and Baremore 2013).
+this suggests the modal reproductive frequency likely still at least triennial.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently.
@@
 For example, due to uncertainty in temporal and spatial frequency of reproduction in sparsely-spotted stingarees, \emph{Urolophus paucimaculatus}, Trinnie \emph{et al}.
-(Trinnie et al.
-2014) estimated maternity parameters for ten plausible scenarios.
+(2014) estimated maternity parameters for ten plausible scenarios.
 In stock assessment of western North Atlantic sandbar sharks a breeding frequency of 2.5 years was assumed to account for uncertainty in the duration of the reproductive cycle (SEDAR 2017).
@@
 
+\bibitem[\citeproctext]{ref-natanson_presence_2019}
+Natanson, L.J., Deacy, B.M., Joyce, W., and Sulikowski, J.
+2019.
+Presence of a resting population of female porbeagles ({Lamna} nasus), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
+Fish.
+B-noaa.
+\textbf{117}.
+
 \bibitem[\citeproctext]{ref-nosal2021}

### 2024-06-02 README.qmd
qmd

@@
 ---
-
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
 
@@
 
+theme_set(theme_bw())
+
 data <- read_rds(here::here("data", "empirical-plumbeus.Rds"))
+
+data %>% 
+  ggplot() + 
+  geom_histogram(aes(x = x, fill = as.factor(z)), binwidth = 5, col = "black") +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition") + 
+  guides(fill = guide_legend(title = "Maternal stage")) + 
+  theme(legend.position = "top")
+
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-brks = seq(40, 220, 7.5)
+brks = seq(40, 220, 10)
 data_binned <- data %>%
@@
   summarise(p = sum(z)/ n())
+
+(p <- ggplot()  + 
+  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition"))
 ```

### 2024-06-02 rmds/figures-supp.Rmd
qmd

--- 2024-05-27 rmds/figures-supp.Rmd
@@
 Bias (per cent relative error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models.
-Each point reflects a mean value from 300 simulated data sets
+Each point reflects a mean value from 300 simulated data sets.
+Point size denotes mean number of females in maternal condition at a given sample size.
+
+\newpage
+
+```{r fig.width = 10, fig.height= 4.5}
+read_rds(here("data", "simulations-summarised.Rds")) %>% 
+  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
+  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
+  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
+  slice_min(abs(rel_error)) %>%
+  ggplot() + 
+  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
+  scale_fill_brewer(palette = "Spectral") + 
+  guides(fill = guide_legend(title = "Method")) + 
+  facet_wrap(~species, nrow = 1) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
+```
+
+Figure S11.
+Performance of alternative maternity functions in minimising bias in calculations of R0.
+The preferred method was that which minimised bias, |relative error| across 300 simulated datasets.
+Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
+\newpage
 

### 2024-06-02 rmds/figures.Rmd
qmd

@@
 Each point represents parameter estimates from one iteration of simulated data (n = 43,129), including all combinations of variables.
-Simulations with longer reproductive cycles and fewer maternal females were associated with higher bias in both parameters $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
+Simulations with longer reproductive cycles and fewer maternal females were associated with higher bias in both $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
 Note: 42 data points were cropped to aid with data visualization (see Figure S13 for uncropped figure).\

### 2024-06-02 rmds/manuscript.Rmd
qmd

@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-However, sample sizes of 100 - 200 maternal females were required for accurate estimation of $P_{Max}$.
-Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
+However, sample sizes of 100 - 200 maternal females were typically required for accurate estimation of $P_{Max}$.
+While maturity parameters could be estimated with greater accuracy, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
@@
 As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$.
-Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).
+Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).
 
@@
 (Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure S11).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in many instances (Figure S11).
 Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10).
@@
 Overall, however, relatively large sample sizes were needed to accurately estimate $P_{Max}$, particularly with longer reproductive cycles.
-This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
+This study also demonstrated that, in some cases, recruitment was overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 This indicates that the both sample size requirements and levels of uncertainty are likely to be larger than conventional logistic maturity analysis [@roa_estimation_1999].
-In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in $R_0$ (Figure S11) Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition.
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in $R_0$ (Figure S11).
+Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, 206 of which were in maternal condition.
 A similar sample size was used by Colonello *et al*.
@@
 $L_{5}$ and $L_{95}$ are 146cm and 176cm – a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013].
-With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
-The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially.
-Given that few females appear to attain this size [@hale_age_2013].
-this suggests the modal reproductive frequency likely still at least triennial.
+With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, while larger sharks may ultimately be capable of reproducing biennially.
+Given that few females appear to attain this size [@hale_age_2013], the modal reproductive frequency is still likely at least triennial within the population.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently.
@@
 Estimation of $P_{Max}$ using the 3PLF-estimated method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans.
-White sharks, *Carcharodon carcharias*, and whale shark, *Rhinchodon typus*, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
+White sharks, *Carcharodon carcharias*, and whale shark, *Rhincodon typus*, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist [@joung_whale_1996; @sato_how_2016].
 In both species lack of data on reproductive frequency is a major impediment to population modelling and the development of management strategies [@huveneers_future_2018; @bowlby_implications_2020].

### 2024-06-02 code/empirical.R

@@ -82,14 +82,14 @@ ggplot() +
   geom_point(data = filter(data_complete2, Stage %in% c(4, 5)), aes(x = month, y = puplen, col = as.factor(Stage))) +
   scale_x_continuous(breaks = seq(0, 12, 2))
 
-data2 <- select(data_complete2, month, Stage, x, z) %>%
+data2 <- select(data_complete2, month, Stage, x, z) 
   filter(!is.na(x), !is.na(z))
 
 # Joing Ivy and Andrew's data
 
 # Tidy up data 
 data <- data %>% 
-  ungroup() %>% select(month, Stage, x, z) %>%
+  ungroup() %>% select(month, Stage, x, z)
   rbind(data2) %>% 
   mutate(y = ifelse(Stage > 2, 1, 0)) %>% 
   select(x, y, z) %>% 

### 2024-06-23 README.qmd
changes

@@
 
-## Example
+# Example
+
+## Load data
 
@@
 
-## Bin data into suitable length categories for plotting
+Bin data into suitable length categories for plotting.
 
@@
 
-(p <- ggplot()  + 
+ggplot()  + 
   geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
   ylim(0, 1) +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition"))
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
 ```
+
+Add marginal rug plots to show individual data points.
+
+```{r}
+ggplot()  + 
+  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
+  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
+  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
+
+```
+
+## Run model
 

### 2024-06-23 reports/revision/figs.tex
changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\subsection{Figures}\label{figures}

\includegraphics[width=1\linewidth]{figure1.png}

Figure 1. Approach used to generate simulated data and test the
performance of four methods for calculating maternity parameters.
Illustrations © R.Swainston/www.anima.net.au

\includegraphics{unnamed-chunk-2-1.pdf}

Figure 2. Bias (per cent relative error) in parameter estimates for
\(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) for the 3PLF maternity function
with \(P_{Max}\) estimated. Each point represents parameter estimates
from one iteration of simulated data (n = 43,129), including all
combinations of variables. Simulations with longer reproductive cycles
and fewer maternal females were associated with higher bias in both
\(\hat{L'_{50}}\) and \(\hat{P_{Max}}\). Note: 42 data points were
cropped to aid with data visualization (see Figure S13 for uncropped
figure).\\
\strut \\
\newpage

\includegraphics{unnamed-chunk-3-1.pdf}

Figure 3. Accuracy (per cent absolute error) in parameter estimates of
\(\hat{P_{Max}}\) for 3PLF methods with high maternal selectivity. Large
sample sizes were needed to accurately estimate \(\hat{P_{Max}}\) and
accuracy decreased as the duration of the reproductive cycle increased.
Each point reflects a mean value from 300 simulated data sets. Point
size denotes mean number of females in maternal condition at a given
sample size.\\
\strut \\
\newpage

\includegraphics{unnamed-chunk-4-1.pdf}

Figure 4. Confidence interval coverage for \(\hat{L'_{50}}\) for 3PLF
methods with high maternal selectivity. Figure shows the percentage of
simulations (n = 300) where the true parameter value fell within the
50\% bootstrap confidence interval. Point size denotes mean number of
females in maternal condition at a given sample size.\\
\strut \\
\newpage

\includegraphics{unnamed-chunk-5-1.pdf}

Figure 5. Accuracy (per cent absolute error) in parameter estimates of
\(\hat{P_{Max}}\) for the 3PLF - estimated method as a function of
number of maternal females. Each point reflects a value from 300
simulated data sets.\\
\newpage

\includegraphics{unnamed-chunk-6-1.pdf}

Figure 6. Performance of four alternative maternity functions in
accurately calculating \(R_0\). The best performing method was that
which minimised mean absolute error across 300 simulated datasets. Note
2PLF-maternity (Annual) scenarios were excluded for this comparison.\\
\newpage

\includegraphics{unnamed-chunk-7-1.pdf}

Figure 7. Comparison of 3PLF-estimated and 3PLF-fixed methods used to
estimate maternal parameters for sandbar shark, \emph{C. plumbeus}, in
the Gulf of Mexico and Western North Atlantic. Solid line is the
expected proportion in maternal condition at length, \(\Psi'(L)\). The
grey shaded region denotes 95\% confidence intervals based on bootstrap
resampling. Black points show proportion in maternal condition in 10cm
length intervals. Marginal rug plots denote raw data that models were
fit to. \(P_{Max}\) was fixed at 0.5 in the lower panel.

\end{document}

### 2024-06-23 reports/revision/main-text.tex
changes

--- 2024-05-31 rmds/manuscript.tex
@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-However, sample sizes of 100 - 200 maternal females were required for accurate estimation of \(P_{Max}\).
-Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
+However, sample sizes of 100 - 200 maternal females were typically required for accurate estimation of \(P_{Max}\).
+While maturity parameters could be estimated with greater accuracy, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
@@
 As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\).
-Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).
+Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).
 
@@
 (Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure S11).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in many instances (Figure S11).
 Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10).
@@
 Overall, however, relatively large sample sizes were needed to accurately estimate \(P_{Max}\), particularly with longer reproductive cycles.
-This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
+This study also demonstrated that, in some cases, recruitment was overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
@@
 1999).
-In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in \(R_0\) (Figure S11) Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, of which 206 were in maternal condition.
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in \(R_0\) (Figure S11).
+Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, 206 of which were in maternal condition.
 A similar sample size was used by Colonello \emph{et al}.
@@
 \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013).
-With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
-The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially.
-Given that few females appear to attain this size (Hale and Baremore 2013).
-this suggests the modal reproductive frequency likely still at least triennial.
+With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, while larger sharks may ultimately be capable of reproducing biennially.
+Given that few females appear to attain this size (Hale and Baremore 2013), the modal reproductive frequency is still likely at least triennial within the population.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently.
@@
 Estimation of \(P_{Max}\) using the 3PLF-estimated method may also be a means to resolving the longstanding question of reproductive frequency in some rare and threatened chondrichthyans.
-White sharks, \emph{Carcharodon carcharias}, and whale shark, \emph{Rhinchodon typus}, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist (Joung et al.
+White sharks, \emph{Carcharodon carcharias}, and whale shark, \emph{Rhincodon typus}, are two examples of intensively studied species where only fragmentary observations of female reproductive biology exist (Joung et al.
 1996, Sato et al.
@@
 2019.
-Presence of a resting population of female porbeagles ({Lamna} nasus), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
+Presence of a resting population of female porbeagles (\emph{{Lamna} nasus}), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
 Fish.

### 2024-06-23 rmds/manuscript.Rmd
changes

@@
 $n = 1$), $Y_i$ reduces to a Bernoulli random variable [@zuur_analyzing_2007].
-$\alpha$ is the upper asymptote of the curve, controlling the maximum value of $Y$ as $x$ approaches infinity.
+The upper asymptote of the curve, $\alpha$, controls the maximum value of $Y$ as $x$ approaches infinity.
 The lesser-used 3PLF reduces to the ubiquitous 2PLF when $\alpha = 1$, as is appropriate for most teleost and chondrichthyan populations if modeling maturity [@roa_estimation_1999; @quinn_quantitative_1999].
@@
 $$
-\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
+C_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
 $$
@@
 When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased.
-This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby true underlying $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2).
+This is seen in the positive correlation between $\widehat{L'_{50}}$ and $\widehat{P_{Max}}$, whereby true underlying $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2).
 With lower values of $P_{Max}$ it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$.
+As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\widehat{P_{Max}}$.
 Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).
@@
 While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6).
-For most simulations, interval coverage for $\hat{L'_{50}}$ was well below the expected level of 50%, irrespective of sample size.
-In contrast, for the 3PLF-estimated method interval coverage oscillated around 50% for both $\hat{L'_{50}}$ and $\hat{P_{Max}}$.
+For most simulations, interval coverage for $\widehat{L'_{50}}$ was well below the expected level of 50%, irrespective of sample size.
+In contrast, for the 3PLF-estimated method interval coverage oscillated around 50% for both $\widehat{L'_{50}}$ and $\widehat{P_{Max}}$.
 
@@
 Maternal data were re-analysed for `r nrow(plumbeus)` sandbar sharks including `r length(which(plumbeus$y == 1))` mature individuals of which `r round(100 * (length(which(plumbeus$z == 1)) / length(which(plumbeus$y == 1))))`% were in maternal condition.
-Using the 3PLF-estimated method, the maximum likelihood estimate for $\hat{P_{Max}}$ was 0.48 (Table 2).
-Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\hat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence (Figure 7).
+Using the 3PLF-estimated method, the maximum likelihood estimate for $\widehat{P_{Max}}$ was 0.48 (Table 2).
+Despite having a sample size of \> 600 mature females, the proportion of maternal individuals at length was still uncertain and $\widehat{P_{Max}}$ was estimated to lie between 0.39 and 0.60 with 95% confidence (Figure 7).
 For 3PLF-fixed models a value of $P_{Max} = 0.5$ was strongly supported ($\Delta_{AIC}$ = 14.15) over the alternative fixed value of $P_{Max} = 0.33$.
@@
 In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
-Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of $\hat{P_{Max}} = 0.48$, considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale [-@baremore_reproduction_2012].
+Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of $\widehat{P_{Max}} = 0.48$, considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale [-@baremore_reproduction_2012].
 Similarly, based on AIC values, a fixed value of $P_{Max} = 0.5$ had a much higher relative plausibility than a value of 0.33 (Table 2).

### 2024-06-23 rmds/tables.Rmd
changes

@@
   pdf_document:
+    keep_tex: true
     latex_engine: "pdflatex"
@@
 tables: true
+
 ---

### 2024-06-23 reports/figs.tex
repo_prep

--- 2024-06-23 reports/revision/figs.tex
@@
 
-\includegraphics[width=1\linewidth]{figure1.png}
+\includegraphics[width=1\linewidth]{../img/figure1}
 
@@
 
-\includegraphics{unnamed-chunk-2-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-3-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-3-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-4-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-4-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-5-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-5-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-6-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-7-1.pdf}
+\includegraphics{/Users/avh/Library/CloudStorage/OneDrive-DepartmentofPrimaryIndustriesAndRegionalDevelopment/manuscripts/maternity-function/reports/figs_files/figure-latex/unnamed-chunk-7-1.pdf}
 

### 2024-06-23 reports/main-text.tex
repo_prep

@@
 Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used.
-Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output.
+However, sample sizes of 100 - 200 maternal females were typically required for accurate estimation of \(P_{Max}\).
+While maturity parameters could be estimated with greater accuracy, substituting them for maternity parameters overestimated lifetime reproductive output.
 Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models.
 In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.
-This study suggests that for sample sizes \textgreater{} 100, estimation of \(P_{Max}\) is generally preferable to using a fixed value.
 
@@
 2007).
-\(\alpha\) is the upper asymptote of the curve, controlling the maximum value of \(Y\) as \(x\) approaches infinity.
+The upper asymptote of the curve, \(\alpha\), controls the maximum value of \(Y\) as \(x\) approaches infinity.
 The lesser-used 3PLF reduces to the ubiquitous 2PLF when \(\alpha = 1\), as is appropriate for most teleost and chondrichthyan populations if modeling maturity (Quinn and Deriso 1999, Roa et al.
@@
 Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions (Cortés et al.
-2012).
+2012, Swenson et al.
+2024).
 It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles.
 Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.~annual, biennial), despite observations of plasticity in this trait within discrete populations (Higgs et al.
-2020).
+2020, Nosal et al.
+2021).
 Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data.
-Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity.
+Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles.
 To date only a single study on spiny dogfish, \emph{Squalus acanthias}, has statistically estimated \(P_{Max}\) from data (Colonello et al.
@@
 \[
-\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
+C_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
 \]
@@
 Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1.
-In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.
-
-Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased.
-The nature of this problem can be seen in the positive correlation between the \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) parameters whereby \(P_{Max}\) affects the magnitude and direction of bias in these parameters (Figure 2).
-With lower values of \(P_{Max}\) it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\) (Figure 2).
+In addition to convergence failures, 12 simulations also generated zero maternal females, and were discarded.
+
+When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased.
+This is seen in the positive correlation between \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2).
+With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
+As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\widehat{P_{Max}}\).
+Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).
 
@@
 While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6).
-For most simulations, interval coverage for \(\hat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size.
-In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\).
-
-Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters.
-Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9).
+For most simulations, interval coverage for \(\widehat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size.
+In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\).
+
+Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters.
+Certain combinations of variables in the simulated data resulted in few females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9).
+This was particularly the case for \(P_{Max}\) which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5).
 These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure.
-For example, in low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).
-As \(P_{Max}\) decreased the proportion of maternal females in the analysis also decreased.
+For example, in biennial low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).
 
@@
 The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios.
-For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10).
+For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure S10).
 Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%.
-(Table S1, Figure S10).
-In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10).
+(Figure S10).
+In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in many instances (Figure S11).
 Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10).
-For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11).
+For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S12).
 For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).
@@
 
-Maternal data were re-analysed for 1087 sandbar sharks including 0 mature individuals of which \ensuremath{\infty{}}\% were in maternal condition.
-Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2).
-Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence.
+Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition.
+Using the 3PLF-estimated method, the maximum likelihood estimate for \(\widehat{P_{Max}}\) was 0.48 (Table 2).
+Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\widehat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7).
 For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\).
 The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86).
-Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data (Burnham and Anderson 2002).
+Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data (Burnham and Anderson 2002).
 
@@
 Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated.
-Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes.
-This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
+Using a fixed value for \(P_{Max}\) in some cases reduced bias.
+Overall, however, relatively large sample sizes were needed to accurately estimate \(P_{Max}\), particularly with longer reproductive cycles.
+This study also demonstrated that, in some cases, recruitment was overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters.
 Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology.
-We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.
+We conclude by discussing the advantages of adopting this approach and future directions.
 
@@
 
-While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately.
-Outcomes of the simulation study suggested that samples sizes of \(\geq500\) would be needed to estimate \(P_{Max}\) with \textless10\% MARE for a triennially reproducing species.
-This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al.
+Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but was difficult to do so accurately.
+For non-annual reproductive cycles, \(> 100\) school sharks and \(> 200\) gummy sharks in maternal condition were typically needed to estimate \(P_{Max}\) with \textless10\% MARE (Figure 5).
+This indicates that the both sample size requirements and levels of uncertainty are likely to be larger than conventional logistic maturity analysis (Roa et al.
 1999).
-In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\).
-Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000.
+In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in \(R_0\) (Figure S11).
+Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, 206 of which were in maternal condition.
 A similar sample size was used by Colonello \emph{et al}.
 (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.
-These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
+
+The simulation study also also identified some situations where it may be preferable to fix \(P_{Max}\), particularly at lower sample sizes (\(\le\) 100 individuals total).
+Providing clear guidelines on this is difficult as model performance was species-specific and influenced by life history and gear selectivity.
+The success of any analysis using a fixed value also ultimately depends on the analyst choosing a suitable value.
+The potential to introduce a greater level of bias through the incorrect selection of \(P_{Max}\) should also be considered if using this approach (Natanson et al.
+2019).
+
+Together, these results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al.
 2010, Tribuzio and Kruse 2012).
-
-In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution
-
-For some species, and data on pregnant females are particularly sparse,
-the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans.
+For smaller samples sizes, placing a prior or penalty on \(P_{Max}\) may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species (Smart and Grammer 2021).
+Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with \(L_{50}\) effectively providing a lower bound for \(L'_{50}\).
+
+For data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option.
 From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging.
@@
 In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level.
-Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\hat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012).
+Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\widehat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012).
 Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).
@@
 \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013).
-With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently.
-The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013).
+With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, while larger sharks may ultimately be capable of reproducing biennially.
+Given that few females appear to attain this size (Hale and Baremore 2013), the modal reproductive frequency is still likely at least triennial within the population.
 Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently.
@@
 Reproduction of the {Sandbar} {Shark} in the {Western} {North} {Atlantic} {Ocean} and {Gulf} of {Mexico}.
-Marine and Coastal Fisheries \textbf{4}(1): 560--572.
+Mar.
+Coast.
+Fish.
+\textbf{4}(1): 560--572.
 doi: \href{https://doi.org/10.1080/19425120.2012.700904}{10.1080/19425120.2012.700904}.
@@
 Reproduction of the {Blacktip} {Shark} in the {Gulf} of {Mexico}.
-Marine and Coastal Fisheries \textbf{5}(1): 127--138.
+Mar.Coast.
+Fish.
+\textbf{5}(1): 127--138.
 doi: \href{https://doi.org/10.1080/19425120.2012.758204}{10.1080/19425120.2012.758204}.
@@
 Implications of life history uncertainty when evaluating status in the {Northwest} {Atlantic} population of white shark (\emph{{Carcharodon} carcharias)}.
-Ecology and Evolution \textbf{10}(11): 4990--5000.
+Eco.
+Evol.
+\textbf{10}(11): 4990--5000.
 doi: \href{https://doi.org/10.1002/ece3.6252}{10.1002/ece3.6252}.
@@
 \bibitem[\citeproctext]{ref-burnham_model_2002}
-Burnham, K.P., and Anderson, D.R. 2002. Model selection and multimodel inference: A practical information-theoretic approach. 2nd edn. Springer, New York, NY.
+Burnham, K.P., and Anderson, D.R. 2002. Model {Selection} and {Multimodel} {Inference}: A {Practical} {Information-theoretic} {Approach}. {Second} {Edition}. Springer, New York, NY.
 
@@
 You can swim but you can't hide: The global status and conservation of oceanic pelagic sharks and rays.
-Aquatic Conservation: Marine and Freshwater Ecosystems.
+Aquat.
+Conserv.
+Mar.
+Freshw.
+Ecosyst.
 
@@
 Life histories and vulnerability to exploitation of elasmobranchs: {Inferences} from elasticity, pertubation and phylogenetic analyses.
-Journal of Northwest Atlantic Fishery Science \textbf{35}: 27--45.
+J.
+Northwest Atl.
+Fish.
+Sci.
+\textbf{35}: 27--45.
 doi: \href{https://doi.org/10.2960/J.v35.m514}{10.2960/J.v35.m514}.
@@
 \textbf{68}(11): 2018--2027.
-Available from \url{https://doi.org/10.1071/MF16101}.
+doi: \href{https://doi.org/10.1071/MF16101}{10.1071/MF16101}.
 
@@
 2013.
-Age and {Growth} of the {Sandbar} {Shark} (\emph{{Carcharhinus} plumbeus}) from the {Northern} {Gulf} of {Mexico} and the {Western} {North} {Atlantic} {Ocean}.
-Gulf of Mexico Science \textbf{1}(2): 28--39.
+Age and {Growth} of the sandbar shark (\emph{{Carcharhinus} plumbeus}) from the {Northern} {Gulf} of {Mexico} and the {Western} {North} {Atlantic} {Ocean}.
+Gulf Mex.
+Sci.
+\textbf{1}(2): 28--39.
 
@@
 Freshwater Res.
-Available from \url{https://doi.org/10.1071/MF20120}.
+doi: \href{https://doi.org/10.1071/MF20120}{10.1071/MF20120}.
 
@@
 Future research directions on the {``{Elusive}''} white shark.
-Frontiers in Marine Science \textbf{5}: 455.
+Front.
+Mar.
+Sci.
+\textbf{5}: 455.
 doi: \href{https://doi.org/10.3389/fmars.2018.00455}{10.3389/fmars.2018.00455}.
@@
 \bibitem[\citeproctext]{ref-jennings_marine_2001}
-Jennings, S., Kaiser, M., and Reynolds, J.D. 2001. Marine fisheries ecology. John Wiley \& Sons, Oxford.
+Jennings, S., Kaiser, M., and Reynolds, J.D. 2001. Marine {Fisheries} {Ecology}. John Wiley \& Sons, Oxford.
 
@@
 \bibitem[\citeproctext]{ref-king_fisheries_2007}
-King, M. 2007. Fisheries biology, assessment and management. Blackwell Publishing, Oxford.
+King, M. 2007. Fisheries {Biology}, {Assessment} and {Management}. Blackwell Publishing, Oxford.
 
@@
 \textbf{160}: 18--32.
-doi: \url{https://doi.org/10.1016/j.fishres.2014.03.003}.
+doi: \href{https://doi.org/10.1016/j.fishres.2014.03.003}{10.1016/j.fishres.2014.03.003}.
 
@@
 1992.
-Age and growth-studies of gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, and school shark, \emph{{Galeorhinus} galeus}({Linnaeus}), from {Southern} {Australian} waters.
+Age and growth-studies of gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, and school shark, \emph{{Galeorhinus} galeus} ({Linnaeus}), from {Southern} {Australian} waters.
 Aust.
@@
 doi: \href{https://doi.org/10.1071/MF9921241}{10.1071/MF9921241}.
+
+\bibitem[\citeproctext]{ref-natanson_presence_2019}
+Natanson, L.J., Deacy, B.M., Joyce, W., and Sulikowski, J.
+2019.
+Presence of a resting population of female porbeagles (\emph{{Lamna} nasus}), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
+Fish.
+Bull.
+\textbf{117}.
+
+\bibitem[\citeproctext]{ref-nosal2021}
+Nosal, A.P., Cartamil, D.P., Ammann, A.J., Bellquist, L.F., Ben-Aderet, N.J., Blincow, K.M., Burns, E.S., Chapman, E.D., Freedman, R.M., Klimley, A.P., Logan, R.K., Lowe, C.G., Semmens, B.X., White, C.F., and Hastings, P.A.
+2021.
+Triennial migration and philopatry in the critically endangered soupfin shark {\emph{Galeorhinus galeus}}.
+J.
+Appl.
+Ecol.
+\textbf{58}(8): 1570--1582.
+doi: \href{https://doi.org/10.1111/1365-2664.13848}{10.1111/1365-2664.13848}.
 
@@
 \textbf{48}(8): 981--992.
-Available from \url{https://doi.org/10.1071/MF97017}.
+doi: \href{https://doi.org/10.1071/MF97017}{10.1071/MF97017}.
 
@@
 \bibitem[\citeproctext]{ref-quinn_quantitative_1999}
-Quinn, I.I., and Deriso, R.B. 1999. Quantitative fish dynamics. Oxford University Press, New York, NY.
+Quinn, I.I., and Deriso, R.B. 1999. Quantitative {Fish} {Dynamics}. Oxford University Press, New York, NY.
 
@@
 Life history, growth, and reproductive biology of four mobulid species in the {Bohol} {Sea}, {Philippines}.
-Frontiers in Marine Science \textbf{5}: 269.
+Front.
+Mar.
+Sci.
+\textbf{5}: 269.
 doi: \href{https://doi.org/10.3389/fmars.2018.00269}{10.3389/fmars.2018.00269}.
@@
 Deep Sea Research Part II: Topical Studies in Oceanography \textbf{115}: 30--40.
-doi: \url{https://doi.org/10.1016/j.dsr2.2013.09.004}.
+doi: \href{https://doi.org/10.1016/j.dsr2.2013.09.004}{10.1016/j.dsr2.2013.09.004}.
 
 \bibitem[\citeproctext]{ref-roa_estimation_1999}
-Roa, R., Ernst, B., and Tapia, F. 1999. Estimation of size at sexual maturity: An evaluation of analytical and resampling procedures. Fish. B-noaa. \textbf{97}: 570--580.
+Roa, R., Ernst, B., and Tapia, F. 1999. Estimation of size at sexual maturity: An evaluation of analytical and resampling procedures. Fish. Bull. \textbf{97}: 570--580.
 
@@
 How great white sharks nourish their embryos to a large size: Evidence of lipid histotrophy in lamnoid shark reproduction.
-Biology Open \textbf{5}(9): 1211--1215.
+Biol.
+Open \textbf{5}(9): 1211--1215.
 doi: \href{https://doi.org/10.1242/bio.017939}{10.1242/bio.017939}.
@@
 \bibitem[\citeproctext]{ref-sedar_sedar_2012}
-SEDAR. 2012. {SEDAR} 29 stock assessment report: {HMS} {Gulf} of {Mexico} blacktip shark. North Charleston, SC.
+SEDAR. 2012. {SEDAR} 29 {Stock} {Assessment} {Report}: {HMS} {Gulf} of {Mexico} {Blacktip} {Shark}. North Charleston, SC.
 
 \bibitem[\citeproctext]{ref-sedar_sedar_2017}
-SEDAR. 2017. {SEDAR} 54 stock assessment report: {HMS} sandbar shark. North Charleston, SC.
+SEDAR. 2017. {SEDAR} 54 {Stock} {Assessment} {Report}: {HMS} {Sandbar} {Shark}. North Charleston, SC.
 
@@
 SEDAR. 2023. {HMS} {Hammerhead} {Sharks}: {Scalloped} {Hammerhead} {Shark}. {Section} {III}: {Assessment} {Report}. North Charleston, SC.
+
+\bibitem[\citeproctext]{ref-10.1371ux2fjournal.pone.0246734}
+Smart, J.J., and Grammer, G.L.
+2021.
+Modernising fish and shark growth curves with {Bayesian} length-at-age models.
+Plos One \textbf{16}(2): 1--21.
+doi: \href{https://doi.org/10.1371/journal.pone.0246734}{10.1371/journal.pone.0246734}.
 
@@
 
+\bibitem[\citeproctext]{ref-swenson2024}
+Swenson, J.D., Brooks, E.N., Kacev, D., Boyd, C., Kinney, M.J., Marcy-Quay, B., Sévêque, A., Feldheim, K.A., and Komoroske, L.M.
+2024.
+Accounting for unobserved population dynamics and aging error in close{-}kin mark{-}recapture assessments.
+Ecol.
+Evol.
+\textbf{14}(2).
+doi: \href{https://doi.org/10.1002/ece3.10854}{10.1002/ece3.10854}.
+
 \bibitem[\citeproctext]{ref-taylor_stockrecruitment_2013}
@@
 \textbf{142}: 15--21.
-doi: \url{https://doi.org/10.1016/j.fishres.2012.04.018}.
+doi: \href{https://doi.org/10.1016/j.fishres.2012.04.018}{10.1016/j.fishres.2012.04.018}.
 
@@
 \textbf{43}(1): 195--212.
-Available from \url{https://doi.org/10.1071/MF9920195}.
+doi: \href{https://doi.org/10.1071/MF9920195}{10.1071/MF9920195}.
 
@@
 \bibitem[\citeproctext]{ref-zuur_analyzing_2007}
-Zuur, A., Ieno, E.N., and Smith, G.M. 2007. Analyzing ecological data. New York, NY.
+Zuur, A., Ieno, E.N., and Smith, G.M. 2007. Analyzing {Ecological} {Data}. New York, NY.
 

### 2024-06-23 code/age-underestimation.R

@@ -1,97 +1,0 @@
-# Housekeeping ----------------------------------------------------------------
-
-source(here::here("code", "setup.R"))
-source(here::here("code", "functions.R"))
-
-compile("code/logistic3.cpp")
-dyn.load(dynlib("code/logistic3"))
-compile("code/logistic2.cpp")
-dyn.load(dynlib("code/logistic2"))
-compile("code/logistic2maternal.cpp")
-dyn.load(dynlib("code/logistic2maternal"))
-
-
-sim_pars_galeus <- list(
-  species = "School shark",
-  
-  # Maximum age
-  AMAX = c(27, 54), 
-  # Natural mortality (Grant et al 1979)
-  M = 0.1075,
-  # Fishing mortality (Nominally set to 25 % of M)
-  F_mort = 0.25 * 0.1075,
-  # von Bertalanffy growth parameters (Grant et al 1979)
-  K = 0.1600,
-  Linf = 1618.3,
-  t0 = -1.2818,
-  cv_l = 0.05,
-  # Maternity parameters (Walker 2005)
-  m50 = 1421,
-  m95 = 1488,
-  c = c(1 / 3), # Proportion maternal
-  # Maturity parameters (Walker 2005)
-  l50 = 1349,
-  l95 = 1502,
-  # Uterine fecundity as function of length (Walker 2005)
-  f = -46.0,
-  g = 0.0491,
-  # Selectivity parameters (Punt and Walker 1998)
-  # Correspond to peak sel at 25% and 75% length maternity
-  mesh = c(7.53),
-  theta1 = 192,
-  theta2 = 67595,
-  
-  # Number of samples
-  Nsamples = c(100, 1000, 2500)
-)
-
-data <-
-  # Generate data frame of all combinations of scenarios
-  cross_df(sim_pars_galeus) %>%
-  # Create some additional variables
-  mutate(mesh_name = case_when(
-    mesh == 5.68 | mesh == 7.27 ~ "low",
-    mesh == 6.57 | mesh == 7.53 ~ "high"
-  )) %>%
-  mutate(repro_freq = case_when(
-    c == 1 ~ "1 (Annual)",
-    c == 3 / 4 ~ "0.75 (Annual / biennial)",
-    c == 2 / 3 ~ "0.66 (Annual / biennial)",
-    c == 0.5 ~ "0.5 (Biennial)",
-    c == 1 / 3 ~ "0.33 (Triennial)",
-    c == 0.25 ~ "0.25 (Quadrennial)"
-  )) %>%
[…]

### 2024-06-23 code/empirical.R

@@ -8,92 +8,16 @@ source(here::here("code", "functions.R"))
 compile("code/logistic3.cpp")
 dyn.load(dynlib("code/logistic3"))
 
-# Load data
-# Complete dataset from Ivy, includes some smaller females that were missing from previous spreadsheet
-data_complete <- readxl::read_xlsx(here("data", "plumbeus-unprocessed.xlsx"))
-
-# Incomplete dataset that has full repro details
-data_incomplete <- readxl::read_xls(here("data", "plumbeus-extra-unprocessed.xls"), sheet = 6) %>%
-  select(
-    ID = `SBAR .`, `OVARY L (mm)`, `OVARY WIDTH`, `OVARY WEIGHT (g)`, `TOTAL OOCYTES`, `. VIT`,
-    `DIAMETER LARGEST OO`, `UTERUS WIDTH`, `NIDAMENTAL WIDTH`, `PUPS?`, `. LEFT`, `. RIGHT`
-  ) %>%
-  arrange(ID) %>%
-  # Remove #358 as it is a duplicate, doesn't seem to have been included in final analysis
-  filter(!row_number() %in% 281)
-
-# Load data for pups
-pups <- read_csv(here("data", "plumbeus-pups-unprocessed.csv")) %>% 
-  rename_all(tolower) %>%
-  select(ID = maternal.id, puplen = `stl.cm`) %>% 
-  mutate(puplen = as.numeric(puplen))
-
-# Treat July as beginning of a new year / season
-
-# SBAR1242 stage 4 just ovulated and pregnant, but June so count as resting for this year
-# SBAR1248 stage 4 just ovulated and pregnant, but June so count as resting for this year
-# SBAR1244 stage 4 just ovulated and pregnant, but June so count as resting for this year
-# SBAR471 non-pregnant, perhaps pupped late?
-# SBAR353 stage 6 postpartum but month is September so would have been maternal last season
-# SBAR303 possible stage 6 but definitely not in maternal condition
-# SBAR376 stage 6 postpartum but month is September so would have been maternal last season
-# SBAR219 stage 6 postpartum but month is August so would have been maternal last season
-# SBAR478 stage 6 postpartum but month is November so would have been maternal last season
-
-data <- left_join(data_complete, data_incomplete) %>%
[…]

### 2024-06-23 code/extra-plots.R

@@ -1,71 +1,0 @@
-data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
-  #filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
-  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
-  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
-  mutate(n_maternal_cut = cut(n_maternal, breaks = seq(0, 1600, 20))) %>% 
-  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
-  slice_min(abs_error)
-
-data %>% 
-  ggplot() + 
-  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
-  #scale_fill_brewer(palette = "Spectral") + 
-  scale_fill_viridis_d() + 
-  guides(fill = guide_legend(title = "Method")) + 
-  facet_wrap(~species, nrow = 1) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90)) + 
-  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
-
-
-
-(p <- ggplot(data = data, aes(x = n_maternal, y = as.numeric(method_long))) + 
-  geom_point() + 
-  geom_smooth() + 
-  scale_x_continuous(breaks = seq(0,1600,200)) + 
-    facet_wrap(~species))
-
-
-(p <- ggplot(data = data, aes(x = n_maternal, y = abs_error, col = repro_freq)) + 
-  geom_point() + 
-  scale_color_brewer(palette = "Spectral") +
-  scale_y_continuous(breaks = seq(0, 90, 10)) +
-  facet_grid( ~ species) + 
-  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
-  theme_bw() + 
-  labs(x = "Median number of maternal females", y = "Mean % absolute relative error") +
-  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") +
-  theme(legend.box = "vertical", legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center"))
-
-
[…]

### 2024-06-23 code/functions.R

@@ -17,101 +17,7 @@ save_sim_data <- function(df, id) {
   write_rds(df, paste0(dir_out, "/population-structure.rds"))
 }
 
-## The continuous time Lotka equation (Xiao and Walker 2000)
-continuous_lotka <- function(a, Lambda, df) {
 
-  # Length at age function
-  len <- function(a) {
-    df$Linf * (1 - exp(-df$K * (a - df$t0)))
-  }
-
-  # Maternity at age function
-  maternity <- function(a) {
-    (df$c / (1 + exp(-log(19) * ((len(a) - df$m50) / (df$m95 - df$m50)))))
-  }
-
-  # Fecundity at age function
-  fecundity <- function(a) {
-    if (df$species == "School shark") {
-      df$f + df$g * len(a)
-    } else {
-      df$f * exp(df$g * len(a))
-    }
-  }
-
-  # Birth function
-  birth <- function(a) {
-    if (df$species == "School shark") {
-      ifelse(a >= 11.2404, fecundity(a) * maternity(a) * (1 / 2), 0)
-    } else {
-      ifelse(a >= 3.9694, fecundity(a) * maternity(a) * (1 / 2), 0)
-    }
-  }
-
-  # Intrinsic rate of population decrease with age
-  LAMBDA <- function(a) {
-    rep(Lambda, length(a))
-  }
-
-  # Sum to optimize
-  birth(a) * exp(sapply(a, function(x) {
-    integrate(LAMBDA, 0, x)$value
-  }))
-}
-
-## Function for solving Lotka equation
-solve_lotka <- function(df, Lambda = 0) {
-  Integral <- integrate(continuous_lotka, 0, Inf, Lambda = Lambda, df = df)$value
-  dif <- abs(1 - Integral)
-  return(dif)
-}
-
-## Function for optimizing Lotka equation
-optimize_lotka <- function(df, lower = -2, upper = 0) {
-  lambda <- optimize(solve_lotka, c(lower, upper), df = df)$minimum
-  return(list(lambda = lambda))
-}
-
-## Function for generating simulated length and age data
-generate_data <- function(df, id, Return = FALSE) {
-  df <- df %>%
-    mutate(lambda = optimize_lotka(df)$lambda)
-
-  # Convert selectivity parameters
-  df <- df %>%
-    mutate(eta = -0.5 * (theta1 * mesh - sqrt(theta1^2 * mesh^2 + 4 * theta2))) %>%
[…]

### 2024-06-23 code/selectivity.R

@@ -1,166 +1,0 @@
-
-
-sel_pars_gummy <- list(mesh = seq(4, 9, 0.01),
-                 
-                 theta1 = 184.3,
-                 theta2 = 29739,
-                 m50 = 1129,
-                 m95 = 1344)
-
-sel_pars_school <- list(mesh = seq(4, 9, 0.01),
-                       
-                        theta1 = 192,
-                        theta2 = 67595,
-                        m50 = 1421,
-                        m95 = 1488)
-
-peak_sel <- function(len, df){
-  1 - (len / (df$rho * df$eta))^df$rho * exp(df$rho - len / df$eta)
-}
-
-sel_df <- cross_df(sel_pars_gummy) %>% 
-  mutate(eta = -0.5 * (theta1 * mesh - sqrt(theta1^2 * mesh^2 + 4 * theta2))) %>% 
-  mutate(rho = (theta1 * mesh) / eta) %>% 
-  mutate(id = row_number()) %>%
-  group_by(id) %>% nest(.key = "df")  %>% 
-  mutate(peak_sel = map_dbl(df, ~ optimize(peak_sel, lower = 0, upper = 2500, df = .)$minimum)) %>% 
-  unnest() %>% 
-  mutate(peak_sel = round(peak_sel, 0)) %>% 
-  mutate(P = (1 / (1 + exp(-log(19) * ((peak_sel - m50) / (m95 - m50)))))) %>% 
-  mutate(P = round(P, 2))
-
-# Selectivity parameters (Kirkwood and Walker 1986)
-
-# Housekeeping ----------------------------------------------------------------
-
-source(here::here("code", "setup.R"))
-source(here::here("code", "functions.R"))
-
-compile("code/logistic3.cpp")
-dyn.load(dynlib("code/logistic3"))
-compile("code/logistic2.cpp")
-dyn.load(dynlib("code/logistic2"))
-compile("code/logistic2maternal.cpp")
-dyn.load(dynlib("code/logistic2maternal"))
-
-# Parameter section  ----------------------------------------------------------
-
-sim_pars_mustelus <- list(
-  species = "Gummy shark",
-  
-  # Maximum age
-  AMAX = 20, 
-  # Natural mortality (Max age reported in Moulton et al 1992, Walker 1992 M)
-  M = 0.1970,
[…]

### 2024-06-23 code/simulation.R

@@ -111,13 +111,7 @@ data <-
   mutate(iteration = row_number()) %>%
   ungroup() %>% 
   # Generate simulated length data
-  mutate(data = map2(pars, id, generate_data))
-
-write_rds(data, "data/generated_data.rds")
-data <- read_rds("data/generated_data.rds")
-
-
-data <- data %>%
+  mutate(data = map2(pars, id, generate_data)) %>%
   # Simulate maternal data
   mutate(data = map2(pars, data, generate_maternal_data)) %>% # How many maternal females are there in the simulated data
   mutate(n_maternal = map_dbl(data, ~ sum(.$z))) %>%
@@ -161,21 +155,6 @@ data <- data %>%
     # 50 % interval coverage
     mutate(int_coverage = ifelse(par_true >= lower & par_true <= upper, TRUE, FALSE)) %>%
     select(-xmin, -xmax, -data)
-
-write_rds(data, here("data", "simulations-school-7.Rds"))
-
-
-data <- read_rds("data/simulations-gummy.rds")
-data1 <- read_rds("data/simulations-school-1.rds")
-data2 <- read_rds("data/simulations-school-2.rds")
-data3 <- read_rds("data/simulations-school-3.rds")
-data4 <- read_rds("data/simulations-school-4.rds")
-data5 <- read_rds("data/simulations-school-5.rds")
-data6 <- read_rds("data/simulations-school-6.rds")
-data7 <- read_rds("data/simulations-school-7.rds")
-
-data <- rbind(data, data1) %>% rbind(data2) %>% rbind(data3) %>% 
-  rbind(data4) %>% rbind(data5) %>% rbind(data6) %>% rbind(data7)
 
 write_rds(data, here("data", "simulations.Rds"))
 

### 2024-06-23 run_all.R

@@ -21,17 +21,3 @@ render("rmds/tables.RmD", output_format = "all", output_file = "tabs.pdf", outpu
 
 # Rmarkdown template
 # http://labrtorian.com/2019/08/26/rmarkdown-template-that-manages-academic-affiliations/
-
-# TODO 
-# - Bubley spiny dogfish - inconsistent definition (any with pups or large developing follicles)
-# - Cotton Squalus mitsikuri uses pregnancy ogive as maternity ogive
-# - Fujimani blue shark
-# - Rochowski paper
-# - Corro-Espinosa rhizoprionodon
-# - mustelus henlei (file:///Users/avh/Downloads/2805-Article%20Text-420420528-1-10-20180628.pdf)
-# - Squalus blainvillei (https://www-publish-csiro-au.libproxy.murdoch.edu.au/mf/pdf/MF19372)
-
-# Notes
-# - https://www.jstor.org/stable/2531207
-# - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5978598/
-# - https://www.researchgate.net/profile/Gunter_Maris/publication/237079179_On_Interpreting_the_Model_Parameters_for_the_Three_Parameter_Logistic_Model/links/5661554508aebae678aa7970.pdf

### 2024-06-24 code/empirical.R
stuff

@@ -15,8 +15,10 @@ dyn.load(dynlib("code/logistic3"))
 # and maternal status was assigned for each individual. Code for this step
 # and the original data files are available on request. 
 
+
+
 # Tidy up data 
-data <- read_rds(here("data", "empirical-plumbeus.Rds")) %>%
+data <- read_csv(here("data", "empirical-plumbeus.csv"))  %>%
   select(-source)
 
 # Run analyses ----------------------------------------------------------------
@@ -221,3 +223,4 @@ summary <- rbind(summary1, summary2, summary3) %>%
   arrange(delta_AIC)
 
 write_rds(summary, here("data", "empirical-summary.Rds"))
+

### 2024-08-16 README.qmd
delete
repo_prep
stuff
readme

@@
 
-# Example
+## Alastair Harry, Ivy Baremore, Andrew Piercy
+
+Published scientific manuscript:
+
+Harry AV, Baremore IE & Piercy AN (2024) Quantifying maternal reproductive output of chondrichthyan fishes. *Canadian Journal of Fisheries and Aquatic Sciences*
+
+![](https://mc06.manuscriptcentral.com/societyimages/cjfas-pubs/scholarOne-CJFAS_en_US.png)
+
+Repository structure
+
+fs::dir_tree()
+
+# Empirical example
 
@@
 
-data <- read_rds(here::here("data", "empirical-plumbeus.Rds"))
+data <- read_csv(here::here("data", "empirical-plumbeus.csv"))
 

### 2024-08-16 rmds/manuscript.Rmd
readme

@@
 empirical <- read_rds(here("data", "empirical-summary.Rds"))
-plumbeus <- read_rds(here("data", "empirical-plumbeus.Rds"))
+plumbeus <- read_csv(here("data", "empirical-plumbeus.csv"))
 
@@
 
+# Funding statement
+
+The authors declare no specific funding for this work.
+
 # References

### 2024-08-25 README.qmd
stuff

@@
 
-data <- read_csv(here::here("data", "empirical-plumbeus.csv"))
+data <- read_csv(here::here("data", "empirical-plumbeus.csv")) %>% 
+  select(-source)
 
@@
 
-## Run model
+## 3 Parameter logistic model
 
-Compile 3 parameter logistic model
+Compile the 3 parameter logistic model. The model
 
@@
 ```
+
+```{r}
+
+# Parameter list
+pars <- list(m50 = 160, m95 = 170, c = 0.5)
+
+# Create objective function
+obj <- MakeADFun(
+  data = data,
+  parameters = pars,
+  DLL = "logistic3"
+)
+
+# Optimize
+opt <- nlminb(obj$par, obj$fn, obj$gr, lower = c(0, 0, 0), upper = c(Inf, Inf, 1))
+```

### 2024-08-25 rmds/manuscript.tex
stuff

@@
 When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased.
-This is seen in the positive correlation between \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2).
+This is seen in the positive correlation between \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2).
 With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote.
-As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\).
-Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).
+As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\widehat{P_{Max}}\).
+Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).
 
@@
 While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6).
-For most simulations, interval coverage for \(\hat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size.
-In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\).
+For most simulations, interval coverage for \(\widehat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size.
+In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\).
 
@@
 Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition.
-Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2).
-Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7).
+Using the 3PLF-estimated method, the maximum likelihood estimate for \(\widehat{P_{Max}}\) was 0.48 (Table 2).
+Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\widehat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7).
 For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\).
@@
 
+\section{Funding statement}\label{funding-statement}
+
+The authors declare no specific funding for this work.
+
 \section*{References}\label{references}
@@
 2019.
-Presence of a resting population of female porbeagles ({Lamna} nasus), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
-Fish.
-B-noaa.
+Presence of a resting population of female porbeagles (\emph{{Lamna} nasus}), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}.
+Fish.
+Bull.
 \textbf{117}.
@@
 2021.
-Modernising fish and shark growth curves with bayesian length-at-age models.
+Modernising fish and shark growth curves with {Bayesian} length-at-age models.
 Plos One \textbf{16}(2): 1--21.
@@
 Accounting for unobserved population dynamics and aging error in close{-}kin mark{-}recapture assessments.
-Ecology and Evolution \textbf{14}(2).
+Ecol.
+Evol.
+\textbf{14}(2).
 doi: \href{https://doi.org/10.1002/ece3.10854}{10.1002/ece3.10854}.

### 2024-09-23 README.qmd
readme

@@
 
-Load data for sandbar shark maternity data set. Rows are data from individuals sharks x is fork length (cm) and z is maternal status (0 = non maternal, 1 = maternal).
+Load sandbar shark maternity data set. Rows are data from individuals sharks.
 
@@
 
+```
+
+*x* is fork length (cm) and *z* is maternal status (0 = non maternal, 1 = maternal).
+
+```{r}
+head(data)
+```
+
+```{r}
 data %>% 
@@
 
-Bin data into suitable length categories for plotting.
+Bin data into suitable length categories for plotting. Add marginal rug plots to show individual data points.
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include=TRUE}
 brks = seq(40, 220, 10)
@@
 ggplot()  + 
-  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
-  ylim(0, 1) +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
-```
-
-Add marginal rug plots to show individual data points.
-
-```{r}
-ggplot()  + 
   geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
@@
   labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
-
 ```

### 2024-09-23 rmds/figures.Rmd
changes
readme

@@
   pdf_document:
+    keep_tex: true
     latex_engine: "pdflatex"
+
 ---
@@
 
-```{r fig.width = 7, fig.height= 5.5}
+```{r fig.width = 7.2, fig.height= 5.5}
 data <- read_rds(here("data", "simulations-pivot.Rds")) %>% 

### 2024-09-24 README.qmd
README

@@
 
-Load sandbar shark maternity data set. Rows are data from individuals sharks.
+Load sandbar shark maternity data set.
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include=TRUE}
 library(tidyverse)
@@
 
-*x* is fork length (cm) and *z* is maternal status (0 = non maternal, 1 = maternal).
+Rows are data from individuals sharks. *x* is fork length (cm) and *z* is maternal status (0 = non maternal, 1 = maternal).
 
@@
 ```
+
+The dataset consists of maturity and maternity at length data for 1087 female sandbar sharks, 640 of which are mature, and 206 of which were in maternal condition.
 
@@
 
-Bin data into suitable length categories for plotting. Add marginal rug plots to show individual data points.
+Bin data into suitable length categories for plotting, in this case 10cm length bins work well. Marginal rug plots are also helpful for displaying the distribution of individual data points.
 
@@
 
-## 3 Parameter logistic model
+## 3 parameter logistic maternity function (3PLF)
 
-Compile the 3 parameter logistic model. The model
+To run the 3PLF model, load the `TMB` library, compile the `C++` file and load the compiled file.
 
@@
 
+Initialize a list of parameters for the model with their starting values, then construct the objective function to be minimised using R.
+
 ```{r}
-
 # Parameter list

### 2024-09-29 README.qmd
readme

@@
 
-Harry AV, Baremore IE & Piercy AN (2024) Quantifying maternal reproductive output of chondrichthyan fishes. *Canadian Journal of Fisheries and Aquatic Sciences*
+Harry AV, Baremore IE & Piercy AN (2024) Quantifying maternal reproductive output of chondrichthyan fishes.
+*Canadian Journal of Fisheries and Aquatic Sciences.* <https://doi.org/10.1139/cjfas-2024-0031>
 
@@
 
-Repository structure
+Repository structure:
 
+```{r, include = TRUE, eval=TRUE, echo=FALSE}
 fs::dir_tree()
+```
 
@@
 
-```{r}
+```{r, include = TRUE, eval=TRUE, echo=FALSE}
+data %>% 
+  filter(!is.na(y)) %>% 
+  ggplot() + 
+  geom_histogram(aes(x = x, fill = as.factor(y)), binwidth = 5, col = "black") +
+  labs(x = "Fork length (cm)", y = "Proportion in mature condition", title = "Length distribution showing female maturity stage") + 
+  guides(fill = guide_legend(title = "Maturity stage")) + 
+  theme(legend.position = c(0.2, 0.75))
+```
+
+```{r, include = TRUE, eval=TRUE, echo=FALSE}
 data %>% 
@@
   geom_histogram(aes(x = x, fill = as.factor(z)), binwidth = 5, col = "black") +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition") + 
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length distribution showing female maternity stage") + 
   guides(fill = guide_legend(title = "Maternal stage")) + 
-  theme(legend.position = "top")
+  theme(legend.position = c(0.2, 0.75))
 
@@
 Bin data into suitable length categories for plotting, in this case 10cm length bins work well. Marginal rug plots are also helpful for displaying the distribution of individual data points.
+
+```{r message=FALSE, warning=FALSE, include=TRUE}
+brks = seq(40, 220, 10)
+data_binned <- data %>%
+  filter(!is.na(y)) %>% 
+  mutate(x_bin = findInterval(x, brks)) %>% 
+  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
+  group_by(x_bin) %>%
+  summarise(p = sum(y)/ n())
+
+ggplot()  + 
+  geom_rug(data = filter(data, y == 1), aes(x = x), sides = "t") +
+  geom_rug(data = filter(data, y == 0), aes(x = x), sides = "b") +
+  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in mature condition", title = "Length at maturity data for female sandbar sharks") 
+```
 
@@
   ylim(0, 1) +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length at maternity data for female sandbar sharks") 
 ```
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, results="hide"}
 library(TMB)
@@
 Initialize a list of parameters for the model with their starting values, then construct the objective function to be minimised using R.
+Next use `nlminb` to optimize the model.
+`nlminb` takes as its arguments a vector of the parameters to be estimated, the function to be optimized, an analytical gradient vector that aids with optimization, and vectors of maximum and minimum allowable values for the model parameters.
 
-```{r}
+```{r message=FALSE, warning=FALSE, results = "hide"}
 # Parameter list
@@
 ```
+
+To check whether the model has successfully converged, run the following:
+
+```{r}
+opt$convergence
+```
+
+The maximum likelihood values of parameters estimated by `TMB` can be obtained as follows:
+
+```{r }
+opt$par
+```
+
+The estimated length at 50% maternity ($L_{50}$) and length at 95% maternity ($L_{95}$) are 160 and 174cm, respectively.
+$L_{50}$ is estimated length at which half of the females in the population were in maternal condition, meaning they would have given birth and contributed to recruitment within that year.
+The final parameter is the asymptotic of the logistic function, $P_{Max}$, which is the maximum proportion of females in a given year in maternal condition, in this case 0.48 (or 48% of the population).

### 2024-09-29 code/simulation.R

@@ -113,7 +113,8 @@ data <-
   # Generate simulated length data
   mutate(data = map2(pars, id, generate_data)) %>%
   # Simulate maternal data
-  mutate(data = map2(pars, data, generate_maternal_data)) %>% # How many maternal females are there in the simulated data
+  mutate(data = map2(pars, data, generate_maternal_data)) %>% 
+  # How many maternal females are there in the simulated data
   mutate(n_maternal = map_dbl(data, ~ sum(.$z))) %>%
   # Are there zero maternal females?
   mutate(no_maternal = ifelse(n_maternal == 0, 1, 0)) %>% 

### 2024-10-12 README.qmd
readme

@@
 
-Load sandbar shark maternity data set.
+Start by loading the example data.
+The dataset is an amalgamation of data from two studies on the reproductive biology of sandbar sharks by [Baremore and Hale (2012)](https://doi.org/10.1080/19425120.2012.700904) and [Piercy et al.
+(2016)](https://doi.org/10.1111/jfb.12945).
 
@@
 
-Rows are data from individuals sharks. *x* is fork length (cm) and *z* is maternal status (0 = non maternal, 1 = maternal).
+In this analysis, the fork length (FL) of the shark is the independent variable, and is continuous number ranging from `r min(data$x)`cm to `r max(data$x)`cm.
+Maturity stage and maternity stage are the dependent variables and are binary (either 0 for non mature / non maternal or 1 for mature / maternal).
+Rows in the data are data from individuals sharks.
+x is FL, y is maturity stage, and z is maternity stage.
 
@@
 
-The dataset consists of maturity and maternity at length data for 1087 female sandbar sharks, 640 of which are mature, and 206 of which were in maternal condition.
+The data consist of maturity and maternity at length data for 1087 female sandbar sharks, 640 of which are mature, and 206 of which were in maternal condition.
 
-```{r, include = TRUE, eval=TRUE, echo=FALSE}
+```{r, include = TRUE, eval=TRUE, warning=FALSE, echo=FALSE}
 data %>% 
@@
   guides(fill = guide_legend(title = "Maturity stage")) + 
-  theme(legend.position = c(0.2, 0.75))
+  theme(legend.position = "inside", legend.position.inside = c(0.2, 0.75))
 ```
 
-```{r, include = TRUE, eval=TRUE, echo=FALSE}
+```{r, include = TRUE, eval=TRUE, warning = FALSE, echo=FALSE}
 data %>% 
@@
   guides(fill = guide_legend(title = "Maternal stage")) + 
-  theme(legend.position = c(0.2, 0.75))
+  theme(legend.position = "inside", legend.position.inside = c(0.2, 0.75))
 
@@
 
-Bin data into suitable length categories for plotting, in this case 10cm length bins work well. Marginal rug plots are also helpful for displaying the distribution of individual data points.
+To visualize the binary response variable, it can be helpful to bin the explanatory variable (FL) into length categories.
+In this case 10cm length bins work well.
+Marginal rug plots are also helpful for displaying the distribution of individual data points.
 
@@
 
-To run the 3PLF model, load the `TMB` library, compile the `C++` file and load the compiled file.
+To run the 3PLF model, begin by loading the `TMB` library, compile the `C++` file and load the compiled file.
 

### 2024-10-25 README.qmd
readme

@@
 
-To visualize the binary response variable, it can be helpful to bin the explanatory variable (FL) into length categories.
+To visualize the binary response variable, it can be helpful to divide the explanatory variable (FL) into discrete length intervals.
 In this case 10cm length bins work well.
@@
 $L_{50}$ is estimated length at which half of the females in the population were in maternal condition, meaning they would have given birth and contributed to recruitment within that year.
-The final parameter is the asymptotic of the logistic function, $P_{Max}$, which is the maximum proportion of females in a given year in maternal condition, in this case 0.48 (or 48% of the population).
+The final parameter is the asymptote of the logistic function, $P_{Max}$, which is the maximum proportion of females in a given year in maternal condition, in this case 0.48 (or 48% of the population).
+
+The estimated proportion of females in maternal condition at a given length can then be calculated using the maximum likelihood values of the parameters from the fitted model.
+
+```{r}
+# Length at 50% maternity
+m50 <- opt$par[[1]]
+
+# Length at 95% maternity
+m95 <- opt$par[[2]]
+
+# PMAX
+pmax <- opt$par[[3]]
+
+# 
+newdata <- tibble(len = min(data$x):max(data$x)) %>% 
+    mutate(matern = (pmax / (1 + exp(-log(19) * ((len - m50) / (m95 - m50))))))
+
+ggplot()  + 
+  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
+  geom_line(data = newdata, aes(x= len, y = matern)) + 
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length at maternity data for female sandbar sharks")
+
+```

### 2024-10-25 README.qmd
readme

@@
   ylim(0, 1) +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length at maternity data for female sandbar sharks")
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Observed and predicted length at maternity for female sandbar sharks")
 

### 2024-11-09 README.qmd
qmd

--- 2024-10-25 README.qmd
@@
 
-Repository structure:
+# Example
 
-```{r, include = TRUE, eval=TRUE, echo=FALSE}
-fs::dir_tree()
-```
-
-# Empirical example
+Instructions and code for running the model described in this paper are outlined below.
 
@@
 
-Start by loading the example data.
+Start by loading the data used in the empirical example in the paper.
 The dataset is an amalgamation of data from two studies on the reproductive biology of sandbar sharks by [Baremore and Hale (2012)](https://doi.org/10.1080/19425120.2012.700904) and [Piercy et al.
@@
 
+# Set theme for plotting data and models
 theme_set(theme_bw())
 
-data <- read_csv(here::here("data", "empirical-plumbeus.csv")) %>% 
+# Get url of empirical dataset used in the study
+url <- RCurl::getURL("https://raw.githubusercontent.com/alharry/maternity-function/refs/heads/main/data/empirical-plumbeus.csv")
+
+# Load data
+data <- read_csv(url) |>
   select(-source)
-
 ```
@@
 ```{r, include = TRUE, eval=TRUE, warning=FALSE, echo=FALSE}
-data %>% 
-  filter(!is.na(y)) %>% 
+data |>
+  filter(!is.na(y)) |>
   ggplot() +
@@
 ```{r, include = TRUE, eval=TRUE, warning = FALSE, echo=FALSE}
-data %>% 
+data |>
   ggplot() +
@@
   theme(legend.position = "inside", legend.position.inside = c(0.2, 0.75))
-
 ```
@@
 ```{r message=FALSE, warning=FALSE, include=TRUE}
-brks = seq(40, 220, 10)
-data_binned <- data %>%
-  filter(!is.na(y)) %>% 
-  mutate(x_bin = findInterval(x, brks)) %>% 
-  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
-  group_by(x_bin) %>%
-  summarise(p = sum(y)/ n())
+brks <- seq(40, 220, 10)
+data_binned <- data |>
+  filter(!is.na(y)) |>
+  mutate(x_bin = findInterval(x, brks)) |>
+  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) |>
+  group_by(x_bin) |>
+  summarise(p = sum(y) / n())
 
@@
 ```{r message=FALSE, warning=FALSE, include=TRUE}
-brks = seq(40, 220, 10)
-data_binned <- data %>%
-  mutate(x_bin = findInterval(x, brks)) %>% 
-  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
-  group_by(x_bin) %>%
-  summarise(p = sum(z)/ n())
+brks <- seq(40, 220, 10)
+data_binned <- data |>
+  mutate(x_bin = findInterval(x, brks)) |>
+  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) |>
+  group_by(x_bin) |>
+  summarise(p = sum(z) / n())
 
@@
 
-## 3 parameter logistic maternity function (3PLF)
+## 3 parameter logistic function - estimated $P_{MAX}$
 
-To run the 3PLF model, begin by loading the `TMB` library, compile the `C++` file and load the compiled file.
+To run the 3PLF-free model described in the paper, begin by loading the `TMB` library, compile the `C++` file and load the compiled file.
 
@@
 #
-newdata <- tibble(len = min(data$x):max(data$x)) %>% 
+newdata <- tibble(len = min(data$x):max(data$x)) |>
   mutate(matern = (pmax / (1 + exp(-log(19) * ((len - m50) / (m95 - m50))))))
@@
   geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") +
-  geom_line(data = newdata, aes(x= len, y = matern)) + 
+  geom_line(data = newdata, aes(x = len, y = matern)) +
   ylim(0, 1) +
-  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length at maternity data for female sandbar sharks")
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "3PLF-free", subtitle = "Observed and predicted length at maternity for female sandbar sharks")
+```
 
+## 3 parameter logistic function - fixed $P_{MAX}$
+
+To run the 3PLF-fixed model, start by creating a new objective function.
+For this model, $P_{Max}$ is chosen by the user and assumed to be known without error.
+The code for the objective function is slightly modified to account for this.
+
+```{r message=FALSE, warning=FALSE, results = "hide"}
+# Chose the pre-specifed (fixed) value of PMAX
+prop_pregnant <- 0.5
+
+# Parameter list
+pars2 <- list(m50 = 160, m95 = 170, c = prop_pregnant)
+
+# Create objective function
+obj2 <- MakeADFun(
+  data = data,
+  parameters = pars2,
+  DLL = "logistic3",
+  map = list(c = factor(NA))
+)
+
+# Optimize
+opt2 <- nlminb(obj2$par, obj2$fn, obj2$gr, lower = c(0, 0, 0), upper = c(Inf, Inf, 1))
 ```
+To check whether the model has successfully converged, run the following:
+
+```{r}
+opt2$convergence
+```
+
+The maximum likelihood values of parameters estimated by `TMB` can be obtained as follows:
+
+```{r}
+opt2$par
+```
+
+For the 3PLF-fixed method with $P_{Max}$ set to 0.5, the estimated length at 50% maternity ($L_{50}$) and length at 95% maternity ($L_{95}$) are 161 and 175cm, respectively.
+
+```{r }
+# Length at 50% maternity
+m50 <- opt2$par[[1]]
+
+# Length at 95% maternity
+m95 <- opt2$par[[2]]
+
+# PMAX (fixed value)
+pmax <- prop_pregnant
+
+# Plot model
+newdata <- tibble(len = min(data$x):max(data$x)) |>
+  mutate(matern = (pmax / (1 + exp(-log(19) * ((len - m50) / (m95 - m50))))))
+
+ggplot() +
+  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") +
+  geom_line(data = newdata, aes(x = len, y = matern)) +
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "3PLF-fixed", subtitle = "Observed and predicted length at maternity for female sandbar sharks")
+```
+
+## Confidence intervals
+Uncertainty in the model predictions and fitted parameters can be quantified using bootstrap resampling. Start by resampling the original dataset 1000 times and re-fitting the model.
+
+```{r message=FALSE, warning=FALSE}
+library(rsample)
+
+output <- bootstraps(data, 1000) |>
+  mutate(obj = map(splits, ~ MakeADFun(analysis(.), parameters = pars, DLL = "logistic3", silent = TRUE))) |>
+  mutate(opt = map(obj, ~ nlminb(start = .$par, objective = .$fn, gradient = .$gr))) |>
+  mutate(pars = map(opt, ~ .$par)) |>
+  mutate(par_name = map(opt, ~ .$par |>names()))
+```
+
+Approximate 95% confidence intervals around the maximum likelihood parameter estimates for the maternity model can then be obtained using the quantile function.
+
+```{r message=FALSE, warning=FALSE}
+summary <- output |>
+  unnest(c(pars, par_name)) |>
+  group_by(par_name) |>
+  summarise(
+    lower = quantile(pars, 0.05 / 2),
+    upper = quantile(pars, 1 - 0.05 / 2)
+  ) |>
+  cbind(est = c(opt$par[3], opt$par[-3])) |>
+  select(par_name, est, lower, upper) |>
+  mutate(across(where(is.double), round, digits = 2))
+
+rownames(summary) <- NULL
+
+knitr::kable(summary)
+```
+
+Confidence intervals for predicted values of maternity at length can also be calculated from the boostrapped data.
+
+```{r message=FALSE, warning=FALSE}
+preds <- output |>
+  mutate(x = map(splits, ~ seq(min(analysis(.)$x) |>floor(), max(analysis(.)$x) |>ceiling(), 1))) |>
+  mutate(y = map2(pars, x, ~ (.x[3] / (1 + exp(-log(19) * ((.y - .x[1]) / (.x[2] - .x[1]))))))) |>
+  dplyr::select(id, x, y) |>
+  unnest(cols = c(x, y)) |>
+  group_by(x) |>
+  summarise(lower = quantile(y, 0.05), upper = quantile(y, 0.975)) |>
+  mutate(y = (opt$par[3] / (1 + exp(-log(19) * ((x - opt$par[1]) / (opt$par[2] - opt$par[1]))))))
+
+ggplot() +
+  geom_ribbon(data = preds, aes(x = x, ymin = lower, ymax = upper), fill = "grey") +
+  geom_line(data = newdata, aes(x = len, y = matern)) +
+  ylim(0, 1) +
+  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "3PLF-free", subtitle = "Predicted length at maternity for female sandbar sharks and 95% confidence intervals")
+```

### 2024-11-09 code/setup.R

@@ -9,7 +9,7 @@ set.seed(3)
 # Load libraries
 libs<-c("MASS", "tidyverse", "knitcitations", "bibtex", "lubridate",
         "forcats", "here", "TMB", "broom", "rsample", "cowplot",
-        "grid", "gridExtra", "xtable")
+        "grid", "gridExtra", "xtable", "rsample")
 
 lapply(libs, library, character.only = TRUE)
 
